### 2021-11-23 PW13Paper.tex
PW13

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PREAMBLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% The following two commands will generate a PDF that follows all the requirements for submission
% and peer review.  Uncomment these commands to generate this output (and comment out the two lines below.)
%
% DOUBLE SPACE VERSION FOR SUBMISSION TO THE AMS
\documentclass[11pt]{article}
\usepackage{ametsoc}
% \usepackage{ametsoc2col}
%\linenumbers
\graphicspath{{/Users/jklymak/PW13/figs/mvp/}}
\usepackage[plain]{fancyref}
\usepackage{xcolor}
\usepackage{lineno}
\newcommand{\tempS}[1]{}
\newcommand{\twowidth}[0]{6.4in}
\newcommand{\onewidth}[0]{3.2in}
\newcommand{\mn}[1]{{\sc #1}}
\DeclareGraphicsExtensions{%
    .pdf,.png}
%
% The following two commands will generate a single space, double column paper that closely
% matches an AMS journal page.  Uncomment these commands to generate this output (and comment
% out the two lines above. FOR AUTHOR USE ONLY. PAPERS SUBMITTED IN THIS FORMAT WILL BE RETURNED
% TO THE AUTHOR for submission with the correct formatting.
%
% TWO COLUMN JOURNAL PAGE LAYOUT FOR AUTHOR USE ONLY
%%%%\documentclass[10pt]{article}
%%%%\usepackage{ametsoc2col}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%
% Enter your Abstract here
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\linenumbers
\newcommand{\myabstract}{}
%
\begin{document}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE
%
% Enter your TITLE here
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\textbf{\large{Evidence for cross-shelf exchange driven by a submarine canyon}}}
%
% Author names, with corresponding author information.
% [Update and move the \thanks{...} block as appropriate.]
%
\author{\textsc{Jody M. Klymak,}
				\thanks{\textit{Corresponding author address:}
				Jody Klymak, University of Victoria, Victoria, BC, Canada
				\newline{E-mail: jklymak@uvic.ca}}\quad
				\\
\textit{\footnotesize{School of Earth and Ocean Sciences and Department of Physics and Astronomy, University of Victoria, Victoria, Canada}}
}
%
% Formatting done here...Authors should skip over this.  See above for abstract.
\ifthenelse{\boolean{dc}}
{
\twocolumn[
\begin{@twocolumnfalse}
\amstitle

% Start Abstract (Enter your Abstract above.  Do not enter any text here)
\begin{center}
\begin{minipage}{13.0cm}
\begin{abstract}
	\myabstract
	\newline
	\begin{center}
		\rule{38mm}{0.2mm}
	\end{center}
\end{abstract}
\end{minipage}
\end{center}
\end{@twocolumnfalse}
]
}
{
\amstitle
\begin{abstract}
\myabstract
\end{abstract}
\newpage
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MAIN BODY OF PAPER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Cross-shelf exchange is important to the health and productivity of continental shelf regions, where offshore nutrients and oxygenated water is exchanged with nutrient and oxygen depleted near-shore water.  Cross-shelf transport requires a level of ageostrophic flow since balanced flow will tend to follow topographic contours, often providing a barrier to lateral exchange.  Mechanisms for cross-shelf exchange include internal waves and instabilities in shelf-break fronts.  Most-dramatically, however, topography can drive cross-slope exchanges, often seen most clearly in surface

%\citet{klymaketal10a}

\section{Site and Methods}
\label{sec:Site}

Data collection was between 21 and 30 August, 2013, aboard the \emph{R/V Falkor}.  The study site was the southern portion of the Vancouver Island Shelf (\fref{fig:LocMapBoth}a), a particularly complicated region due to the bathymetry and varied forcing.

\begin{figure*}[htbp]
  \begin{center}

     \includegraphics[width=6.2in]{LocMapBoth.pdf}
    \caption{a) Study site on the Vancouver Island Shelf.  The blue arrow indicates the direction of the Vancouver Island Current.  The pink arrow indicates southward flow of the coastal upwelling current.  The dashed box indicates the approximate limits of the study area.  b) The study area with standard hydrographic casts indicated in yellow dots, and Moving Vessel Profiler casts as red dots.
      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProc.ipynb ;
        /Users/jklymak/PW13/figs/mvp//LocMapBoth.pdf}
      \label{fig:LocMapBoth} }
  \end{center}
\end{figure*}

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=6.2in]{SpiceO2264}
    \caption{Spatial overview of a) local spice (temperature anomaly on an isopycnal: see text), and b) oxygen saturation on the $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.  Grey cross-slope lines are cross sections indicated in \fref{fig:CrossSectionsSpice}.  Along-slope grey lines are every 20 km in the cross-slope direction, with $x=0\ \mathrm{km}$ near the 100-m isobath at the north end of the observation area.

      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProcNew.ipynb ;
        /Users/jklymak/PW13/figs/mvp//SpiceO2264.pdf}
      \label{fig:SpiceO2264} }
  \end{center}
\end{figure*}

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=6.2in]{CrossSectionsSpice}
    \caption{
      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProcNew.ipynb ;
        /Users/jklymak/PW13/figs/mvp//CrossSectionsSpice.pdf}
      \label{fig:CrossSectionsSpice} }
  \end{center}
\end{figure*}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=3in]{TsLineDefn}
    \caption{
      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProcNew ;
        /Users/jklymak/PW13/figs/mvp//TsLineDefn.pdf}
      \label{fig:TsLineDefn} }
  \end{center}
\end{figure}

%\begin{appendix}

%\subsection{First appendix secondary heading}
%
%\subsection{Second appendix secondary heading}
%
%\subsubsection{First appendix tertiary heading}
%
%\subsubsection{Second appendix tertiary heading}
%
%\paragraph{First appendix quaternary heading}
%
%\paragraph{Second appendix quaternary heading}
%
%\end{appendix}

% Create a bibliography directory and place your .bib file there.
% -REMOVE ALL DIRECTORY PATHS TO REFERENCE FILES BEFORE SUBMITTING TO THE AMS FOR PEER REVIEW
\ifthenelse{\boolean{dc}}
{}
{\clearpage}
\bibliographystyle{ametsocjmk}
\bibliography{main}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIGURES-REMOVE ALL DIRECTORY PATHS TO FIGURE FILES BEFORE SUBMITTING TO THE AMS FOR PEER REVIEW
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
\end{document}

### 2022-02-05 PW13Paper.tex
DOC: rough version to coauthors

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% PREAMBLE
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
 %
-% The following two commands will generate a PDF that follows all the requirements for submission
-% and peer review.  Uncomment these commands to generate this output (and comment out the two lines below.)
+% This file includes commands and instructions
+% given in the order necessary to produce a final output that will
+% satisfy AGU requirements, including customized APA reference formatting.
 %
-% DOUBLE SPACE VERSION FOR SUBMISSION TO THE AMS
-\documentclass[11pt]{article}
-\usepackage{ametsoc}
-% \usepackage{ametsoc2col}
-%\linenumbers
-\graphicspath{{/Users/jklymak/PW13/figs/mvp/}}
-\usepackage[plain]{fancyref}
-\usepackage{xcolor}
+% You may copy this file and give it your
+% article name, and enter your text.
+%
+%
+% Step 1: Set the \documentclass
+%
+%
+
+%% To submit your paper:
+\documentclass[draft]{agujournal2019}
+\usepackage{url} %this package should fix any errors with URLs in refs.
 \usepackage{lineno}
-\newcommand{\tempS}[1]{}
-\newcommand{\twowidth}[0]{6.4in}
-\newcommand{\onewidth}[0]{3.2in}
-\newcommand{\mn}[1]{{\sc #1}}
+% \graphicspath{{/Users/jklymak/Dropbox/PW13/satellite/doc/},
+%               {/Users/jklymak/Dropbox/PW13/figs/mvp/},
+%               {/Users/jklymak/Dropbox/PW13/analysis/mvp/doc/},
+%               {/Users/jklymak/Dropbox/PW13/LaPerouse13/doc/},
+%               {/Users/jklymak/Dropbox/PW13/auxdata/doc/}}
+
+\graphicspath{{./figs/}}
 \DeclareGraphicsExtensions{%
               .pdf,.png}
+
+\usepackage[plain]{fancyref}
+% \usepackage[]{hyperref}
+%\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
+\usepackage{soul}
+\linenumbers
+
+\draftfalse
+
+\journalname{JGR: Oceans}
+
+\begin{document}
+
+%% ------------------------------------------------------------------------ %%
+%  Title
 %
-% The following two commands will generate a single space, double column paper that closely
-% matches an AMS journal page.  Uncomment these commands to generate this output (and comment
-% out the two lines above. FOR AUTHOR USE ONLY. PAPERS SUBMITTED IN THIS FORMAT WILL BE RETURNED
-% TO THE AUTHOR for submission with the correct formatting.
+% (A title should be specific, informative, and brief. Use
+% abbreviations only if they are defined in the abstract. Titles that
+% start with general keywords then specific terms are optimized in
+% searches)
 %
-% TWO COLUMN JOURNAL PAGE LAYOUT FOR AUTHOR USE ONLY
-%%%%\documentclass[10pt]{article}
-%%%%\usepackage{ametsoc2col}
+%% ------------------------------------------------------------------------ %%
+
+% Example: \title{This is a test title}
+
+\title{Separation of an upwelling current bounding the Juan de Fuca Eddy, off Vancouver Island}
+
+%% ------------------------------------------------------------------------ %%
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ABSTRACT
+%  AUTHORS AND AFFILIATIONS
 %
-% Enter your Abstract here
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\linenumbers
-\newcommand{\myabstract}{}
+%% ------------------------------------------------------------------------ %%
+
+% Authors are individuals who have significantly contributed to the
+% research and preparation of the article. Group authors are allowed, if
+% each author in the group is separately identified in an appendix.)
+
+% List authors by first name or initial followed by last name and
+% separated by commas. Use \affil{} to number affiliations, and
+% \thanks{} for author notes.
+% Additional author notes should be indicated with \thanks{} (for
+% example, for current addresses).
+
+% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
+% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
+
+\authors{Jody M. Klymak\affil{1} and Everyone }
+
+% \affiliation{1}{First Affiliation}
+% \affiliation{2}{Second Affiliation}
+% \affiliation{3}{Third Affiliation}
+% \affiliation{4}{Fourth Affiliation}
+\affiliation{1}{School of Earth and Ocean Sciences and Department of Physcics \& Astronomy, University of Victoria, B.C.\ Canada}
+%(repeat as many times as is necessary)
+
+%% Corresponding Author:
+% Corresponding author mailing address and e-mail address:
+
+\correspondingauthor{J.\ Klymak}{jklymak@uvic.ca}
+
+%% Keypoints, final entry on title page.
+
+%  List up to three key points (at least one is required)
+%  Key Points summarize the main points and conclusions of the article
+%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
+
+% Example:
+% \begin{keypoints}
+% \item	List up to three key points (at least one is required)
+% \item	Key Points summarize the main points and conclusions of the article
+% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
+% \end{keypoints}
+
+\begin{keypoints}
+\item The shelfbreak current along Vancouver Island separates downstream of a submarine bank.
+\item Offshore water is drawn onto the shelf and forms a sharp semi-persisnte front with the Juan de Fuca eddy.
+\item The Juan de Fuca eddy shows evidence of long residence times, and little evidence of deep-water origin.
+\end{keypoints}
+
+%% ------------------------------------------------------------------------ %%
 %
-\begin{document}
+%  ABSTRACT and PLAIN LANGUAGE SUMMARY
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TITLE
+% A good Abstract will begin with a short description of the problem
+% being addressed, briefly describe the new data or analyses, then
+% briefly states the main conclusion(s) and how they are supported and
+% uncertainties.
+
+% The Plain Language Summary should be written for a broad audience,
+% including journalists and the science-interested public, that will not have
+% a background in your field.
 %
-% Enter your TITLE here
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{\textbf{\large{Evidence for cross-shelf exchange driven by a submarine canyon}}}
+% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
+% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
+% see http://sharingscience.agu.org/creating-plain-language-summary/)
 %
-% Author names, with corresponding author information.
-% [Update and move the \thanks{...} block as appropriate.]
+%% ------------------------------------------------------------------------ %%
+
+%% \begin{abstract} starts the second page
+
+\begin{abstract}
+
+\end{abstract}
+
+\section*{Plain Language Summary}
+
+The southern Vancouver Island continental shelf is very productive due to high nutrient input from the Strait of Juan de Fuca and Salish Sea estuarine system, and substantial cross-shelf transport due to the complicated topography.
+The estuarine water flows poleward, hugging the coast as the Vancouver Island current, and in the summer shelf water flows equatorward.
+Trapped between these two currents is the Juan de Fuca eddy, a stagnant region of water that occupies the shelf where it abruptly widens just north of the Strait of Juan de Fuca.
+
+Here we present intensive sampling of the Juan de Fuca eddy region.
+The observations show the eddy is low in oxygen, and we find has undergone substantial vertical and lateral mixing.
+This indicates that the waters have a surprisingly long residence time for a coastal feature, and that this region extends the depth of the water column.
+In contrast to previous literature we find that the water likely has low oxygen from being used for respiration rather than being pulled from deep in the California Undercurrent.
+
+The sampling also shows a remarkable flow separation of the equatorward shelf current  The current is shown to detach and is pushed offshore.
+Such events are readily seen in satellite imagery, but the results here indicate that the separation extends the depth of the water column on the shelf, and that this separation may be partially driven by the local bathymetry.
+The separation is a very strong cross-shelf exchange event, and likely transports substantial nutrient-rich coastal water offshore to drive productivity in the deeper ocean adjacent to the continental slope.
+
+%% ------------------------------------------------------------------------ %%
 %
-\author{\textsc{Jody M. Klymak,}
-				\thanks{\textit{Corresponding author address:}
-				Jody Klymak, University of Victoria, Victoria, BC, Canada
-				\newline{E-mail: jklymak@uvic.ca}}\quad
-				\\
-\textit{\footnotesize{School of Earth and Ocean Sciences and Department of Physics and Astronomy, University of Victoria, Victoria, Canada}}
-}
+%  TEXT
 %
-% Formatting done here...Authors should skip over this.  See above for abstract.
-\ifthenelse{\boolean{dc}}
-{
-\twocolumn[
-\begin{@twocolumnfalse}
-\amstitle
+%% ------------------------------------------------------------------------ %%
 
-% Start Abstract (Enter your Abstract above.  Do not enter any text here)
-\begin{center}
-\begin{minipage}{13.0cm}
-\begin{abstract}
-	\myabstract
-	\newline
-	\begin{center}
-		\rule{38mm}{0.2mm}
-	\end{center}
-\end{abstract}
-\end{minipage}
-\end{center}
-\end{@twocolumnfalse}
-]
-}
-{
-\amstitle
-\begin{abstract}
-\myabstract
-\end{abstract}
-\newpage
-}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Cross-shelf exchange is important to the health and productivity of continental shelf regions, where offshore nutrients and oxygenated water is exchanged with nutrient and oxygen depleted near-shore water.
-Cross-shelf transport requires a level of ageostrophic flow since balanced flow will tend to follow topographic contours, often providing a barrier to lateral exchange.
+Cross-shelf exchange is important to the health and productivity of continental shelf regions, where oxygenated water is exchanged with nutrient-rich, but oxygen-depleted nearshore water.
+Cross-shelf transport usually requires ageostrophic flow, since balanced flow will tend to follow topographic contours, often providing a barrier to lateral exchange \cite{brink16}.
 Mechanisms for cross-shelf exchange include internal waves and instabilities in shelf-break fronts.
-Most-dramatically, however, topography can drive cross-slope exchanges, often seen most clearly in surface
+Most-dramatically, however, topography can drive cross-slope exchanges, often seen most clearly in surface imagery, but occasionally observed \emph{in-situ} \cite{barthetal00}.
 
-%\citet{klymaketal10a}
+Here we present detailed observations from the southern Vancouver Island shelf, collected in summer 2013, and sampled with a rapid profiling vehicle equipped with a CTD and oxygen sensor, supplemented by traditional hydrographic surveys.
+The shelf has complicated bathymetry \fref{fig:LocMapBoth}, with a somewhat typical shelf poleward of our study site, but then widening in the lee of La Perouse Bank, and then hitting the Juan de Fuca canyon, equatorward.
+Water from the strait of Juan de Fuca flows poleward as a buoyant current that hugs the coast \cite{thomsonetal89, hickeyetal91}, while shelf water flows from poleward in the summer under upwelling conditions, both forced by local winds and via tele-connections with the long, homogenous shelves equatorward off Washington and Oregon \cite{hickeyetal91,thomsonkrassovski15,engidaetal16}.
+Trapped between these two currents is a region of relatively homogenous water that has been termed the Juan de Fuca eddy \cite{freelanddenman82,freelandmcintosh89,foremanetal07,macfadyenhickey10}.
+
+One of our findings below is that the water in the Juan de Fuca eddy is apparently quite old.
+The low oxygen found in the Juan de Fuca eddy has caused it to be inferred to be upwelled from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
+This inference was based on the oxygen being conservative, and hence having to come from close to the oxygen minimum zone found further offshore \cite{mackasetal87}.
+This lead to hypotheses that perhaps the eddy was fed by waters being drawn up a spur canyon of the Juan de Fuca canyon (often called the Tully Canyon) via ageostrophic transport from offshore to onshore due to the low pressure in the eddy center \cite{weaverhsieh87}.
+Below, we will argue that there is not substantial evidence of such transport, and that oxygen is likely low because of consumption due to respiration.
+
+A second important finding is that the water flowing equatorward along the shelf from upstream is all displaced from the shelf and pushed offshore and replaced by offshore water that floods the southern shelf.
+Such events are important as they tend to transport nutrient rich shelf water out into the open ocean, fueling offshore ecosystems.
+This cross-shelf exchange has not, to our knowledge, been remarked upon for the Vancouver Island shelf.
+Mesoscale separations have been observed further to the south off Cape Blanco \cite{barthetal00}, and was also inferred at other locations along the coast.
+Similar observations have been made elsewhere \cite[e.g.]{relvasbarton05}, though usually in the presence of a distinct coastal headland, a feature not found at our study location.
+The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics \cite{dalebarth01} leading to along-shore trapping of coastal waves,  wind stress curl variations \cite{castelaobarth07}, and induced relative vorticity due to stretching of parcels that inertially overshoot their initial isobaths due to a sudden change in downstream bathymetry.
+
+In this paper we present our observations from 2013 by describing our at-sea observation tools (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes, with a focus on the offshore side of the eddy region (\fref{sec:Observations}) where we consider the age of the eddy, the steadiness of the front separating the eddy and the offshore water, properties along the spur canyon that was believed to feed the eddy, and finally describe the large offshore separation of the shelf-break current.
+We discuss the origin of the eddy and the implications of the separating jet.
 
@@
 
-Data collection was between 21 and 30 August, 2013, aboard the \emph{R/V Falkor}.
+Observations from three summer cruises in 2013 are discussed here.
+Two were hydrographic surveys carried out on the \emph{CCGS Tully}, one in June, and the other in late September.
+The third was a finescale survey made in the Vancouver Island shelf between 21 and 30 August 2013, aboard the \emph{R/V Falkor}.
 The study site was the southern portion of the Vancouver Island Shelf (\fref{fig:LocMapBoth}a), a particularly complicated region due to the bathymetry and varied forcing.
+
+\begin{figure*}[htbp]
+  \includegraphics[width=5in]{LocMapBoth.pdf}
+  \caption{a) Study site on the Vancouver Island Shelf.
+The blue arrow indicates the direction of the Vancouver Island Current.
+The pink arrow indicates southward flow of the coastal upwelling current.
+The dashed box indicates the approximate limits of the study area.
+b) The study area with hydrographic casts from the La Perouse cruises along the LB and LC Lines are green dots, hydrographic casts during the Falkor cruise indicated in yellow dots, and Moving Vessel Profiler casts as red dots.
+The coordinate system used for this paper is shown with alternating grey and white bands at 10-km intervals.}
+  \label{fig:LocMapBoth}
+\end{figure*}
+
+At the south end of the study site is the Juan de Fuca canyon, which feeds dense water into Juan de Fuca Strait.
+This dense water is mixed with fresh water from the Fraser River at the sills and archipelagos further inland, and fluxes out the Strait again, where it turns poleward along Vancouver Island to form the Vancouver Island Current.
+The Juan de Fuca Canyon has a notable spur canyon, the Tully Canyon, that incises the shelf towards the north into the study site (\fref{fig:LocMapBoth}b).
+The rest of the shelf is punctuated by a series of banks and shallow basins.
+Notable is La Perouse bank, running along the shelf from the north, and moving isobaths offshore until it abruptly ends at around 48.5 N.
+The shelf break is characterized by a series of submarine canyons, in particular Nitnat Canyon and Barkley Canyon.
+
+Observations from the LB and LC lines were collected aboard the \emph{CCGS Tully} from 2013-05-30 to 2013-05-31, which we will call ``June'', and 2013-09-07 to 2013-09-09, which we will call ``September''.
+These lines span the depths from about 50 m to deep offshore.
+The data comes from a lowered Seabird 9-11 CTD, with an SBE 43 oxygen sensor.
+Oxygen data have been checked against bottle casts, and the CTD corrected for sensor offsets and thermal lags.
+
+Observations aboard the \emph{R/V Falkor} were carried out from 2013-08-21 to 2013-08-30, starting two weeks before the September Tully cruise.
+Data were collected with an ODIM Brooke-Ocean Moving Vessel Profiler, profiling to 200 m.
+The MVP was equipped with an AML Oceanographic CTD, and a Rinko Oxygen sensor with a 7-s response time foil.
+The MVP profiles at vertical fall speeds of $3\ \mathrm{m\,s^{-1}}$ to depths of 200 m or to within 5 m of the seafloor, whichever is shallower.
+Data collection took place while the ship cruised at speeds between 5 and 8 kts, usually at around 6 kts to enable fine horizontal spacing of the casts, with typical spacing less than 800 m.
+Data is collected during down- and up-casts, but is only reported for down-casts because the path through the water is highly asymmetric on the up-cast, whereas the downcast is largely vertical.
+The rapid speed of the profiling makes the oxygen measurements somewhat coarse, and probably biased due to the phase lag of the sensor, so we treat these largely qualitatively in this paper.
+
+Unfortunately, neither vessel had an operational acoustic Doppler profiler during the cruises with which to make current measurements.
+
+Winds during the cruises were typical for the west coast of Vancouver Island, with down-welling favorable winds during July and early August, and then intermittent winds before and during the Falkor cruise.
+Note that this locale is strongly affected by coastally trapped waves from further south, so upwelling usually persists despite local wind forcing \cite{thomsonkrassovski15, engidaetal16}.
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=5.5in]{LaPeWind}
+    \caption{
+      Wind from the La Perouse buoy \cite{DFOWind2013C46206}.
+The vertical direction is along-shelf (chosen as a heading of 315 N).
+The wind components has been low-pass filtered to one-day averages.
+The timing of the cruises discussed in the paper are shown as colored bands.
+      \label{fig:LaPeWind} }
+  \end{center}
+\end{figure*}
+
+\section{Observations}
+\label{sec:Observations}
+
+\subsection{Early and late summer broad-scale surveys}
+
+Hydrographic sections along the LB and LC lines highlight summer conditions on the Southern Vancouver Island Shelf, and indicate some of the features we are focusing on in this paper (\fref{fig:LaPerouse2013Ctd}).
+During both cruises and at both lines there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m to shallower than 85 m over the shelf.
+This dense water tends to be low in oxygen and cool.
+Further onshore, isopycnals tilt back down, and warm water fills the water column, evidence of the Vancouver Island Current that hugs the coast here.
+
+\begin{figure*}[htbp]
+  \begin{center}
+     \includegraphics[width=130mm]{LaPerouse2013Ctd}
+     \caption{Observations along LC and LB lines in Jun and Sep 2013.
+X is across-shelf distance in the coordinate system shown in \fref{fig:LocMapBoth}.
+Potential density is contoured every $0.2\ \mathrm{kg\,m^{-3}}$, with the $26.4\ \mathrm{kg\,m^{-3}}$.
+Potential temperature $\theta$, spice anomaly $\gamma$ and oxygen concentration $O_2$ are colored for each section.}
+     \label{fig:LaPerouse2013Ctd}
+  \end{center}
+\end{figure*}
+
+The ``Juan de Fuca Eddy'' region is along the LB Line, mostly inshore of 0 km.
+This water is cooler and lower in oxygen than along the same isopycnal further offshore.
+Also note that during the June cruise the water in this region was cooler and more oxygenated than later in the summer during the September cruise.
+Near the surface, the isopycnals are domed, but the low-oxygen anomaly extends as high as the thin surface mixed layer.
+
+In contrast, the water upstream, along the LC line, does not show as much drawdown of oxygen, and is somewhat warmer, at least offshore of La Perouse bank (\fref{fig:LaPerouse2013Ctd} top three rows).
+Water does not appear to make it over La Perouse bank into the basin onshore, or if it does, it does so intermittently.
+There appears to be some mixing of the deeper water and shallow water near $26.4\ \mathrm{kg\,m^{-3}}$, with the water along this isopycnal becoming cooler and lower in oxygen where it intersects the shelf.
+
+The $\theta$--$S$ properties show many of the features demonstrated above (\fref{fig:LaPerouse2013TS}a).
+Below $26.6\ \mathrm{kg\,m^{-3}}$ or so, the deep water masses are found offshore, and are largely homogenous.
+Shallower, there are distinct differences between the water found on the shelf and water found more offshore.
+These anomalies are hard to parse directly, so in this paper we use a spice anomaly, relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}).
+The line passes the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$, and spice anomaly for a given water sample at density $\sigma_{\theta}$ is $\gamma = \alpha \left(T - T_0\left(\sigma_{\theta}\right)\right) + \beta \left(S - S_0\left(\sigma_{\theta}\right)\right)$ where $T_0\left(\sigma_{\theta}\right)$ and $S_0\left(\sigma_{\theta}\right)$ are the temperature and salinity along the mixing line at the same density as the water sample, with the sign convention that positive is warmer and saltier than data along the mixing line.
+It is hard to see from the $\theta$--$S$ plot in \fref{fig:LaPerouse2013TS}, but this mixing line is a strong feature of the water masses discussed in this paper.
+
+\begin{figure}[htbp]
+  \begin{center}
+     \includegraphics[width=5in]{LaPerouse2013TS}
+     \caption{Potential temperature versus salinity for the La Perouse data shown in \fref{fig:LaPerouse2013Ctd}.
+Potential density is contoured in the same way.
+A definition of spice anomaly is shown in this plot, and discussed in the text.
+The rectangle is the $\theta-S$ range used in \fref{fig:TSdensSpice} below.}
+     \label{fig:LaPerouse2013TS}
+  \end{center}
+\end{figure}
+
+The spice anomaly defined this way is mapped in the middle row of each set of plots in \fref{fig:LaPerouse2013Ctd}.
+Offshore water tends to be high is absolute spice anomaly, with a positive-anomaly layer (red) centered at $26.4\ \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and below.
+On the shelf ($-10\, \mathrm{km} < X < 30\, \mathrm{km}$), the distinct $\theta$--$S$ masses are attenuated and much closer to the mixing line than the offshore water.
+Hugging the coast ($X > 30\, \mathrm{km}$), is the Vancouver Island Current, with a negative spice anomaly in June, and a positive anomaly in summer, consistent with the warming of this water during the summer.
+
+The June and September cruises are quite similar in water properties, particularly in the offshore waters.
+Water has upwelled from deeper depths in September, but has most of the same water mass characteristics as earlier in the year.
+The water on the shelf is somewhat warmer than earlier in the year, and by September has started to fall off the mixing line.
+
+In terms of oxygen, both cruises have a similar dichotomy between the onshelf and offshelf water, with water found in the eddy tracking at least $100\ \mathrm{\mu mol\, kg^{-1}}$ lower on the shelf than offshelf (\fref{fig:LaPerouse2013Ctd}, \fref{fig:LaPerouse2013TS}b).
+The very deepest water ($S\approx 33.9\ \mathrm{psu}$) on the shelf shows a further $50\ \mathrm{\mu mol\, kg^{-1}}$ drop between June and September along the LB line.
+
+\subsection{Fine-scale surveys}
+
+\subsubsection{Overview}
+
+As shown in \fref{fig:LocMapBoth},  we conducted a large spatial survey of the shelf using the Moving Vessel Profiler, collecting over 2000 CTD casts in a relatively synoptic manner.
+This amount of data allows a reasonable census of the water masses that motivated our choice of the mixing line shown in \fref{fig:LaPerouse2013TS}.
+Binning in temperature and salinity, we can look at sample frequency of these observations \fref{fig:TSdensSpice}, and we see that there are distinct water masses with few observations between them.
+Looking along the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal, we see that there is a large population along the mixing line.
+At the opposite extreme, we see that there is a large population with a positive spice anomaly of approximately $0.2\ \mathrm{kg\,m^{-3}}$, representing the offshore water.
+Between these two end points there is a third water mass, somewhat indistinct from the offshore water mass, but very distinct from the water mass along the mixing line.
+These three water masses exist on deeper and shallower isopycnals as well, but become hard to distinguish from $\theta$--$S$ properties alone as the properties converge near $26.1\,\mathrm{kg\,m^{-3}}$.
+
+% TODO: shallower?
 
@@
 
-     \includegraphics[width=6.2in]{LocMapBoth.pdf}
-    \caption{a) Study site on the Vancouver Island Shelf.
-The blue arrow indicates the direction of the Vancouver Island Current.
-The pink arrow indicates southward flow of the coastal upwelling current.
-The dashed box indicates the approximate limits of the study area.
-b) The study area with standard hydrographic casts indicated in yellow dots, and Moving Vessel Profiler casts as red dots.
-      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProc.ipynb ;
-        /Users/jklymak/PW13/figs/mvp//LocMapBoth.pdf}
-      \label{fig:LocMapBoth} }
+     \includegraphics[width=3.5in]{TSdensSpice.pdf}
+    \caption{Sample density of salinity and potential density data from the cruise, with a logarithmic color scale.
+Grey contours are potential density relative to the surface, every $0.1\ \mathrm{kg\,m^{-3}}$, and the magenta contour is the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.
+Spice anomaly, $\gamma$ is defined as temperature anomaly scaled by $\alpha$ along isopycnals from the mixing line found in the well-mixed region onshore, in $\mathrm{kg\,m^{-3}}$.
+    \label{fig:TSdensSpice}}
   \end{center}
 \end{figure*}
+
+These three water masses are quite coherent in space.
+While we did not do the specific La Perouse sections, our spatial coverage is such that we can make consistent synthetic sections (\fref{fig:CrossSectionsSpice}), taken at $y=42, 28, 14$, and $0\, \mathrm{km}$ (see \fref{fig:LocMapBoth}).
+During the August cruise, the upstream water along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$ has moderate spice anomaly of  $\gamma \approx 0.1\,\mathrm{kg\,m^{-3}}$ (\fref{fig:CrossSectionsSpice}a), and in this representation, the water appears "pink".
+There is some higher-anomaly water offshore of $x \approx -40\, \mathrm{km}$ along this upstream section.
+
+By $y\approx 14\, \mathrm{km}$, this ``pink'' water along $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ is gone and is replaced by high spice anomaly water $\gamma \approx 0.2 \,\mathrm{kg\,m^{-3}}$ (``red'') that is directly next to well-mixed water ($\gamma \approx 0.0 \,\mathrm{kg\,m^{-3}}$, ``white''), a situation that persists further down the shelf (\fref{fig:CrossSectionsSpice}c, d).
+The onshore waters ($-5 < X < 40\, \mathrm{km}$) are remarkably well-mixed in $\theta$--$S$ space, despite still retaining significant density stratification.
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=6.2in]{CrossSectionsSpice}
+    \caption{Synthetic sections of spice anomaly from the MVP data, projected along four lines from poleward (upstream of coastal current) a) to equatorward (downstream); the lines are shown in \fref{fig:LocMapBoth}b).
+Isopycnals are contoured every $\sigma_{\theta} = 0.2\,\mathrm{kg\,m^{-3}}$ , with the $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ isopycnal highlighted in magenta.
+Colors are spice anomaly as defined in \fref{fig:TSdensSpice}.
+"BC" in panel (c) refers to Barkley Canyon.
+      \label{fig:CrossSectionsSpice} }
+  \end{center}
+\end{figure*}
+
+Oxygen profiles, computed in the same manner, show similar trends, with the low-spice anomaly water clearly having a very low oxygen signature, while the strong negative spice anomaly shallower than $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ has very high oxygen, and the deeper water is also more oxygenated.
+Of note is that compared with the coarser sampling along the LC Line (\fref{fig:LaPerouse2013Ctd}), we see that there is a very abrupt transition between the low spice anomaly/low-O2 water and the offshore water over the ``Eddy'' region.
+We will see below that this transition is even sharper than indicated by these synthetic sections, since they alias some lateral and temporal irregularity in the $\theta$--$S$-compensated front.
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=6.2in]{CrossSectionsO2}
+    \caption{Synthetic sections of Oxygen, approximate percent saturation from the MVP data.
+      \label{fig:CrossSectionsO2} }
+  \end{center}
+\end{figure*}
+
+The spatial patterns are very clear when considering a map-view of properties along the $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ isopycnal (\fref{fig:SpiceO2264}).
+There is a region onshore of the south tip of La Perouse Bank (approximately 44.5 N, 125.75 W), that consists of water that is found along the mixing line (spice anomaly $\gamma \approx 0 \mathrm{kg\,m^{-3}}$) with low oxygen saturations; this water is called the Juan de Fuca eddy (or Tully eddy), though here an eddy shape is not readily apparent.
+Offshore of this region is water that is very warm and salty in comparison ($\gamma > 0.15 \mathrm{kg\,m^{-3}}$), and relatively high in oxygen (saturations of approximately 60\%).
+The region between these two water masses is very abrupt, and is concave with respect to the isobaths on the shelf,  stretching from La Perouse Bank to a shallow bank just above the Juan de Fuca canyon (48.3 N and 125.4 W).
+
+Upstream of La Perouse bank, and Barkley Canyon, the water along the shelf has a weaker spice anomaly than the water further downstream ($\gamma \approx 0.7$, light pink colors).
+This water is also somewhat lower in oxygen than water from offshore, though not as depleted as the water in the "eddy".
+This water appears to be pushed offshore just upstream of Barkley Canyon, and replaced on the shelf by the warmer offshore water.
 
@@
 Along-slope grey lines are every 20 km in the cross-slope direction, with $x=0\ \mathrm{km}$ near the 100-m isobath at the north end of the observation area.
-
-      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProcNew.ipynb ;
-        /Users/jklymak/PW13/figs/mvp//SpiceO2264.pdf}
-      \label{fig:SpiceO2264} }
+   \label{fig:SpiceO2264}
+    }
   \end{center}
@@
 
+\subsubsection{Frontal survey}
+
+A systematic survey through the front between the offshore water and the ``eddy'' water demonstrates the structure of the front (\fref{fig:Frontsurvey}).
+The survey started close to shore and passed through the Vancouver Island Current (casts 2030--2080).
+This water is buoyant and signficiantly fresher and cooler than water at the same density near the surface.
+There is strong evidence of isopycnal mixing of the surface water and the outflow water, with warm-salty water in the relatively wide front between the ``eddy'' water and the water in the current.
+
+The front between the offshore water and the ``eddy'' water was passed through 10 times, showing its evolution following the along-shore equatorward flow.
+First, as noted in the composite sections, isopycnals slope up from offshore onto the shelf.
+This is in agreement with a shelf-break current, though the isopycnals flatten out in the ``eddy'' region.
+The front starts out very sharp, (along-track distance $50\ \mathrm{km}$) though two small tendrils can be seen separating from the front on the inshore side.
+Similar tendrils are found on the second pass, perhaps a bit more separated from the front ($\approx 80\ \mathrm{km}$), and on the third pass ($\approx 95\ \mathrm{km}$).
+Note that these tendrils are made of up partially mixed water.
+The subsequent passes have more of this partially mixed water, such that the partially mixed region is up to 5-km wide by the fifth pass ($\approx 150\ \mathrm{km}$).
+However, note that deeper isopycnals retain a sharp front, and indeed the front appears sharp again by the seventh pass at all depths.
+
+There is evidence of some warmer water swirling into ``eddy'', particularly along isopycnals deeper than $26.4\,\mathrm{kg\,m^{-3}}$.
+Regions of warmer (and more oxygenated) water are found in tendrils at these depths (e.g.\ $\approx\,170\ \mathrm{km}$ and $\approx 185\,\mathrm{km}$).
+
+The overall effect is similar to what was seen in the Gulf Stream with similar observations \cite{klymaketal16}; there are two quite distinct water masses, as seen in the $\theta$--$S$ plot (\fref{fig:Frontsurvey}c).
+There is only a small population of samples between these two, indicative of substantial isopycnal and vertical mixing, but even these populations are relatively cut off from the large water masses, indicating that they are well-mixed on their own, in short episodic events.
+
 \begin{figure*}[htbp]
   \begin{center}
-    \includegraphics[width=6.2in]{CrossSectionsSpice}
-    \caption{
-      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProcNew.ipynb ;
-        /Users/jklymak/PW13/figs/mvp//CrossSectionsSpice.pdf}
-      \label{fig:CrossSectionsSpice} }
+    \includegraphics[width=6.2in]{Frontsurvey}
+    \caption{Data from a survey focusing on the front region (2013-08-28T15:11 to 2013-08-29T09:18) a) spice anomaly along $26.4\,\mathrm{kg\,m^{-3}}$.
+Grey-white alternating lines are the coordinate system, with 10-km alternating shades.
+b) oxygen saturation; magenta dots correspond to distance along-track.
+c) density of data points (logscale) in this section of data.
+d) cross-section of spice anomaly along track.
+Grey isopycnals are contoured every $0.5\,\mathrm{kg\,m^{-3}}$, and the $26.4\,\mathrm{kg\,m^{-3}}$ isopycnal is shown in magenta.
+e) cross-section of oxygen saturation.
+      \label{fig:Frontsurvey} }
   \end{center}
@@
 
-\begin{figure}[htbp]
+\subsection{Tully Canyon}
+
+The seasurface is low in the middle of the ``Eddy'', so water has been hypothesized to move up Tully Canyon due to ageostrophic motion \cite{weaverhsieh87,freelanddenman82}.
+This is difficult to infer from water properties alone.
+Three transects up the canyon indicate that deep isopycnals slope up into the canyon (\fref{fig:CanyonPropertiesSpice}, to approximately 30 km).
+Across the eddy, isopycnals are largely flat until they intersect the Vancouver Island current ($x\approx 80\ \mathrm{km}$).
+The deeper isopycnals do not make it into the deeper basin northeast of La Perouse bank ($x\approx 70\ \mathrm{km}$).
+
+\begin{figure*}[htbp]
   \begin{center}
-    \includegraphics[width=3in]{TsLineDefn}
+    \includegraphics[width=6.2in]{CanyonPropertiesSpice}
+    \caption{Magenta line in map inset is path taken during a cross-canyon survey.
+      \label{fig:CanyonPropertiesSpice} }
+  \end{center}
+\end{figure*}
+
+Water properties along the canyon go from modified offshore water, to eddy water, with a relatively consistent transition at about 30 km.
+Oxygen behaves in a similar manner, though some of the incoming water has slightly lower oxygen than water inside the eddy (not shown).
+Based on these sections, it is difficult to infer water motion up the canyon in particular.
+
+Much of the modified water in the Tully Canyon appears to come from the shelf to the west, but heavily modified by tidal moxing.
+A repeated tidal survey over the bank on the west side of the canyon shows a strong hydraulic response (\fref{fig:HydraulicsCanyon}) during onshore flow (17:58--21:00).
+Note the tide here is largely diurnal, so this onslope flow only occurs once a day.
+Given the stratification of $N\approx 6\times10^{-3} \ \mathrm{rad\,s^{-1}}$ and an overturning scale of 50 m, we might expect dissipation rates reaching $\epsilon \sim L^2N^{3} \approx 5\times10^{-4}\ \mathrm{m^2\,s^{-3}}$, which is more than three orders of magnitude higher than dissipation observed west of this location by \cite{deweycrawford88}.
+The effect on the properties in the canyon is for water to spill over from the offshore front into the canyon during the tide.
+This water is rapidly mixed with surrounding water such that its strong offshore spice values are attenuated.
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=4in]{HydraulicsCanyon}
+    \caption{Repeated, tide-resolving survey across the Tully Canyon, time is indicated in the upper left of each plot (29 August, 2013).
+      \label{fig:HydraulicsCanyon}
+    }
+  \end{center}
+\end{figure*}
+
+This cross-canyon mixing makes it ambiguous if there is water moving up the canyon or not.
+There is indeed a general tendency along the canyon for higher spice water to be found offshore (\fref{fig:CanyonPropertiesSpice}) but it seems likely the source of the higher spice is over the bank rather than water being advected up the canyon.
+
+\subsection{Offshore tongue}
+
+The very narrow front on the offshore side of the ``eddy'' region persists, despite there being considerable ``intermediate'' water found upstream along the shelf.
+This water must go somewhere (or be accelerated unreasonably fast in the frontal region), and indeed it appears to largely be pushed offshore just upstream of Barkley Canyon (\fref{fig:IsopycnalsDepth}).
+There is a tongue of partially mixed ($\gamma$ is pink along $26.4\,\mathrm{kg\,m^{-3}}$) that flows almost due south just west of 126 W.
+We had trouble crossing the whole tongue with the scale of our surveys, but it is at least 30 km wide.
+It also appears to end at approximately 48.2 N.
+This partially mixed water reaches from relatively shallow isopycnals to at least $26.6\,\mathrm{kg\,m^{-3}}$ (\fref{fig:IsopycnalsDepth}, left panel).
+Unfortunately, we cannot track the fate of the isopycnals deeper than this because they tilt down offshore, below the depth limit of the MVP.
+The shallower isopycnal appears to have the tongue closer to the shelf than at $26.4\,\mathrm{kg\,m^{-3}}$, indicating strong three-dimensionality to this feature.
+
+This feature is embedded in the larger scale isopycnal tilt caused by the upwelling (\fref{fig:IsopycnalsDepth} right panels), so it is difficult to see dynamically what is driving this offshore push.
+One possibility is that it is simple flow separation as the water is not able to make the sharp turn around Swiftsure Bank.
+Whatever causes it, the water downstream is replaced by offshore water with much higher spice values.
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=6.2in]{IsopycnalsDepth}
     \caption{
-      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProcNew ;
-        /Users/jklymak/PW13/figs/mvp//TsLineDefn.pdf}
-      \label{fig:TsLineDefn} }
+      \label{fig:IsopycnalsDepth} }
+  \end{center}
+\end{figure*}
+
+We can attempt to track this feature in satellite imagery, though there are lots of clouds during this time period.
+Sea-surface temperatures are generally colder coming out of the Strait of Juan de Fuca (\fref{fig:SST}).
+Water flowing equatorward along the shelf tends to be cooler that offshore water, though this water is significantly saltier than the Juan de Fuca water hugging the shelf.
+On 2013-08-26, there is a cold tongue of water separating from La Perouse bank, and pointing south at 125.8 W.
+Further downstream, there is the hint of a slightly cooler tendril streaming west at 48 N.
+This feature is not as well-developed in the previous image, 2013-08-20, perhaps indicating that it is an evolving mesoscale feature (or a consequence of the cloud cover making for smeared 1-d composites)
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=4in]{SST}
+    \caption{Sea surface temperature composites \cite{GHRSST, ChinEtAl17} from the observation period.
+Note these composites are during a time of year when there is substantial cloud cover.\label{fig:SST}}
   \end{center}
@@
 
-%\begin{appendix}
+A similar impression is given by ocean chlorophyll estimates (\fref{fig:ChlA}).
+The 8-d composites show a region of high chlorophyll  on the west side of the eddy.
+It also shows a relatively  high-chlorophyll squirt to the west, again exiting the study region at approximately 48 degrees N.
+The feature is relatively long-lived, on the order of one month.
+Inspection of images before 2013-08-05 were too obscured by clouds, or did not show this feature, and even by 2013-09-06 we see the feature fading from the satellite.
+However, note that this feature is somewhat south of tongue that we observe deeper in the water column, but is indicative of substantial offshore transport in the region.
+On the scale of mesoscale features along the coast it is not unusually large, but is a facet of the local oceanography not often commented on.
 
-%\subsection{First appendix secondary heading}
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=6.2in]{ChlA}
+    \caption{Surface chlorophyll density estimated from ocean color \cite{Huetal12,MODISChlL3} over 8-day windows in 4-km bins. Gray regions had too many clouds for acceptable averages.
+      \label{fig:ChlA} }
+  \end{center}
+\end{figure*}
+
+\section{Summary and Discussion}
+\label{sec:Summary}
+
+The results of this fine-scaled survey perhaps pose more questions than they answer, but the questions are important to set the processes that should be looked for in understanding the circulation of the southern Vancouver Island shelf.
+
+\subsection{Age and source of the eddy water}
+
+The water we have been referring to as the Juan de Fuca Eddy is a relatively well-mixed water mass, as evidenced by the water therein lying along a mixing-line in $\theta$--$S$ space relative to water offshore (\fref{fig:LaPerouse2013TS} and \fref{fig:TSdensSpice}).
+Water coming equatorward along the shelf is closer to offshore water, though also shows clear signs of modification due to mixing.
+
+Unfortunately, this data set cannot definitively track where the water in the eddy originated from.
+The deepest water in the eddy could originate along the $\theta$--$S$ line from approximately 5.5 degrees to 7.5 degrees (\fref{fig:LaPerouse2013TS}a), which in the open ocean spans depths from 420 m to 70 m.
+\cite{mackasetal87} attempted to determine the origin of the water by including oxygen as a third variable to resolve the ambiguous $\theta$--$S$ relation.
+However, as \fref{fig:LaPerouse2013TS}b makes clear, oxygen is definitely not a conserved property in the eddy, with concentrations up to $150 \ \mathrm{\mu mol\ kg^{1}}$ lower in the eddy than the water found offshore, and in a way that definitely cannot result from conservative mixing.
+As a best guess, if the water in the deepest part of the eddy came from a vertical mixture of the water at $26.5\,\mathrm{kg\,m^{-3}}$ and an equal distance down in density space of $26.7\,\mathrm{kg\,m^{-3}}$, then the deepest water in the eddy would be coming from a depth of 250 m.
+This is actually a relatively typical upwelling depth for coastal flows, and may not require extra input up Tully canyon as posited by \cite{freelanddenman82}.
+
+We found little evidence of vigorous flow up the Tully Canyon, or if there is, then the mixing in the canyon is quick enough to remove the $\theta$--$S$ signature of offshore water within 20 km (\fref{fig:CanyonPropertiesSpice}).
+We did find substantial evidence of mixing in the canyon, but the primary route of high-spice water into the canyon appears to be due to tidal flow over the banks on the west side (\fref{fig:HydraulicsCanyon}), rather than evidence of flow up the canyon.
+
+There is evidence of aging of the water in the eddy between late spring and late summer (\fref{fig:LaPerouse2013Ctd}), with a reduction of oxygen in the eddy.
+If we posited that the reduction was all from the same water, then the oxygen consumption rate over the time between the late-May and early September cruises would be on the order of $0.5\ \mathrm{\mu mol\, kg^{-1} d^{-1}}$.
+This seems on the low side of continental upwellng system estimates of apparent oxygen utilization rates, which are between $1$ and $5\  \mathrm{\mu mol\, kg^{-1} d^{-1}}$ \cite{dortchetal94}.
+\marginpar{More citations?
+Is AOU really poorly estimated on shelves?}  So it seems possible the eddy has enough exchange with the surrounding water to have a shorter residence time.
+
+There is partial mixing of the water in the eddy, though it remains vertically stratified.
+The amount of turbulence required to homogenize 100 m of water over 90 days is $\kappa_{\rho} \approx 10^{-3}\ \mathrm{m^2\,s^{-1}}$.
+Given that the diffusivity implied in the cross-channel surveys was very locally on the order of $\kappa_{\rho} = 10 \ \mathrm{m^2\,s^{-1}}$, this number is not outrageous if we think that such high dissipation is found in $10^{-4}$ of the water column.
+If the diffusivity were higher, it could support a shorter residence time in the eddy, but there would need to be enough turbulence found on the shelf to support this.
+
+The sharpness of the front with the eddy and the offshore water is also intriguing.
+It was persistent for the duration of our detailed survey (\fref{fig:SpiceO2264}), and, so far as we can tell with the limited resolution, was present during the La Perouse cruises.
+There is not any substantial bathymetry blocking the onshore incursion of water at this location, so there must be a dynamic barrier.
+In contrast, numerical simulations of this region reported by \cite{sahuetal22} using a NEMO 36th-degree regional model indicate that there is relatively rapid exchange between the deep eddy water and the rest of the coastal ocean.
+The region where the eddy resides has velocities equal or greater than other parts of the shelf, and water has an approximate residence time of less than 20 days.
+Consequently, the eddy water does not develop a distinct $\theta$--$S$ signature from the surrounding water.
+
+Overall, it would be an improvement to our understanding of the eddy if we could sample the shelf more persistently.
+The eddy was already well-formed by the June La Perouse cruise, and seems to evolve slowly during that time.
+Capturing its formation, presumably earlier in the spring, and its evolution through the year could answer
+
+\subsection{Offshore exchange of shelf water}
+
+The (apparent) displacement of shelf water at this location is relatively dramatic.
+Eddies have been known to separate from irregular coastal topography, both at the surface \cite{barthetal00}, and deeper in the water column \cite{pellandetal13}.
+However, the wholesale replacement of shelf water by a new water mass from offshore, as shown here is relatively rare.
+In our situation it is clear that water from as deep as 150 m is separating from the shelf and moving offshore (\fref{fig:SpiceO2264}, \fref{fig:IsopycnalsDepth}).
+
+Satellite imagery makes it clear that there is often an exchange between coastal and deepwater along the Vancouver Island shelf (\fref{fig:SSTLateAug}).
+Most years there are three of four large excursions from the shelf into the open ocean, many of them over 100 km long.
+It is possible that there is even some spatial consistency to these features, with a persistent separation at the north tip of the Island, and a strong tendency for one at 49.5 N.
+The study area also has evidence of a squirt in most of the years, with 2011 being the only clear exception.
+
+Large-scale mixing between the shelf and open ocean has been evident since the satellite era.
+However, here we demonstrate that the flow is originating on the shelf and separating from the bathymetry and being injected into the interior down to the bottom of the water column.
+A very similar observation was made by \citeA{barthetal00} downstream of Cape Blanco, Oregon, where the coastal current was observed to detach from the shelf in the lee of the cape, and squirt into the interior.
+They hypothesized that as the current moved offshore, it deepened, stretching isopycnals and creating cyclonic relative vorticity that would tend to push the current back onshelf, but then it was caught in the undercurrent and stalled, being pushed offshore.
+It is also possible that coastally trapped waves in the region experience a hydraulic control, and these meanders are the response \cite{dalebarth01}.
+
+General baroclinic instability of the upwelling front is a possible mechanism to drive offshore exchange \cite<e.g.>{durskiallen05}, but this tends to be shallow, with smaller-scale instabilities that will not extend as far into the interior ocean as observed here.
+Rather it seems likely that the topographic change engendered by the sudden turn to the east of La Perouse bank catalyses a larger scale instability at this location.
+In \citeA{durskiallen05}, when including realistic shelf bathymetry including Heceta Bank, the region around the bank catalyzes intermittent large-scale instabilities similar to the feature here and along the coast.
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=6in]{SSTLateAug}
+    \caption{
+      Late August sea-surface temperature, 8-day composites \cite{MODISSST8d}.
+From 2011 to 2021, missing years had too much cloud cover or no satellite coverage (note the 2013 data in \fref{fig:SST} are from daily composites).
+      \label{fig:SSTLateAug} }
+  \end{center}
+\end{figure*}
+
+Regardless of the dynamics, the offshore transport is substantial.
+If we assume the coastal current is approximately $0.1\ \mathrm{m\,s^{-1}}$ over 100 m in the vertical and 20 km in the horizontal, it represents 0.2 Sv of nutrient-rich shelf water transported offshore.
+We do not have velocity measurements for the water that replaces it, but assuming that water also flows along-shelf, there is a large replacement of low O2, low nutrient water with offshore water at this location, with potentially important implications for coastal productivity that are not accounted for in 2-dimensional concepts of Ekman-driven cross-shelf dynamics.
+
+\clearpage
+%\section{Here Is Appendix Title}
+% will show
+% A: Here Is Appendix Title
 %
-%\subsection{Second appendix secondary heading}
+%\appendix
+%\section{Here is a sample appendix}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%\subsubsection{First appendix tertiary heading}
+% Optional Glossary, Notation or Acronym section goes here:
 %
-%\subsubsection{Second appendix tertiary heading}
+%%%%%%%%%%%%%%
+% Glossary is only allowed in Reviews of Geophysics
+%  \begin{glossary}
+%  \term{Term}
+%   Term Definition here
+%  \term{Term}
+%   Term Definition here
+%  \term{Term}
+%   Term Definition here
+%  \end{glossary}
+
 %
-%\paragraph{First appendix quaternary heading}
+%%%%%%%%%%%%%%
+% Acronyms
+%   \begin{acronyms}
+%   \acro{Acronym}
+%   Definition here
+%   \acro{EMOS}
+%   Ensemble model output statistics
+%   \acro{ECMWF}
+%   Centre for Medium-Range Weather Forecasts
+%   \end{acronyms}
+
 %
-%\paragraph{Second appendix quaternary heading}
+%%%%%%%%%%%%%%
+% Notation
+%   \begin{notation}
+%   \notation{$a+b$} Notation Definition here
+%   \notation{$e=mc^2$}
+%   Equation in German-born physicist Albert Einstein's theory of special
+%  relativity that showed that the increased relativistic mass ($m$) of a
+%  body comes from the energy of motion of the body—that is, its kinetic
+%  energy ($E$)—divided by the speed of light squared ($c^2$).
+%   \end{notation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%\end{appendix}
+%  ACKNOWLEDGMENTS
+%
+% The acknowledgments should list:
+%
+% 	All funding sources related to this work from all authors
+%
+% 	Any real or perceived financial conflicts of interests for any
+%	author
+%
+% 	Other affiliations for any author that may be perceived as
+% 	having a conflict of interest with respect to the results of this
+% 	paper.
+%
+%   It is also the appropriate place to thank colleagues and other contributors.
+%   AGU does not normally allow dedications.
 
-% Create a bibliography directory and place your .bib file there.
-% -REMOVE ALL DIRECTORY PATHS TO REFERENCE FILES BEFORE SUBMITTING TO THE AMS FOR PEER REVIEW
-\ifthenelse{\boolean{dc}}
-{}
-{\clearpage}
-\bibliographystyle{ametsocjmk}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section*{Open Research}
+
+%Open Research
+% AGU requires an Availability Statement for the underlying data needed to understand, evaluate, and build upon the reported research at the time of peer review and publication.
+
+%Additionally, authors should include an Availability Statement for the software that has a significant impact on the research.
+Details and templates are in the Availability Statement section of the Data & Software for Authors Guidance:
+% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#availability
+
+%For physical samples, use the IGSN persistent identifier, see the International Geo Sample Numbers section:
+% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#IGSN
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\acknowledgments
+This section is optional. Include any Acknowledgments here.
+
+%% ------------------------------------------------------------------------ %%
+%% References and Citations
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% \bibliography{<name of your .bib file>} don't specify the file extension
+%
+% don't specify bibliographystyle
+
+% In the References section, cite the data/software described in the Availability Statement (this includes primary and processed data used for your research).
+For details on data/software citation as well as examples, see the Data & Software Citation section of the Data & Software for Authors guidance
+% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#citation
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \bibliography{main}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURES-REMOVE ALL DIRECTORY PATHS TO FIGURE FILES BEFORE SUBMITTING TO THE AMS FOR PEER REVIEW
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Reference citation instructions and examples:
+%
+% Please use ONLY \cite and \citeA for reference citations.
+% \cite for parenthetical references
+% ...as shown in recent studies (Simpson et al., 2019)
+% \citeA for in-text citations
+% ...Simpson et al. (2019) have shown...
+%
+%
+%...as shown by \citeA{jskilby}.
+%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
+%...has been shown \cite{jskilbye}.
+%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
+%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
+%...has been shown by \cite <e.g.,>[and others]{lewin76}.
+%
+% apacite uses < > for prenotes and [ ] for postnotes
+% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \citealp, etc.).
+% \nocite is okay to use to add references from your Supporting Information
+%
 
+\end{document}
+
+More Information and Advice:
+
+%% ------------------------------------------------------------------------ %%
 %
-\end{document}
+%  SECTION HEADS
+%
+%% ------------------------------------------------------------------------ %%
+
+% Capitalize the first letter of each word (except for
+% prepositions, conjunctions, and articles that are
+% three or fewer letters).
+
+% AGU follows standard outline style; therefore, there cannot be a section 1 without
+% a section 2, or a section 2.3.1 without a section 2.3.2.
+% Please make sure your section numbers are balanced.
+% ---------------
+% Level 1 head
+%
+% Use the \section{} command to identify level 1 heads;
+% type the appropriate head wording between the curly
+% brackets, as shown below.
+%
+%An example:
+%\section{Level 1 Head: Introduction}
+%
+% ---------------
+% Level 2 head
+%
+% Use the \subsection{} command to identify level 2 heads.
+%An example:
+%\subsection{Level 2 Head}
+%
+% ---------------
+% Level 3 head
+%
+% Use the \subsubsection{} command to identify level 3 heads
+%An example:
+%\subsubsection{Level 3 Head}
+%
+%---------------
+% Level 4 head
+%
+% Use the \subsubsubsection{} command to identify level 3 heads
+% An example:
+%\subsubsubsection{Level 4 Head} An example.
+%
+%% ------------------------------------------------------------------------ %%
+%
+%  IN-TEXT LISTS
+%
+%% ------------------------------------------------------------------------ %%
+%
+% Do not use bulleted lists; enumerated lists are okay.
+% \begin{enumerate}
+% \item
+% \item
+% \item
+% \end{enumerate}
+%
+%% ------------------------------------------------------------------------ %%
+%
+%  EQUATIONS
+%
+%% ------------------------------------------------------------------------ %%
+
+% Single-line equations are centered.
+% Equation arrays will appear left-aligned.
+
+Math coded inside display math mode \[ ...\]
+ will not be numbered, e.g.,:
+ \[ x^2=y^2 + z^2\]
+
+ Math coded inside \begin{equation} and \end{equation} will
+ be automatically numbered, e.g.,:
+ \begin{equation}
+ x^2=y^2 + z^2
+ \end{equation}
+
+% To create multiline equations, use the
+% \begin{eqnarray} and \end{eqnarray} environment
+% as demonstrated below.
+\begin{eqnarray}
+  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
+        && + (y - y_{0}) \sin \Theta  \nonumber \\
+  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
+        && + (y - y_{0}) \cos \Theta.
+\end{eqnarray}
+
+%If you don't want an equation number, use the star form:
+%\begin{eqnarray*}...\end{eqnarray*}
+
+% Break each line at a sign of operation
+% (+, -, etc.) if possible, with the sign of operation
+% on the new line.
+
+% Indent second and subsequent lines to align with
+% the first character following the equal sign on the
+% first line.
+
+% Use an \hspace{} command to insert horizontal space
+% into your equation if necessary. Place an appropriate
+% unit of measure between the curly braces, e.g.
+% \hspace{1in}; you may have to experiment to achieve
+% the correct amount of space.
+
+%% ------------------------------------------------------------------------ %%
+%
+%  EQUATION NUMBERING: COUNTER
+%
+%% ------------------------------------------------------------------------ %%
+
+% You may change equation numbering by resetting
+% the equation counter or by explicitly numbering
+% an equation.
+
+% To explicitly number an equation, type \eqnum{}
+% (with the desired number between the brackets)
+% after the \begin{equation} or \begin{eqnarray}
+% command.  The \eqnum{} command will affect only
+% the equation it appears with; LaTeX will number
+% any equations appearing later in the manuscript
+% according to the equation counter.
+%
+
+% If you have a multiline equation that needs only
+% one equation number, use a \nonumber command in
+% front of the double backslashes (\\) as shown in
+% the multiline equation above.
+
+% If you are using line numbers, remember to surround
+% equations with \begin{linenomath*}...\end{linenomath*}
+
+%  To add line numbers to lines in equations:
+%  \begin{linenomath*}
+%  \begin{equation}
+%  \end{equation}
+%  \end{linenomath*}

### 2022-02-05 gather_figs.py
Update on Overleaf.

@@ -1,0 +1,11 @@
+import os
+import re
+import shutil
+
+with open('PW13Paper.log') as fin:
+    for l in fin:
+        searchst = r'\<use (.+?)\>'
+        name = re.search(searchst,l)
+        if name:
+            print(name.group(1))
+            shutil.copy(name.group(1), './figs')

### 2022-02-05 gather_figs.py
Fix bib

@@ -8,4 +8,12 @@ with open('PW13Paper.log') as fin:
         name = re.search(searchst,l)
         if name:
             print(name.group(1))
-            shutil.copy(name.group(1), './figs')
+            try:
+                shutil.copy(name.group(1), './figs')
+            except:
+                pass
+
+try:
+    shutil.copy('/Users/jklymak/texmf/bibtex/bib/main.bib', './main.bib')
+except:
+    pass

### 2022-04-09 PW13Paper.tex
New version

@@
 The southern Vancouver Island continental shelf is very productive due to high nutrient input from the Strait of Juan de Fuca and Salish Sea estuarine system, and substantial cross-shelf transport due to the complicated topography.
-The estuarine water flows poleward, hugging the coast as the Vancouver Island current, and in the summer shelf water flows equatorward.
+The estuarine water flows poleward, hugging the coast as the Vancouver Island Coastal Current, and in the summer shelf water flows equatorward.
 Trapped between these two currents is the Juan de Fuca eddy, a stagnant region of water that occupies the shelf where it abruptly widens just north of the Strait of Juan de Fuca.
@@
 The shelf has complicated bathymetry \fref{fig:LocMapBoth}, with a somewhat typical shelf poleward of our study site, but then widening in the lee of La Perouse Bank, and then hitting the Juan de Fuca canyon, equatorward.
-Water from the strait of Juan de Fuca flows poleward as a buoyant current that hugs the coast \cite{thomsonetal89, hickeyetal91}, while shelf water flows from poleward in the summer under upwelling conditions, both forced by local winds and via tele-connections with the long, homogenous shelves equatorward off Washington and Oregon \cite{hickeyetal91,thomsonkrassovski15,engidaetal16}.
-Trapped between these two currents is a region of relatively homogenous water that has been termed the Juan de Fuca eddy \cite{freelanddenman82,freelandmcintosh89,foremanetal07,macfadyenhickey10}.
+Water from the strait of Juan de Fuca flows poleward as a buoyant current that hugs the coast \cite{thomsonetal89, hickeyetal91}, while shelf water flows equatorward in the summer under upwelling conditions, both forced by local winds and via tele-connections with the long, homogenous shelves equatorward off Washington and Oregon \cite{hickeyetal91,thomsonkrassovski15,engidaetal16}.
+Trapped between these two currents is a region of relatively homogenous water that has been termed the ``Juan de Fuca'' eddy \cite{freelanddenman82,freelandmcintosh89,foremanetal07,macfadyenhickey10}.
 
-One of our findings below is that the water in the Juan de Fuca eddy is apparently quite old.
+One of our findings below is that the water in the deep pool of the Juan de Fuca eddy is either  quite old or undergone significant vertical mixing.
 The low oxygen found in the Juan de Fuca eddy has caused it to be inferred to be upwelled from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
 This inference was based on the oxygen being conservative, and hence having to come from close to the oxygen minimum zone found further offshore \cite{mackasetal87}.
-This lead to hypotheses that perhaps the eddy was fed by waters being drawn up a spur canyon of the Juan de Fuca canyon (often called the Tully Canyon) via ageostrophic transport from offshore to onshore due to the low pressure in the eddy center \cite{weaverhsieh87}.
+This lead to hypotheses that perhaps the eddy was fed by waters being drawn up a spur canyon of the Juan de Fuca canyon (called the Spur Canyon) via ageostrophic transport from offshore to onshore due to the low pressure in the eddy center \cite{weaverhsieh87}.
 Below, we will argue that there is not substantial evidence of such transport, and that oxygen is likely low because of consumption due to respiration.
 
-A second important finding is that the water flowing equatorward along the shelf from upstream is all displaced from the shelf and pushed offshore and replaced by offshore water that floods the southern shelf.
+A second important finding is that the water flowing equatorward along the shelf from upstream all detaches from the shelf and is pushed offshore to be replaced by offshore water that floods the southern shelf.
 Such events are important as they tend to transport nutrient rich shelf water out into the open ocean, fueling offshore ecosystems.
-This cross-shelf exchange has not, to our knowledge, been remarked upon for the Vancouver Island shelf.
-Mesoscale separations have been observed further to the south off Cape Blanco \cite{barthetal00}, and was also inferred at other locations along the coast.
-Similar observations have been made elsewhere \cite[e.g.]{relvasbarton05}, though usually in the presence of a distinct coastal headland, a feature not found at our study location.
-The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics \cite{dalebarth01} leading to along-shore trapping of coastal waves,  wind stress curl variations \cite{castelaobarth07}, and induced relative vorticity due to stretching of parcels that inertially overshoot their initial isobaths due to a sudden change in downstream bathymetry.
+Similar offshore motion has been remarked in the region using satellite images \cite{ikedaemery84,thomsongower98}, and observed further to the south off Cape Blanco \cite{barthetal00}, and other coasts \cite[e.g.]{relvasbarton05}.
+Net cross-shore transport of nutrients and chlorophyll have been found in hydrographic surveys off Vancouver Island \cite{mackasyelland99}, and associated with mesoscale features both in geostrophic velocities, and in satellites.
+The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics \cite{dalebarth01} leading to along-shore trapping of coastal waves,  wind stress curl variations \cite{castelaobarth07}, induced relative vorticity due to stretching of parcels that inertially overshoot their initial isobaths due to a sudden change in downstream bathymetry, and the non-linear breaking of large-scale meanders due to baroclinic instability between the wind-driven current and the California Undercurrent \cite{ikedaetal84, batteen97}.
 
-In this paper we present our observations from 2013 by describing our at-sea observation tools (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes, with a focus on the offshore side of the eddy region (\fref{sec:Observations}) where we consider the age of the eddy, the steadiness of the front separating the eddy and the offshore water, properties along the spur canyon that was believed to feed the eddy, and finally describe the large offshore separation of the shelf-break current.
-We discuss the origin of the eddy and the implications of the separating jet.
+In this paper we present  observations from 2013 (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes, with a focus on the offshore side of the eddy region (\fref{sec:Observations}) where we consider the age of the eddy, the steadiness of the front separating the eddy and the offshore water, properties along the spur canyon that was believed to feed the eddy, and finally describe the large offshore separation of the shelf-break current and its intrusion into the interior.
+We discuss the origin of the eddy and the implications of the separating jet (\fref{sec:Summary}).
 
@@
   \caption{a) Study site on the Vancouver Island Shelf.
-The blue arrow indicates the direction of the Vancouver Island Current.
+The blue arrow indicates the direction of the Vancouver Island Coastal Current.
 The pink arrow indicates southward flow of the coastal upwelling current.
@@
 At the south end of the study site is the Juan de Fuca canyon, which feeds dense water into Juan de Fuca Strait.
-This dense water is mixed with fresh water from the Fraser River at the sills and archipelagos further inland, and fluxes out the Strait again, where it turns poleward along Vancouver Island to form the Vancouver Island Current.
-The Juan de Fuca Canyon has a notable spur canyon, the Tully Canyon, that incises the shelf towards the north into the study site (\fref{fig:LocMapBoth}b).
+This dense water is mixed with fresh water from the Fraser River at the sills and archipelagos further inland, and fluxes out the Strait again, where it turns poleward along Vancouver Island to form the Vancouver Island Coastal Current.
+The Juan de Fuca Canyon has a notable spur canyon (Spur Canyon), that incises the shelf towards the north into the study site (\fref{fig:LocMapBoth}b).
 The rest of the shelf is punctuated by a series of banks and shallow basins.
@@
 
-Observations aboard the \emph{R/V Falkor} were carried out from 2013-08-21 to 2013-08-30, starting two weeks before the September Tully cruise.
+Observations aboard the \emph{R/V Falkor} were carried out from 2013-08-21 to 2013-08-30, starting two weeks before the September \emph{Tully} cruise.
 Data were collected with an ODIM Brooke-Ocean Moving Vessel Profiler, profiling to 200 m.
@@
 
-Winds during the cruises were typical for the west coast of Vancouver Island, with down-welling favorable winds during July and early August, and then intermittent winds before and during the Falkor cruise.
-Note that this locale is strongly affected by coastally trapped waves from further south, so upwelling usually persists despite local wind forcing \cite{thomsonkrassovski15, engidaetal16}.
+Winds during the cruises were typical for the west coast of Vancouver Island, with upwelling favorable winds during July and early August (\fref{fig:LaPeWind}).
+During the cruise, and for the week previous, the winds were intermittently downwelling favorable.
+Note that this locale is strongly affected by coastally trapped waves from further south, so doming of near-bottom isopycnals often persists despite local wind forcing \cite{thomsonkrassovski15, engidaetal16}, and takes a finite amount of time to spin down.
 
@@
 This dense water tends to be low in oxygen and cool.
-Further onshore, isopycnals tilt back down, and warm water fills the water column, evidence of the Vancouver Island Current that hugs the coast here.
+Further onshore, isopycnals tilt back down, and warm water fills the water column, evidence of the Vancouver Island Coastal Current that hugs the coast here.
 
@@
 On the shelf ($-10\, \mathrm{km} < X < 30\, \mathrm{km}$), the distinct $\theta$--$S$ masses are attenuated and much closer to the mixing line than the offshore water.
-Hugging the coast ($X > 30\, \mathrm{km}$), is the Vancouver Island Current, with a negative spice anomaly in June, and a positive anomaly in summer, consistent with the warming of this water during the summer.
+Hugging the coast ($X > 30\, \mathrm{km}$), is the Vancouver Island Coastal Current, with a negative spice anomaly in June, and a positive anomaly in summer, consistent with the warming of this water during the summer.
 
@@
 A systematic survey through the front between the offshore water and the ``eddy'' water demonstrates the structure of the front (\fref{fig:Frontsurvey}).
-The survey started close to shore and passed through the Vancouver Island Current (casts 2030--2080).
+The survey started close to shore and passed through the Vancouver Island Coastal Current (casts 2030--2080).
 This water is buoyant and signficiantly fresher and cooler than water at the same density near the surface.
@@
 
-\subsection{Tully Canyon}
+\subsection{Spur Canyon}
 
-The seasurface is low in the middle of the ``Eddy'', so water has been hypothesized to move up Tully Canyon due to ageostrophic motion \cite{weaverhsieh87,freelanddenman82}.
+The seasurface is low in the middle of the ``Eddy'', so water has been hypothesized to move up the Spur Canyon due to ageostrophic motion \cite{weaverhsieh87,freelanddenman82}.
 This is difficult to infer from water properties alone.
 Three transects up the canyon indicate that deep isopycnals slope up into the canyon (\fref{fig:CanyonPropertiesSpice}, to approximately 30 km).
-Across the eddy, isopycnals are largely flat until they intersect the Vancouver Island current ($x\approx 80\ \mathrm{km}$).
+Across the eddy, isopycnals are largely flat until they intersect the Vancouver Island Coastal Current ($x\approx 80\ \mathrm{km}$).
 The deeper isopycnals do not make it into the deeper basin northeast of La Perouse bank ($x\approx 70\ \mathrm{km}$).
@@
 
-Much of the modified water in the Tully Canyon appears to come from the shelf to the west, but heavily modified by tidal moxing.
+Much of the modified water in the Spur Canyon appears to come from the shelf to the west, but heavily modified by tidal moxing.
 A repeated tidal survey over the bank on the west side of the canyon shows a strong hydraulic response (\fref{fig:HydraulicsCanyon}) during onshore flow (17:58--21:00).
@@
     \includegraphics[width=4in]{HydraulicsCanyon}
-    \caption{Repeated, tide-resolving survey across the Tully Canyon, time is indicated in the upper left of each plot (29 August, 2013).
+    \caption{Repeated, tide-resolving survey across the Spur Canyon, time is indicated in the upper left of each plot (29 August, 2013).
       \label{fig:HydraulicsCanyon}
@@
 Sea-surface temperatures are generally colder coming out of the Strait of Juan de Fuca (\fref{fig:SST}).
-Water flowing equatorward along the shelf tends to be cooler that offshore water, though this water is significantly saltier than the Juan de Fuca water hugging the shelf.
-On 2013-08-26, there is a cold tongue of water separating from La Perouse bank, and pointing south at 125.8 W.
-Further downstream, there is the hint of a slightly cooler tendril streaming west at 48 N.
-This feature is not as well-developed in the previous image, 2013-08-20, perhaps indicating that it is an evolving mesoscale feature (or a consequence of the cloud cover making for smeared 1-d composites)
+Water flowing equatorward along the shelf tends to be cooler than offshore water, though this water is significantly saltier than the Juan de Fuca water hugging the shelf.
+On Aug-25, there is a cold tongue of water separating from La Perouse bank, and pointing south at 125.8 W with a cooler tendril streaming west at 48 N.
+This feature is not as well-developed in the previous image, Aug 21, perhaps indicating that it is an evolving mesoscale feature.
+By 31 Aug, there is no surface expression of the feature, though tendrils of cooler water can be seen separating from La Perouse Bank.
+By Sep-05 the water gas significantly warmed, but the offshore pushing anomally does not appear to have a surface signature.
 
@@
   \begin{center}
-    \includegraphics[width=4in]{SST}
-    \caption{Sea surface temperature composites \cite{GHRSST, ChinEtAl17} from the observation period.
-Note these composites are during a time of year when there is substantial cloud cover.\label{fig:SST}}
+    \includegraphics[width=6in]{SSTSnaps}
+    \caption{Sea surface temperature images from the observation period.
+Grey areas are clouds; dashed gray line is the study area.
+Depths are contoured in thin gray lines at 200, 150 and 100 m.
+\cite{L2PMetopA}\label{fig:SST}}
   \end{center}
@@
 
-A similar impression is given by ocean chlorophyll estimates (\fref{fig:ChlA}).
-The 8-d composites show a region of high chlorophyll  on the west side of the eddy.
+Ccean chlorophyll estimates show the same feature (\fref{fig:ChlA}) demostrating the advection of a region of high chlorophyll to the west side of the eddy.
 It also shows a relatively high-chlorophyll squirt to the west, again exiting the study region at approximately 48 degrees N.
@@
 However, note that this feature is somewhat south of tongue that we observe deeper in the water column, but is indicative of substantial offshore transport in the region.
-On the scale of mesoscale features along the coast it is not unusually large, but is a facet of the local oceanography not often commented on.
 
@@
 The deepest water in the eddy could originate along the $\theta$--$S$ line from approximately 5.5 degrees to 7.5 degrees (\fref{fig:LaPerouse2013TS}a), which in the open ocean spans depths from 420 m to 70 m.
-\cite{mackasetal87} attempted to determine the origin of the water by including oxygen as a third variable to resolve the ambiguous $\theta$--$S$ relation.
+\citeA{mackasetal87} attempted to determine the origin of the water by including oxygen as a third variable to resolve the ambiguous $\theta$--$S$ relation.
 However, as \fref{fig:LaPerouse2013TS}b makes clear, oxygen is definitely not a conserved property in the eddy, with concentrations up to $150 \ \mathrm{\mu mol\ kg^{1}}$ lower in the eddy than the water found offshore, and in a way that definitely cannot result from conservative mixing.
 As a best guess, if the water in the deepest part of the eddy came from a vertical mixture of the water at $26.5\,\mathrm{kg\,m^{-3}}$ and an equal distance down in density space of $26.7\,\mathrm{kg\,m^{-3}}$, then the deepest water in the eddy would be coming from a depth of 250 m.
-This is actually a relatively typical upwelling depth for coastal flows, and may not require extra input up Tully canyon as posited by \cite{freelanddenman82}.
+This is actually a relatively typical upwelling depth for coastal flows, and may not require extra input up Spur canyon as posited by \citeA{freelanddenman82}.
 
-We found little evidence of vigorous flow up the Tully Canyon, or if there is, then the mixing in the canyon is quick enough to remove the $\theta$--$S$ signature of offshore water within 20 km (\fref{fig:CanyonPropertiesSpice}).
+We found little evidence of vigorous flow up the Spur Canyon, or if there is, then the mixing in the canyon is quick enough to remove the $\theta$--$S$ signature of offshore water within 20 km (\fref{fig:CanyonPropertiesSpice}).
 We did find substantial evidence of mixing in the canyon, but the primary route of high-spice water into the canyon appears to be due to tidal flow over the banks on the west side (\fref{fig:HydraulicsCanyon}), rather than evidence of flow up the canyon.
+However, it is also of note that upwelling winds had ceased at this point (\fref{fig:LaPeWind}), meaning that the offshore pressure gradient may be reduced, leading to less ageostrophic upwelling in the canyon.
+Whether the cessation of winds would also lead to a reduction of the low sea level height in the center of the eddy is an open question.
 
@@
 This seems on the low side of continental upwellng system estimates of apparent oxygen utilization rates, which are between $1$ and $5\  \mathrm{\mu mol\, kg^{-1} d^{-1}}$ \cite{dortchetal94}.
-\marginpar{More citations?
-Is AOU really poorly estimated on shelves?}  So it seems possible the eddy has enough exchange with the surrounding water to have a shorter residence time.
+So it seems possible the eddy has enough exchange with the surrounding water to have a shorter residence time.
 
-There is partial mixing of the water in the eddy, though it remains vertically stratified.
-The amount of turbulence required to homogenize 100 m of water over 90 days is $\kappa_{\rho} \approx 10^{-3}\ \mathrm{m^2\,s^{-1}}$.
-Given that the diffusivity implied in the cross-channel surveys was very locally on the order of $\kappa_{\rho} = 10 \ \mathrm{m^2\,s^{-1}}$, this number is not outrageous if we think that such high dissipation is found in $10^{-4}$ of the water column.
-If the diffusivity were higher, it could support a shorter residence time in the eddy, but there would need to be enough turbulence found on the shelf to support this.
+There is partial mixing of the water in the eddy, though it remains vertically stratified, and the amount of homogenization is such that either the mixing is very strong, or the water is retained in the eddy for a long time.
+The amount of turbulence required to homogenize 100 m of water over 90 days is $\kappa \approx 10^{-3}\ \mathrm{m^2\,s^{-1}}$.
+Given that the diffusivity implied in the cross-channel surveys was (very) locally on the order of $\kappa_{\rho} = 10 \ \mathrm{m^2\,s^{-1}}$, this number is not outrageous if we think that such high dissipation is found in $10^{-4}$ of the water column.
+We can more carefully quantify this by considering a synthetic profile or temperature and salinity based on an offshore profile, extrapolated from the bottom of the 200-m cast to 250 m, assuming that the temperature and salinity at 250 m are $6.2\ \mathrm{[^oC]}$ and $34\ \mathrm{psu}$ respectively (\fref{fig:TSExercise}).
+Water at the offshore station was warmer than onshore, so the profile was also linearly interpolated to a surface value of ($31.6\ \mathrm{psu}$, $12\ \mathrm{^oC}$).
+The profile was then linearly compressed into a depth of 150 m representative of the shelf depth in the dense pool, and subjected to mixing with a constant diffusivity of $\kappa = 10^{-3}\ \mathrm{m^2\,s^{-1}}$, with the surface and bottom values pinned under the assumption that the near-bottom source and surface waters are replenished from a large reservoir.
+Similar to the naive scaling, the T-S relationship does not approach a straight line until after approximately 60-100 days, or until the mixing affects a vertical length scale of $\lambda = \left(\tau \kappa\right)^{1/2}$ that is greater than $\approx 70\,\mathrm{m}$.
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=4in]{TSExercise}
+    \caption{Mixing model assuming constant eddy diffusivity of $\kappa = 10^{-3}\ \mathrm{m^2\,s^{-1}}$ acting on an offshore temperature and salinity profiles compressed from 250 thick to 150 m thick.
+Profiles are pinned to the deep value at ($34\ \mathrm{psu}$, $6.2\ \mathrm{^oC}$), and a shallow one at ($31.6\ \mathrm{psu}$, $12\ \mathrm{^oC}$).
+      \label{fig:TSExercise} }
+  \end{center}
+\end{figure*}
 
@@
 In contrast, numerical simulations of this region reported by \cite{sahuetal22} using a NEMO 36th-degree regional model indicate that there is relatively rapid exchange between the deep eddy water and the rest of the coastal ocean.
-The region where the eddy resides has velocities equal or greater than other parts of the shelf, and water has an approximate residence time of less than 20 days.
+In this model, the region where the eddy resides has velocities equal or greater than other parts of the shelf, and water has an approximate residence time of less than 20 days.
 Consequently, the eddy water does not develop a distinct $\theta$--$S$ signature from the surrounding water.
@@
 The eddy was already well-formed by the June La Perouse cruise, and seems to evolve slowly during that time.
-Capturing its formation, presumably earlier in the spring, and its evolution through the year could answer
+Capturing its formation, presumably earlier in the spring, and its evolution through the year would be valuable in undertanding retention and exchange on this productive part of the shelf.
 
@@
 
-The (apparent) displacement of shelf water at this location is relatively dramatic.
+The (apparent) displacement of shelf water from La Perousse Bank is relatively dramatic.
 Eddies have been known to separate from irregular coastal topography, both at the surface \cite{barthetal00}, and deeper in the water column \cite{pellandetal13}.
-However, the wholesale replacement of shelf water by a new water mass from offshore, as shown here is relatively rare.
+It has been recognized that instabilities lead to exchange between the open ocean and the shelf at this location \cite{ikedaemery84, ikedaemery84}.
+However, observations of the wholesale replacement of shelf water by a new water mass from offshore, is relatively rare.
 In our situation it is clear that water from as deep as 150 m is separating from the shelf and moving offshore (\fref{fig:SpiceO2264}, \fref{fig:IsopycnalsDepth}).
@@
 Most years there are three of four large excursions from the shelf into the open ocean, many of them over 100 km long.
-It is possible that there is even some spatial consistency to these features, with a persistent separation at the north tip of the Island, and a strong tendency for one at 49.5 N.
-The study area also has evidence of a squirt in most of the years, with 2011 being the only clear exception.
+The lengthscale in these images is somewhat larger than that inferred for this region by \citeA{ikedaetal84}.
+It is possible that there is even some spatial locking of these features, with a persistent separation at the north tip of the Island, and a strong tendency for one at 49.5 N.
+Our study area also has evidence of a squirt in most of the years, with 2011 being the only clear exception.
+General baroclinic instability of the upwelling front is a possible mechanism to drive offshore exchange \cite<e.g.>{ikedaetal84,durskiallen05}, but this tends to be shallow, with smaller-scale instabilities that will not extend as far into the interior ocean as observed here.
+Rather it seems likely that the topographic change engendered by the sudden turn to the east of La Perouse bank catalyses a larger scale instability at this location.
+In \citeA{durskiallen05}, when including realistic shelf bathymetry including Heceta Bank, the region around the bank catalyzes intermittent large-scale instabilities similar to the feature here and along the coast \cite{batteen97}.
 
 Large-scale mixing between the shelf and open ocean has been evident since the satellite era.
-However, here we demonstrate that the flow is originating on the shelf and separating from the bathymetry and being injected into the interior down to the bottom of the water column.
-A very similar observation was made by \citeA{barthetal00} downstream of Cape Blanco, Oregon, where the coastal current was observed to detach from the shelf in the lee of the cape, and squirt into the interior.
+Here we demonstrate that the flow is originating on the shelf and separating from the bathymetry and being injected into the interior down to the bottom of the water column.
+A  similar observation was made by \citeA{barthetal00} downstream of Cape Blanco, Oregon, where the coastal current was observed to detach from the shelf in the lee of the cape, and squirt into the interior.
 They hypothesized that as the current moved offshore, it deepened, stretching isopycnals and creating cyclonic relative vorticity that would tend to push the current back onshelf, but then it was caught in the undercurrent and stalled, being pushed offshore.
 It is also possible that coastally trapped waves in the region experience a hydraulic control, and these meanders are the response \cite{dalebarth01}.
-
-General baroclinic instability of the upwelling front is a possible mechanism to drive offshore exchange \cite<e.g.>{durskiallen05}, but this tends to be shallow, with smaller-scale instabilities that will not extend as far into the interior ocean as observed here.
-Rather it seems likely that the topographic change engendered by the sudden turn to the east of La Perouse bank catalyses a larger scale instability at this location.
-In \citeA{durskiallen05}, when including realistic shelf bathymetry including Heceta Bank, the region around the bank catalyzes intermittent large-scale instabilities similar to the feature here and along the coast.
 
@@
     \caption{
-      Late August sea-surface temperature, 8-day composites \cite{MODISSST8d}.
-From 2011 to 2021, missing years had too much cloud cover or no satellite coverage (note the 2013 data in \fref{fig:SST} are from daily composites).
+      Late August sea-surface temperature, 8-day composites \cite{MODISSST8d}.   From 2011 to 2021; missing years had too much cloud cover or no satellite coverage.
       \label{fig:SSTLateAug} }
@@
 Regardless of the dynamics, the offshore transport is substantial.
-If we assume the coastal current is approximately $0.1\ \mathrm{m\,s^{-1}}$ over 100 m in the vertical and 20 km in the horizontal, it represents 0.2 Sv of nutrient-rich shelf water transported offshore.
-We do not have velocity measurements for the water that replaces it, but assuming that water also flows along-shelf, there is a large replacement of low O2, low nutrient water with offshore water at this location, with potentially important implications for coastal productivity that are not accounted for in 2-dimensional concepts of Ekman-driven cross-shelf dynamics.
+If we assume the coastal current is approximately $0.1\ \mathrm{m\,s^{-1}}$ over 100 m in the vertical and 20 km in the horizontal, it represents 0.2 Sv of nutrient- and chlorophyll-rich shelf water transported offshore.
+Our observations are a finer detailed representation of the kind of cross-shore transports inferred by \citeA{mackasyelland99} from hydrographic surveys, and definitively show that this water can originate from the shelf from relatively deep depths and be transported offshore.
+We do not have velocity measurements for the water that replaces it, but assuming that water also flows along-shelf, there is a large replacement of low-O2, low-nutrient water with offshore water at this location.
+This emphasizes the importance of three dimensional observations and modeling of cross-shelf dynamics when thinking about biological processes on the shelf.
 

### 2022-04-25 PW13Paper.tex
FIX
Abstract

@@
 \begin{abstract}
+
+Comprehensive water column observations of temperature, salinity, and oxygen on the productive southern Vancouver Island shelf offer a unique characterization of a number of processes affecting exchange on the shelf.
+A semi-permanent cyclonic ``eddy'' occupies the region in the lee of a bank where the shelf widens abruptly.
+The observations indicate that the water in this wide part of the shelf has distinct water mass properties, and is clearly a mixture of offshore water and water in a coastal current.
+The water in the ``eddy'' is either rapidly mixed (high $\kappa$), or has a long residence time (large $\tau$), such that $(\kappa \tau)^{1/2} \sim 50 -- 100\ \mathrm{m}$.
+The offshore edge of this ``eddy'' water was very abrupt, with distinct water properties showing little lateral mixing between offshore and ``eddy'' water, with the front no more than 1-km wide.
+Previous work had hypothesized the role of a spur canyon incising the shelf in feeding water to the ``eddy''; little evidence of this was found during the observations, however, these were collected during a relaxation and reversal of upwelling winds, which may have slowed the flow of water into the eddy.
+Tidal observations in the spur canyon also show very strong hydraulic flows in the cross-canyon direction accompanied by very high localized mixing rates in 50-m hydraulic jumps possibly exceeding $\kappa \sim 1 \ \mathrm{m^2s^{-1}}$.
+Upstream of the observation there is evidence of an along-shelf current flowing equatorward.
+This water is readily identified by partially mixed water properties, accumulated while on the shelf.
+This water is clearly observed to all separate from the shelf to a depth of approximately 150 m, and is replaced offshore of the ``eddy'' by distinct offshore water.
+This event can be seen in sea-surface temperatures in satellite images as tongue of water that is ejected offshore, probably with the relaxation of upwelling, with the flow heading directly offshore.
+The cause of this offshore separation event are not determined, but likely include instability between the shelfbreak current and the California Undercurrent, and possible are catalyzed by separation from the abrupt underwater bank at the polward end of the study region.
 
@@
 Given the stratification of $N\approx 6\times10^{-3} \ \mathrm{rad\,s^{-1}}$ and an overturning scale of 50 m, we might expect dissipation rates reaching $\epsilon \sim L^2N^{3} \approx 5\times10^{-4}\ \mathrm{m^2\,s^{-3}}$, which is more than three orders of magnitude higher than dissipation observed west of this location by \cite{deweycrawford88}.
-The effect on the properties in the canyon is for water to spill over from the offshore front into the canyon during the tide.
+In terms of a diapycnal diffusivity $\kappa = \gamma \epsilon / N^2 \approx 1\ \mathrm{m^2\,s^{-1}}$, assuming a mixing efficiency of $\gamma=0.2$ The effect on the properties in the canyon is for water to spill over from the offshore front into the canyon during the tide.
 This water is rapidly mixed with surrounding water such that its strong offshore spice values are attenuated.

### 2022-04-25 PW13Paper.tex
Author

@@
 
-\authors{Jody M. Klymak\affil{1} and Everyone }
+\authors{Jody M. Klymak\affil{1}\affil{2} and Susan E. Allen\affil{3}\affil{4} and Others}
 
-% \affiliation{1}{First Affiliation}
-% \affiliation{2}{Second Affiliation}
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
-\affiliation{1}{School of Earth and Ocean Sciences and Department of Physcics \& Astronomy, University of Victoria, B.C.\ Canada}
+\affiliation{1}{School of Earth and Ocean Sciences, University of Victoria, BC, Canada}
+\affiliation{2}{Department of Physcics \& Astronomy, University of Victoria, BC, Canada}
+\affiliation{3}{Department of Earth, Ocean and Atmospheric Sciences, University of British Columbia, Vancouver, BC, Canada}
+\affiliation{4}{Institute of Applied Mathematics, University of British Columbia, Vancouver, BC, Canada}
 %(repeat as many times as is necessary)

### 2022-06-07 PW13Paper.tex
Author
FIX
DOC: version after SNW suggestions

--- 2022-04-09 PW13Paper.tex
@@
 
-\title{Separation of an upwelling current bounding the Juan de Fuca Eddy, off Vancouver Island}
+\title{Separation of an upwelling current bounding the Juan de Fuca Eddy}
 
@@
 
-\authors{Jody M. Klymak\affil{1} and Everyone }
+\authors{Jody M. Klymak\affil{1,2}, Susan E. Allen\affil{3,4}, Stephanie N. Waterman\affil{3}}
 
-% \affiliation{1}{First Affiliation}
-% \affiliation{2}{Second Affiliation}
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
-\affiliation{1}{School of Earth and Ocean Sciences and Department of Physcics \& Astronomy, University of Victoria, B.C.\ Canada}
+\affiliation{1}{School of Earth and Ocean Sciences, University of Victoria, BC, Canada}
+\affiliation{2}{Department of Physcics \& Astronomy, University of Victoria, BC, Canada}
+\affiliation{3}{Department of Earth, Ocean and Atmospheric Sciences, University of British Columbia, Vancouver, BC, Canada}
+\affiliation{4}{Institute of Applied Mathematics, University of British Columbia, Vancouver, BC, Canada}
 %(repeat as many times as is necessary)
@@
 \begin{keypoints}
-\item The shelfbreak current along Vancouver Island separates downstream of a submarine bank.
-\item Offshore water is drawn onto the shelf and forms a sharp semi-persisnte front with the Juan de Fuca eddy.
-\item The Juan de Fuca eddy shows evidence of long residence times, and little evidence of deep-water origin.
+\item The shelf break current along Vancouver Island separates downstream of a submarine bank.
+\item Offshore water is drawn onto the shelf and forms a sharp semi-persistent front with the Juan de Fuca Eddy.
+\item The Eddy shows evidence of long residence times, and little evidence of deep-water origin.
 \end{keypoints}
 
-%% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT and PLAIN LANGUAGE SUMMARY
-%
-% A good Abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties.
+\begin{abstract}
 
-% The Plain Language Summary should be written for a broad audience,
-% including journalists and the science-interested public, that will not have
-% a background in your field.
-%
-% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
-% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
-% see http://sharingscience.agu.org/creating-plain-language-summary/)
-%
-%% ------------------------------------------------------------------------ %%
+Comprehensive water column observations of temperature, salinity, and oxygen on the productive southern Vancouver Island shelf offer a unique characterization of processes affecting exchange on the shelf.
+A semi-permanent cyclonic recirculation (the Juan de Fuca Eddy) occupies the region in the lee of a bank where the shelf widens abruptly.
+The observations indicate that the water in this Eddy is a mixture of offshore water and water from a buoyant coastal current.
+The water in the  in the Eddy is well-mixed along a mixing line in $\theta$--$S$ space, though it retains stratification, and is either rapidly mixed (high $\kappa$), or has a long residence time (large $\tau$), such that $(\kappa \tau)^{1/2} \sim 50 -- 100\ \mathrm{m}$.
+There is a sharp temperature-salinity front on the offshore side of of this well-mixed water, with the front no more than 1-km wide, and no sign of instabilities.
+Previous work had hypothesized the role of a spur canyon incising the shelf in feeding water to the Eddy; little evidence of this was found during the observations, however, these were collected during a relaxation and reversal of upwelling winds, which may have slowed the flow of water into the Eddy.
+Tidally resolved observations in the spur canyon also show very strong hydraulic flows in the cross-canyon direction accompanied by very high localized mixing rates in 50-m hydraulic jumps possibly exceeding $\kappa \sim 1 \ \mathrm{m^2s^{-1}}$.
 
-%% \begin{abstract} starts the second page
-
-\begin{abstract}
+Upstream of the Eddy, there is an along-shelf current flowing equatorward with partially mixed water properties compared to the well-mixed water in the Eddy.
+The shelf current was observed to separate from the shelf in the lee of a bank, and is replaced by the distinct offshore water mass that forms the offshore front with the Eddy.
+The separation event can be seen in sea-surface temperatures in satellite images as tongue of water that is ejected offshore, possibly with the relaxation of upwelling, with the flow heading directly offshore.
+The cause of this offshore separation event is not determined, but likely involves instability between the shelfbreak current and the California Undercurrent, and possibly is catalyzed by separation from the abrupt underwater bank at the poleward end of the study region.
 
@@
 
-Cross-shelf exchange is important to the health and productivity of continental shelf regions, where oxygenated water is exchanged with nutrient-rich, but oxygen-depleted nearshore water.
-Cross-shelf transport usually requires ageostrophic flow, since balanced flow will tend to follow topographic contours, often providing a barrier to lateral exchange \cite{brink16}.
+Cross-shelf exchange is important to the health and productivity of continental shelf regions, where offshore oxygenated water is exchanged with nutrient-rich, but oxygen-depleted nearshore water.
+Cross-shelf transport usually requires ageostrophic flow, since geostrophically balanced flow will tend to follow topographic contours, often providing a barrier to lateral exchange \cite{brink16}.
 Mechanisms for cross-shelf exchange include internal waves and instabilities in shelf-break fronts.
@@
 
-Here we present detailed observations from the southern Vancouver Island shelf, collected in summer 2013, and sampled with a rapid profiling vehicle equipped with a CTD and oxygen sensor, supplemented by traditional hydrographic surveys.
+Here we present detailed \emph{in-situ} observations from the southern Vancouver Island shelf, collected in summer 2013, and sampled with a rapid profiling vehicle equipped with a CTD and oxygen sensor, supplemented by traditional hydrographic surveys.
 The shelf has complicated bathymetry \fref{fig:LocMapBoth}, with a somewhat typical shelf poleward of our study site, but then widening in the lee of La Perouse Bank, and then hitting the Juan de Fuca canyon, equatorward.
 Water from the strait of Juan de Fuca flows poleward as a buoyant current that hugs the coast \cite{thomsonetal89, hickeyetal91}, while shelf water flows equatorward in the summer under upwelling conditions, both forced by local winds and via tele-connections with the long, homogenous shelves equatorward off Washington and Oregon \cite{hickeyetal91,thomsonkrassovski15,engidaetal16}.
-Trapped between these two currents is a region of relatively homogenous water that has been termed the ``Juan de Fuca'' eddy \cite{freelanddenman82,freelandmcintosh89,foremanetal07,macfadyenhickey10}.
+Trapped between these two currents is a region of relatively homogenous water that has been termed the Juan de Fuca Eddy \cite{freelanddenman82,freelandmcintosh89,foremanetal07,macfadyenhickey10}.
 
-One of our findings below is that the water in the deep pool of the Juan de Fuca eddy is either  quite old or undergone significant vertical mixing.
-The low oxygen found in the Juan de Fuca eddy has caused it to be inferred to be upwelled from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
-This inference was based on the oxygen being conservative, and hence having to come from close to the oxygen minimum zone found further offshore \cite{mackasetal87}.
-This lead to hypotheses that perhaps the eddy was fed by waters being drawn up a spur canyon of the Juan de Fuca canyon (called the Spur Canyon) via ageostrophic transport from offshore to onshore due to the low pressure in the eddy center \cite{weaverhsieh87}.
+A goal of our study was to understand how the Juan de Fuca Eddy persists, and how it exchanges water properties with offshore water.
+Low oxygen has previously been found in the Eddy, causing it to be inferred that water in the eddy is upwelled from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
+This inference was based on the oxygen being conservative, and hence having to come from close to the oxygen minimum zone found further offshore \cite{mackasetal87} and lead to hypotheses that perhaps the eddy was fed by waters being drawn up a spur canyon (called the Spur Canyon) via ageostrophic transport from offshore to onshore due to the low pressure in the eddy center \cite{weaverhsieh87}.
 Below, we will argue that there is not substantial evidence of such transport, and that oxygen is likely low because of consumption due to respiration.
 
-A second important finding is that the water flowing equatorward along the shelf from upstream all detaches from the shelf and is pushed offshore to be replaced by offshore water that floods the southern shelf.
+A second goal was to better understand cross-shore exchange with offshore water.
+While the typical schematic of shelf exchange is that there are Ekman layers near the boundaries, there is also evidence of filaments and wholesale separation of shelf currents into the deep ocean, both at this site, and at others.
 Such events are important as they tend to transport nutrient rich shelf water out into the open ocean, fueling offshore ecosystems.
-Similar offshore motion has been remarked in the region using satellite images \cite{ikedaemery84,thomsongower98}, and observed further to the south off Cape Blanco \cite{barthetal00}, and other coasts \cite[e.g.]{relvasbarton05}.
-Net cross-shore transport of nutrients and chlorophyll have been found in hydrographic surveys off Vancouver Island \cite{mackasyelland99}, and associated with mesoscale features both in geostrophic velocities, and in satellites.
+Along the Vancouver Island shelf, large filaments or instabilities of the coastal current have been inferred using satellite images \cite{ikedaemery84,thomsongower98}.
+Direct observations of such filaments have been made further to the south off Cape Blanco \cite{barthetal00}, and other coasts \cite[e.g.]{relvasbarton05}.
+Net cross-shore transport of nutrients and chlorophyll have been found in hydrographic surveys off Vancouver Island \cite{mackasyelland99}, and associated with mesoscale features both in geostrophic velocities, and in satellite observations.
 The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics \cite{dalebarth01} leading to along-shore trapping of coastal waves,  wind stress curl variations \cite{castelaobarth07}, induced relative vorticity due to stretching of parcels that inertially overshoot their initial isobaths due to a sudden change in downstream bathymetry, and the non-linear breaking of large-scale meanders due to baroclinic instability between the wind-driven current and the California Undercurrent \cite{ikedaetal84, batteen97}.
 
-In this paper we present  observations from 2013 (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes, with a focus on the offshore side of the eddy region (\fref{sec:Observations}) where we consider the age of the eddy, the steadiness of the front separating the eddy and the offshore water, properties along the spur canyon that was believed to feed the eddy, and finally describe the large offshore separation of the shelf-break current and its intrusion into the interior.
+In this paper we present  observations from 2013 (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes, with a focus on the offshore side of the eddy region (\fref{sec:Observations}) where we consider the age of the eddy, the steadiness of the front separating the eddy and the offshore water, properties along the spur canyon that was believed to feed the eddy, and finally describe a large offshore tongue of the shelf-break current and its intrusion into the interior.
 We discuss the origin of the eddy and the implications of the separating jet (\fref{sec:Summary}).
@@
 
-Observations from the LB and LC lines were collected aboard the \emph{CCGS Tully} from 2013-05-30 to 2013-05-31, which we will call ``June'', and 2013-09-07 to 2013-09-09, which we will call ``September''.
-These lines span the depths from about 50 m to deep offshore.
+We consider routine hydrographic surveys along the LB and LC lines, collected aboard the \emph{CCGS Tully} from 2013-05-30 to 2013-05-31  (``June''), and 2013-09-07 to 2013-09-09 (``September'').
+These lines span the depths from about 50 m to deep offshore, with casts every 7.5 km across shelf.
 The data comes from a lowered Seabird 9-11 CTD, with an SBE 43 oxygen sensor.
@@
 
-Observations aboard the \emph{R/V Falkor} were carried out from 2013-08-21 to 2013-08-30, starting two weeks before the September \emph{Tully} cruise.
-Data were collected with an ODIM Brooke-Ocean Moving Vessel Profiler, profiling to 200 m.
+We focus on fine-scale surveys carried out from 2013-08-21 to 2013-08-30.
+Data were collected from the \emph{R/V Falkor} with an ODIM Brooke-Ocean Moving Vessel Profiler (MVP).
 The MVP was equipped with an AML Oceanographic CTD, and a Rinko Oxygen sensor with a 7-s response time foil.
-The MVP profiles at vertical fall speeds of $3\ \mathrm{m\,s^{-1}}$ to depths of 200 m or to within 5 m of the seafloor, whichever is shallower.
+The MVP profiles to depths of 200 m or to within 5 m of the seafloor, whichever is shallower, and drops at a speed of $3\ \mathrm{m\,s^{-1}}$.
 Data collection took place while the ship cruised at speeds between 5 and 8 kts, usually at around 6 kts to enable fine horizontal spacing of the casts, with typical spacing less than 800 m.
-Data is collected during down- and up-casts, but is only reported for down-casts because the path through the water is highly asymmetric on the up-cast, whereas the downcast is largely vertical.
+Data is reported for the downcast, which mostly follows a vertical path.
 The rapid speed of the profiling makes the oxygen measurements somewhat coarse, and probably biased due to the phase lag of the sensor, so we treat these largely qualitatively in this paper.
@@
 Winds during the cruises were typical for the west coast of Vancouver Island, with upwelling favorable winds during July and early August (\fref{fig:LaPeWind}).
-During the cruise, and for the week previous, the winds were intermittently downwelling favorable.
+During the finescale cruise, and for the week previous, the winds were intermittently downwelling favorable.
 Note that this locale is strongly affected by coastally trapped waves from further south, so doming of near-bottom isopycnals often persists despite local wind forcing \cite{thomsonkrassovski15, engidaetal16}, and takes a finite amount of time to spin down.
@@
       Wind from the La Perouse buoy \cite{DFOWind2013C46206}.
-The vertical direction is along-shelf (chosen as a heading of 315 N).
+The vertical direction is along-shelf (chosen as a heading of 315 N), so vectors pointing straight down represent upwelling winds.
+.
 The wind components has been low-pass filtered to one-day averages.
@@
 
-Hydrographic sections along the LB and LC lines highlight summer conditions on the Southern Vancouver Island Shelf, and indicate some of the features we are focusing on in this paper (\fref{fig:LaPerouse2013Ctd}).
-During both cruises and at both lines there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m to shallower than 85 m over the shelf.
+Hydrographic sections along the LB and LC lines highlight summer conditions on the Southern Vancouver Island Shelf, and indicate some of the features we are focusing on in this paper (\Fref{fig:LaPerouse2013Ctd}).
+During both cruises and at both lines there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m depth to shallower than 85 m over the shelf.
 This dense water tends to be low in oxygen and cool.
-Further onshore, isopycnals tilt back down, and warm water fills the water column, evidence of the Vancouver Island Coastal Current that hugs the coast here.
+Further onshore, isopycnals tilt back down, and the water warms, evidence of the Vancouver Island Coastal Current that hugs the coast here.
 
@@
      \includegraphics[width=130mm]{LaPerouse2013Ctd}
-     \caption{Observations along LC and LB lines in Jun and Sep 2013.
+     \caption{Observations along LC (top three rows) and LB lines (bottom three rows) in Jun (left column) and Sep 2013 (right column).
 X is across-shelf distance in the coordinate system shown in \fref{fig:LocMapBoth}.
-Potential density is contoured every $0.2\ \mathrm{kg\,m^{-3}}$, with the $26.4\ \mathrm{kg\,m^{-3}}$.
-Potential temperature $\theta$, spice anomaly $\gamma$ and oxygen concentration $O_2$ are colored for each section.}
+Potential density is contoured every $0.2\ \mathrm{kg\,m^{-3}}$, with the $26.4\ \mathrm{kg\,m^{-3}}$ colored magenta.
+Potential temperature, $\theta$, spice anomaly, $\gamma$ and oxygen concentration, $O_2$, are colored for each section.}
      \label{fig:LaPerouse2013Ctd}
@@
 
-The ``Juan de Fuca Eddy'' region is along the LB Line, mostly inshore of 0 km.
+The Juan de Fuca Eddy region is along the LB Line (\fref{fig:LaPerouse2013Ctd}, bottom three rows), mostly inshore of 0 km.
 This water is cooler and lower in oxygen than along the same isopycnal further offshore.
@@
 
-In contrast, the water upstream, along the LC line, does not show as much drawdown of oxygen, and is somewhat warmer, at least offshore of La Perouse bank (\fref{fig:LaPerouse2013Ctd} top three rows).
+The water upstream of the Eddy region, along the LC line appears less mixed than the Eddy water.
+It does not show as much drawdown of oxygen, and is warmer, at least offshore of La Perouse bank (\fref{fig:LaPerouse2013Ctd}, top three rows).
 Water does not appear to make it over La Perouse bank into the basin onshore, or if it does, it does so intermittently.
-There appears to be some mixing of the deeper water and shallow water near $26.4\ \mathrm{kg\,m^{-3}}$, with the water along this isopycnal becoming cooler and lower in oxygen where it intersects the shelf.
+There appears to be some mixing of the deeper and shallow water near $26.4\ \mathrm{kg\,m^{-3}}$, with the water along this isopycnal becoming cooler and lower in oxygen where it intersects the shelf, consistent with enhanced mixing near the shelf
 
@@
 Below $26.6\ \mathrm{kg\,m^{-3}}$ or so, the deep water masses are found offshore, and are largely homogenous.
-Shallower, there are distinct differences between the water found on the shelf and water found more offshore.
+Shallower, there are distinct differences between the water found offshore, with it being warmer and saltier than water on the shelf at the same densities.
 These anomalies are hard to parse directly, so in this paper we use a spice anomaly, relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}).
-The line passes the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$, and spice anomaly for a given water sample at density $\sigma_{\theta}$ is $\gamma = \alpha \left(T - T_0\left(\sigma_{\theta}\right)\right) + \beta \left(S - S_0\left(\sigma_{\theta}\right)\right)$ where $T_0\left(\sigma_{\theta}\right)$ and $S_0\left(\sigma_{\theta}\right)$ are the temperature and salinity along the mixing line at the same density as the water sample, with the sign convention that positive is warmer and saltier than data along the mixing line.
-It is hard to see from the $\theta$--$S$ plot in \fref{fig:LaPerouse2013TS}, but this mixing line is a strong feature of the water masses discussed in this paper.
+This line represents an approximate mixing line for water in the Eddy, and passes the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$.
+Spice anomaly for a given water sample at density $\sigma_{\theta}$ is given by $\gamma = \alpha \left(T - T_0\left(\sigma_{\theta}\right)\right) + \beta \left(S - S_0\left(\sigma_{\theta}\right)\right)$ where $T_0\left(\sigma_{\theta}\right)$, where $S_0\left(\sigma_{\theta}\right)$ are the temperature and salinity along the mixing line at the same density as the water sample, with the sign convention that positive is warmer and saltier than data along the mixing line.
 
@@
      \includegraphics[width=5in]{LaPerouse2013TS}
-     \caption{Potential temperature versus salinity for the La Perouse data shown in \fref{fig:LaPerouse2013Ctd}.
-Potential density is contoured in the same way.
+     \caption{Potential temperature versus salinity for the La Perouse data shown in \fref{fig:LaPerouse2013Ctd}; potential density is contoured every $0.2\ \mathrm{kg\,m^{-3}}$, with the $26.4\ \mathrm{kg\,m^{-3}}$ colored magenta.
 A definition of spice anomaly is shown in this plot, and discussed in the text.
@@
 
-The spice anomaly defined this way is mapped in the middle row of each set of plots in \fref{fig:LaPerouse2013Ctd}.
-Offshore water tends to be high is absolute spice anomaly, with a positive-anomaly layer (red) centered at $26.4\ \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and below.
-On the shelf ($-10\, \mathrm{km} < X < 30\, \mathrm{km}$), the distinct $\theta$--$S$ masses are attenuated and much closer to the mixing line than the offshore water.
-Hugging the coast ($X > 30\, \mathrm{km}$), is the Vancouver Island Coastal Current, with a negative spice anomaly in June, and a positive anomaly in summer, consistent with the warming of this water during the summer.
+Offshore water tends to be high is absolute spice anomaly (\fref{fig:LaPerouse2013Ctd}, middle rows), with a positive-anomaly layer (red) centered at $26.4\ \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and below.
+On the shelf ($-10\, \mathrm{km} < X < 30\, \mathrm{km}$), the distinct $\theta$--$S$ masses are attenuated and much closer to the defined mixing line than the offshore water (closer to white in color).
+Hugging the coast ($X > 30\, \mathrm{km}$), is the Vancouver Island Coastal Current, with a negative spice anomaly in June, and a positive anomaly in summer, consistent with seasonal warming.
 
@@
 Grey contours are potential density relative to the surface, every $0.1\ \mathrm{kg\,m^{-3}}$, and the magenta contour is the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.
-Spice anomaly, $\gamma$ is defined as temperature anomaly scaled by $\alpha$ along isopycnals from the mixing line found in the well-mixed region onshore, in $\mathrm{kg\,m^{-3}}$.
+Spice anomaly, $\gamma$, is defined as temperature anomaly scaled by the thermal expansion co-efficient, $\alpha$, along isopycnals from the mixing line found in the well-mixed region in the Eddy (white line), in $\mathrm{kg\,m^{-3}}$.
     \label{fig:TSdensSpice}}
@@
 
-These three water masses are quite coherent in space.
-While we did not do the specific La Perouse sections, our spatial coverage is such that we can make consistent synthetic sections (\fref{fig:CrossSectionsSpice}), taken at $y=42, 28, 14$, and $0\, \mathrm{km}$ (see \fref{fig:LocMapBoth}).
-During the August cruise, the upstream water along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$ has moderate spice anomaly of  $\gamma \approx 0.1\,\mathrm{kg\,m^{-3}}$ (\fref{fig:CrossSectionsSpice}a), and in this representation, the water appears "pink".
-There is some higher-anomaly water offshore of $x \approx -40\, \mathrm{km}$ along this upstream section.
+In synthetic cross sections of spice anomaly, the Eddy is clearly identifiable as water that is along the straight mixing line, and thus has low-spice anomaly ($\gamma \approx 0\ \mathrm{kg\,m^{-3}}$, \fref{fig:CrossSectionsSpice}c, d).
+This signature of well-mixed water extends from the seafloor to approximately 30 m depth, onshore of $x \approx 0\ \mathrm{km}$.
+Despite lying along a mixing line, the Eddy water is still stratified.
+This well-mixed water contrasts with water directly offshore on the other side of a sharp $\theta-S$ compensated front that is warmer and saltier than the well mixed water ($\gamma > 0.2\ \mathrm{kg\,m^{-3}}$) deeper than $26\,\mathrm{kg\,m^{-3}}$, and colder and fresher shallower.
+The thickness of this front is even sharper in individual sections than in these composite sections (see \fref{sec:frontsurvey}).
 
-By $y\approx 14\, \mathrm{km}$, this ``pink'' water along $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ is gone and is replaced by high spice anomaly water $\gamma \approx 0.2 \,\mathrm{kg\,m^{-3}}$ (``red'') that is directly next to well-mixed water ($\gamma \approx 0.0 \,\mathrm{kg\,m^{-3}}$, ``white''), a situation that persists further down the shelf (\fref{fig:CrossSectionsSpice}c, d).
-The onshore waters ($-5 < X < 40\, \mathrm{km}$) are remarkably well-mixed in $\theta$--$S$ space, despite still retaining significant density stratification.
+In contrast, the third population of partially mixed water mass identified in \fref{fig:TSdensSpice} is all found upstream of the Eddy region (\fref{fig:CrossSectionsSpice}a, b) as a weaker spice anomaly ($\gamma \approx 0.1\ \mathrm{kg\,m^{-3}}$, ``pink'' colors).
+This water is generally found near the intersection of the upwelling water and La Perouse Bank, and is not as warm as offshore water, indicating some mixing has taken place.
+Of note, is that this intermediate water mass appears completely absent in the sections further equatorward (\fref{fig:CrossSectionsSpice}c, d), indicating that it is either mixed away, or that it is advected elsewhere; we argue below that it is advected offshore in a large-scale tongue.
 
@@
 
-Oxygen profiles, computed in the same manner, show similar trends, with the low-spice anomaly water clearly having a very low oxygen signature, while the strong negative spice anomaly shallower than $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ has very high oxygen, and the deeper water is also more oxygenated.
-Of note is that compared with the coarser sampling along the LC Line (\fref{fig:LaPerouse2013Ctd}), we see that there is a very abrupt transition between the low spice anomaly/low-O2 water and the offshore water over the ``Eddy'' region.
-We will see below that this transition is even sharper than indicated by these synthetic sections, since they alias some lateral and temporal irregularity in the $\theta$--$S$-compensated front.
+Oxygen section show similar trends, with the low-spice anomaly water having low oxygen saturations, while the strong negative spice anomaly shallower than $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ has very high oxygen, and the deeper water is also more oxygenated.
+Of note is that compared with the coarser sampling along the LC Line (\fref{fig:LaPerouse2013Ctd}), we see that the abrupt transition between offshore and Eddy water is also present in oxygen.
 
@@
 The spatial patterns are very clear when considering a map-view of properties along the $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ isopycnal (\fref{fig:SpiceO2264}).
-There is a region onshore of the south tip of La Perouse Bank (approximately 44.5 N, 125.75 W), that consists of water that is found along the mixing line (spice anomaly $\gamma \approx 0 \mathrm{kg\,m^{-3}}$) with low oxygen saturations; this water is called the Juan de Fuca eddy (or Tully eddy), though here an eddy shape is not readily apparent.
+There is a region onshore of the south tip of La Perouse Bank (approximately 44.5 N, 125.75 W), that consists of water that is found along the mixing line (spice anomaly $\gamma \approx 0 \mathrm{kg\,m^{-3}}$) with low oxygen saturations.
 Offshore of this region is water that is very warm and salty in comparison ($\gamma > 0.15 \mathrm{kg\,m^{-3}}$), and relatively high in oxygen (saturations of approximately 60\%).
@@
 
-Upstream of La Perouse bank, and Barkley Canyon, the water along the shelf has a weaker spice anomaly than the water further downstream ($\gamma \approx 0.7$, light pink colors).
-This water is also somewhat lower in oxygen than water from offshore, though not as depleted as the water in the "eddy".
+Upstream of La Perouse bank and Barkley Canyon, the water along the shelf the weaker spice anomaly of the third water mass ($\gamma \approx 0.1\,\mathrm{kg\,m^{-3}}$, light pink colors).
+This water is also somewhat lower in oxygen than water from offshore, though not as depleted as the water in the Eddy.
 This water appears to be pushed offshore just upstream of Barkley Canyon, and replaced on the shelf by the warmer offshore water.
@@
 \subsubsection{Frontal survey}
+\label{sec:frontsurvey}
 
-A systematic survey through the front between the offshore water and the ``eddy'' water demonstrates the structure of the front (\fref{fig:Frontsurvey}).
-The survey started close to shore and passed through the Vancouver Island Coastal Current (casts 2030--2080).
-This water is buoyant and signficiantly fresher and cooler than water at the same density near the surface.
-There is strong evidence of isopycnal mixing of the surface water and the outflow water, with warm-salty water in the relatively wide front between the ``eddy'' water and the water in the current.
+A systematic survey through the front between the offshore water and the Eddy water demonstrates the structure of the front (\fref{fig:Frontsurvey}).
+The survey started close to shore and passed through the Vancouver Island Coastal Current (along-track 0-10 km).
+The coastal current water is buoyant and significantly fresher and cooler than water at the same density near the surface and forms a classic density front with
 
@@
 First, as noted in the composite sections, isopycnals slope up from offshore onto the shelf.
-This is in agreement with a shelf-break current, though the isopycnals flatten out in the ``eddy'' region.
+This is in agreement with a shelf-break current, though the isopycnals flatten out in the Eddy region.
 The front starts out very sharp, (along-track distance $50\ \mathrm{km}$) though two small tendrils can be seen separating from the front on the inshore side.
@@
 The subsequent passes have more of this partially mixed water, such that the partially mixed region is up to 5-km wide by the fifth pass ($\approx 150\ \mathrm{km}$).
-However, note that deeper isopycnals retain a sharp front, and indeed the front appears sharp again by the seventh pass at all depths.
+However, note that the deeper isopycnals retain a sharp front, and indeed the front appears sharp again by the seventh pass at all depths.
 
-There is evidence of some warmer water swirling into ``eddy'', particularly along isopycnals deeper than $26.4\,\mathrm{kg\,m^{-3}}$.
+There is evidence of some warmer water swirling into Eddy, particularly along isopycnals deeper than $26.4\,\mathrm{kg\,m^{-3}}$.
 Regions of warmer (and more oxygenated) water are found in tendrils at these depths (e.g.\ $\approx\,170\ \mathrm{km}$ and $\approx 185\,\mathrm{km}$).
 
-The overall effect is similar to what was seen in the Gulf Stream with similar observations \cite{klymaketal16}; there are two quite distinct water masses, as seen in the $\theta$--$S$ plot (\fref{fig:Frontsurvey}c).
-There is only a small population of samples between these two, indicative of substantial isopycnal and vertical mixing, but even these populations are relatively cut off from the large water masses, indicating that they are well-mixed on their own, in short episodic events.
+The overall effect is similar to what was seen in the Gulf Stream with similar observations \cite{klymaketal16}; there are two quite distinct water masses, the Eddy water and the offshore water, as seen in the $\theta$--$S$ plot (\fref{fig:Frontsurvey}c).
+There is only a small population of samples between these two, indicative of substantial isopycnal and vertical mixing, but even these populations are relatively cut off from the main water masses, indicating that they are well-mixed on their own, in short episodic events.
 
@@
 
-The seasurface is low in the middle of the ``Eddy'', so water has been hypothesized to move up the Spur Canyon due to ageostrophic motion \cite{weaverhsieh87,freelanddenman82}.
+The Spur Canyon leading from the Strait of Juan de Fuca has been implicated in allowing dense water to be upwelled into the Eddy.
+The seasurface is low in the middle of the Eddy, so water has been hypothesized to move up the Spur Canyon due to ageostrophic motion \cite{weaverhsieh87,freelanddenman82}.
 This is difficult to infer from water properties alone.
@@
     \includegraphics[width=6.2in]{CanyonPropertiesSpice}
-    \caption{Magenta line in map inset is path taken during a cross-canyon survey.
+    \caption{Three surveys up the Spur Canyon, where $X$ is along-canyon as defined in the map.
+Grey isopycnals are contoured every $0.5\,\mathrm{kg\,m^{-3}}$, and the $26.4\,\mathrm{kg\,m^{-3}}$ isopycnal is shown in magenta.
+The seafloor is indicated with the thick black line.
+      Map shows the path taken during each survey in chronological order blue, orange, and green.  Magenta line is path taken during a cross-canyon survey (\fref{fig:HydraulicsCanyon}).
       \label{fig:CanyonPropertiesSpice} }
@@
 
-Water properties along the canyon go from modified offshore water, to eddy water, with a relatively consistent transition at about 30 km.
+Water properties along the canyon go from offshore water, to Eddy water.
 Oxygen behaves in a similar manner, though some of the incoming water has slightly lower oxygen than water inside the eddy (not shown).
@@
 
-Much of the modified water in the Spur Canyon appears to come from the shelf to the west, but heavily modified by tidal moxing.
+Much of the modified water in the Spur Canyon appears to come from the shelf to the west, but heavily modified by tidal mixing.
 A repeated tidal survey over the bank on the west side of the canyon shows a strong hydraulic response (\fref{fig:HydraulicsCanyon}) during onshore flow (17:58--21:00).
-Note the tide here is largely diurnal, so this onslope flow only occurs once a day.
-Given the stratification of $N\approx 6\times10^{-3} \ \mathrm{rad\,s^{-1}}$ and an overturning scale of 50 m, we might expect dissipation rates reaching $\epsilon \sim L^2N^{3} \approx 5\times10^{-4}\ \mathrm{m^2\,s^{-3}}$, which is more than three orders of magnitude higher than dissipation observed west of this location by \cite{deweycrawford88}.
-The effect on the properties in the canyon is for water to spill over from the offshore front into the canyon during the tide.
+Dense ocean water passes from the offshore side into the Canyon, and plunges down the side wall before rebounding downstream.
+Note that the tide here is largely diurnal, so this onslope flow only occurs once a day.
+Given the stratification of $N\approx 6\times10^{-3} \ \mathrm{rad\,s^{-1}}$ and an overturning scale of 50 m, we might expect dissipation rates reaching $\epsilon \sim L^2N^{3} \approx 5\times10^{-4}\ \mathrm{m^2\,s^{-3}}$, which is more than three orders of magnitude higher than dissipation observed west of this location by \citeA{deweycrawford88}.
+In terms of a diapycnal diffusivity $\kappa = \gamma \epsilon / N^2 \approx 1\ \mathrm{m^2\,s^{-1}}$, assuming a mixing efficiency of $\gamma=0.2$ The effect on the properties in the canyon is for water to spill over from the offshore front into the canyon during the tide.
 This water is rapidly mixed with surrounding water such that its strong offshore spice values are attenuated.
@@
     \caption{Repeated, tide-resolving survey across the Spur Canyon, time is indicated in the upper left of each plot (29 August, 2013).
+See magenta line on map \fref{fig:CanyonPropertiesSpice} for location of survey.
       \label{fig:HydraulicsCanyon}
@@
 
-The very narrow front on the offshore side of the ``eddy'' region persists, despite there being considerable ``intermediate'' water found upstream along the shelf.
-This water must go somewhere (or be accelerated unreasonably fast in the frontal region), and indeed it appears to largely be pushed offshore just upstream of Barkley Canyon (\fref{fig:IsopycnalsDepth}).
+As discussed above, the very sharp T/S front between the well-mixed Eddy water and the offshore water  persists, despite there being considerable partially-mixed water found upstream along the shelf.
+The partially-mixed water must go somewhere, and indeed it appears to largely be pushed offshore just upstream of Barkley Canyon (\fref{fig:IsopycnalsDepth}).
 There is a tongue of partially mixed ($\gamma$ is pink along $26.4\,\mathrm{kg\,m^{-3}}$) that flows almost due south just west of 126 W.
@@
 This partially mixed water reaches from relatively shallow isopycnals to at least $26.6\,\mathrm{kg\,m^{-3}}$ (\fref{fig:IsopycnalsDepth}, left panel).
-Unfortunately, we cannot track the fate of the isopycnals deeper than this because they tilt down offshore, below the depth limit of the MVP.
+Unfortunately, we cannot track the fate of this water mass because isopycnals tilt down offshore, below the depth limit of the MVP.
 The shallower isopycnal appears to have the tongue closer to the shelf than at $26.4\,\mathrm{kg\,m^{-3}}$, indicating strong three-dimensionality to this feature.
@@
 This feature is embedded in the larger scale isopycnal tilt caused by the upwelling (\fref{fig:IsopycnalsDepth} right panels), so it is difficult to see dynamically what is driving this offshore push.
-One possibility is that it is simple flow separation as the water is not able to make the sharp turn around Swiftsure Bank.
-Whatever causes it, the water downstream is replaced by offshore water with much higher spice values.
+One possibility is that it is simple flow separation as the water is not able to make the sharp turn around La Perouse Bank.
+Whatever causes it, downstream of the separation, the water is replaced by offshore water with much higher spice values.
 
@@
     \includegraphics[width=6.2in]{IsopycnalsDepth}
-    \caption{
+    \caption{Isopycnal slices through showing the vertical structure of the water separating from the shelf, with the first row along $25.5 \,\mathrm{kg\,m^{-3}}$, second along $26.4\,\mathrm{kg\,m^{-3}}$, and the third at $26.6\,\mathrm{kg\,m^{-3}}$.
+First column is the spice anomaly, second is oxygen saturation, and the last column is depth of each isopycnal.
       \label{fig:IsopycnalsDepth} }
@@
 
-We can attempt to track this feature in satellite imagery, though there are lots of clouds during this time period.
-Sea-surface temperatures are generally colder coming out of the Strait of Juan de Fuca (\fref{fig:SST}).
-Water flowing equatorward along the shelf tends to be cooler than offshore water, though this water is significantly saltier than the Juan de Fuca water hugging the shelf.
-On Aug-25, there is a cold tongue of water separating from La Perouse bank, and pointing south at 125.8 W with a cooler tendril streaming west at 48 N.
-This feature is not as well-developed in the previous image, Aug 21, perhaps indicating that it is an evolving mesoscale feature.
-By 31 Aug, there is no surface expression of the feature, though tendrils of cooler water can be seen separating from La Perouse Bank.
-By Sep-05 the water gas significantly warmed, but the offshore pushing anomally does not appear to have a surface signature.
+There is a clear surface expression of the separating tongue in satellite imagery (\fref{fig:SST}).
+Water flowing equatorward along the shelf tends to be cooler than offshore water, likely mixing with the colder coming out of the Strait of Juan de Fuca.
+On Aug 25, there is a cold tongue of water separating from La Perouse bank, crossing isobaths and pointing south at 125.8 W.
+There is cooler tendril streaming west at 48 N off the south end of this tongue.
+This feature is not as well-developed in the previous image, Aug 21, perhaps indicating that it is an evolving feature.
+By 31 Aug, there is no surface expression of the feature, though small tendrils of cooler water can be seen separating from La Perouse Bank.
+By Sep-05 the water has significantly warmed, but the offshore anomaly does not appear to have a surface signature.
 
@@
     \includegraphics[width=6in]{SSTSnaps}
-    \caption{Sea surface temperature images from the observation period.
+    \caption{Sea surface temperature snapshots from the observation period.
 Grey areas are clouds; dashed gray line is the study area.
@@
 
-Ccean chlorophyll estimates show the same feature (\fref{fig:ChlA}) demostrating the advection of a region of high chlorophyll to the west side of the eddy.
-It also shows a relatively high-chlorophyll squirt to the west, again exiting the study region at approximately 48 degrees N.
+Satellite-based surface chlorophyll estimates show the same feature (\fref{fig:ChlA}) demonstrating the advection of high chlorophyll to the west side of the Eddy.
+It also shows a relatively high-chlorophyll tendril to the west, again exiting the study region at approximately 48 degrees N.
 The feature is relatively long-lived, on the order of one month.
-Inspection of images before 2013-08-05 were too obscured by clouds, or did not show this feature, and even by 2013-09-06 we see the feature fading from the satellite.
-However, note that this feature is somewhat south of tongue that we observe deeper in the water column, but is indicative of substantial offshore transport in the region.
+Inspection of images before 2013-08-05 were too obscured by clouds, or did not show this feature.
+By 2013-09-06, we see the feature fading from the satellite.
+However, note that this feature is centered 0.2 degrees of latitude south of tongue that we observe deeper in the water column, again indicating that there is some depth-dependence to the feature.
 
@@
 
-The results of this fine-scaled survey perhaps pose more questions than they answer, but the questions are important to set the processes that should be looked for in understanding the circulation of the southern Vancouver Island shelf.
+The intensive sampling discussed here has demonstrated a few important features of the South Vancouver Island Shelf.
+First, there is a sharp and persistent temperature-salinity compensated front between offshore water, and water that has been termed Eddy water.
+This is readily determined in $\theta-S$ space, with the Eddy water lying along a straight mixing line, and the offshore water having a sharp kink centered at $26.4\,\mathrm{kg\,m^{-3}}$.
+Second, there was not strong evidence of water moving up the Spur Canyon during our observations, but the Spur Canyon was a site of hydraulic cross-canyon flows in which we infer significant mixing.
+Finally, upstream of the Eddy region water in the equatorward shelf current is partially mixed in $\theta-S$ space, and is seen to separate from the shelf at the point of an abrupt bend in an underwater bank.
+The water mass crosses isobaths and is ejected into the interior; this can also be observed from satellites.
+Thus the water that is offshore of the Eddy appears to have been brought onto the shelf in exchange with this offshore tongue.
 
@@
 
-The water we have been referring to as the Juan de Fuca Eddy is a relatively well-mixed water mass, as evidenced by the water therein lying along a mixing-line in $\theta$--$S$ space relative to water offshore (\fref{fig:LaPerouse2013TS} and \fref{fig:TSdensSpice}).
-Water coming equatorward along the shelf is closer to offshore water, though also shows clear signs of modification due to mixing.
-
-Unfortunately, this data set cannot definitively track where the water in the eddy originated from.
+The source of water and formation of the Juan de Fuca Eddy has received substantial attention, however, the observations of a tight mixing line in $\theta$--$S$ space presented here are a new observation.
 The deepest water in the eddy could originate along the $\theta$--$S$ line from approximately 5.5 degrees to 7.5 degrees (\fref{fig:LaPerouse2013TS}a), which in the open ocean spans depths from 420 m to 70 m.
@@
 As a best guess, if the water in the deepest part of the eddy came from a vertical mixture of the water at $26.5\,\mathrm{kg\,m^{-3}}$ and an equal distance down in density space of $26.7\,\mathrm{kg\,m^{-3}}$, then the deepest water in the eddy would be coming from a depth of 250 m.
-This is actually a relatively typical upwelling depth for coastal flows, and may not require extra input up Spur canyon as posited by \citeA{freelanddenman82}.
+This is a typical upwelling depth for coastal flows, and may not require extra input up Spur canyon as posited by \citeA{freelanddenman82}.
 
-We found little evidence of vigorous flow up the Spur Canyon, or if there is, then the mixing in the canyon is quick enough to remove the $\theta$--$S$ signature of offshore water within 20 km (\fref{fig:CanyonPropertiesSpice}).
+We found little evidence of vigorous flow up the Spur Canyon, or if there is, then the mixing in the canyon is quick enough to remove the $\theta$--$S$ signature of offshore water within 20 km of the canyon mouth (\fref{fig:CanyonPropertiesSpice}).
 We did find substantial evidence of mixing in the canyon, but the primary route of high-spice water into the canyon appears to be due to tidal flow over the banks on the west side (\fref{fig:HydraulicsCanyon}), rather than evidence of flow up the canyon.
-However, it is also of note that upwelling winds had ceased at this point (\fref{fig:LaPeWind}), meaning that the offshore pressure gradient may be reduced, leading to less ageostrophic upwelling in the canyon.
+However, it is also of note that upwelling winds had ceased at this point (\fref{fig:LaPeWind}), meaning that the offshore surface pressure gradient may be reduced, leading to reduced ageostrophic upwelling in the canyon.
 Whether the cessation of winds would also lead to a reduction of the low sea level height in the center of the eddy is an open question.
@@
 There is evidence of aging of the water in the eddy between late spring and late summer (\fref{fig:LaPerouse2013Ctd}), with a reduction of oxygen in the eddy.
-If we posited that the reduction was all from the same water, then the oxygen consumption rate over the time between the late-May and early September cruises would be on the order of $0.5\ \mathrm{\mu mol\, kg^{-1} d^{-1}}$.
-This seems on the low side of continental upwellng system estimates of apparent oxygen utilization rates, which are between $1$ and $5\  \mathrm{\mu mol\, kg^{-1} d^{-1}}$ \cite{dortchetal94}.
-So it seems possible the eddy has enough exchange with the surrounding water to have a shorter residence time.
+If we posited that the reduction was all in the same water, then the oxygen consumption rate over the time between the late-May and early September cruises would be on the order of $0.5\ \mathrm{\mu mol\, kg^{-1} d^{-1}}$.
+This seems on the low side of continental upwelling system estimates of apparent oxygen utilization rates, which are between $1$ and $5\  \mathrm{\mu mol\, kg^{-1} d^{-1}}$ \cite{dortchetal94}.
+So it seems possible the eddy has enough exchange with the surrounding water for the residence time to be less than the whole time period from late May to September.
+Note also that during the June cruise, the water in the Eddy is slightly cooler than the mixing line, and during the September cruise is slightly warmer than the mixing line, so the water in the eddy is evolving seasonally, and probably affected by the water temperature in the Vancouver Island Coastal Current, and the amount of water getting through the front.
 
-There is partial mixing of the water in the eddy, though it remains vertically stratified, and the amount of homogenization is such that either the mixing is very strong, or the water is retained in the eddy for a long time.
+The water is  mixed enough in the Eddy that it falls along a mixing line, though it remains vertically stratified.
+The amount of homogenization is such that either the mixing is very strong, or the water is retained in the eddy for a long time.
 The amount of turbulence required to homogenize 100 m of water over 90 days is $\kappa \approx 10^{-3}\ \mathrm{m^2\,s^{-1}}$.
@@
 We can more carefully quantify this by considering a synthetic profile or temperature and salinity based on an offshore profile, extrapolated from the bottom of the 200-m cast to 250 m, assuming that the temperature and salinity at 250 m are $6.2\ \mathrm{[^oC]}$ and $34\ \mathrm{psu}$ respectively (\fref{fig:TSExercise}).
-Water at the offshore station was warmer than onshore, so the profile was also linearly interpolated to a surface value of ($31.6\ \mathrm{psu}$, $12\ \mathrm{^oC}$).
-The profile was then linearly compressed into a depth of 150 m representative of the shelf depth in the dense pool, and subjected to mixing with a constant diffusivity of $\kappa = 10^{-3}\ \mathrm{m^2\,s^{-1}}$, with the surface and bottom values pinned under the assumption that the near-bottom source and surface waters are replenished from a large reservoir.
+Water at the offshore station was warmer than onshore, so the profile was also linearly interpolated to a surface value of ($12\ \mathrm{^oC}$, $31.6\ \mathrm{psu}$).
+The profile was then linearly compressed into a depth range of 150 m representative of the shelf depth in the dense pool, and subjected to mixing with a constant diffusivity of $\kappa = 10^{-3}\ \mathrm{m^2\,s^{-1}}$, with the surface and bottom values pinned under the assumption that the near-bottom source and surface waters are replenished from a large reservoir.
 Similar to the naive scaling, the T-S relationship does not approach a straight line until after approximately 60-100 days, or until the mixing affects a vertical length scale of $\lambda = \left(\tau \kappa\right)^{1/2}$ that is greater than $\approx 70\,\mathrm{m}$.
@@
 There is not any substantial bathymetry blocking the onshore incursion of water at this location, so there must be a dynamic barrier.
-In contrast, numerical simulations of this region reported by \cite{sahuetal22} using a NEMO 36th-degree regional model indicate that there is relatively rapid exchange between the deep eddy water and the rest of the coastal ocean.
+In contrast, numerical simulations of this region reported by \cite{sahuetal22} using a NEMO 36th-degree regional model  indicate that there is more rapid exchange (order 40 days) between the deep eddy water and the rest of the coastal ocean.
 In this model, the region where the eddy resides has velocities equal or greater than other parts of the shelf, and water has an approximate residence time of less than 20 days.
-Consequently, the eddy water does not develop a distinct $\theta$--$S$ signature from the surrounding water.
+The eddy water in the model does develop a distinct $\theta$--$S$ signature, but not nearly as strong as observed.
+It also has a front with the offshore water, but the front is much wider and more gradual than that observed here.
 
@@
 
-The (apparent) displacement of shelf water from La Perousse Bank is relatively dramatic.
+The (apparent) displacement of shelf water from La Perouse Bank is a dramatic departure from geostrophically balanced isobath-following flow..
 Eddies have been known to separate from irregular coastal topography, both at the surface \cite{barthetal00}, and deeper in the water column \cite{pellandetal13}.
 It has been recognized that instabilities lead to exchange between the open ocean and the shelf at this location \cite{ikedaemery84, ikedaemery84}.
-However, observations of the wholesale replacement of shelf water by a new water mass from offshore, is relatively rare.
+However, observations of the wholesale replacement of shelf water by a new water mass from offshore, is rare.
 In our situation it is clear that water from as deep as 150 m is separating from the shelf and moving offshore (\fref{fig:SpiceO2264}, \fref{fig:IsopycnalsDepth}).
@@
 Most years there are three of four large excursions from the shelf into the open ocean, many of them over 100 km long.
-The lengthscale in these images is somewhat larger than that inferred for this region by \citeA{ikedaetal84}.
+This lengthscale is somewhat larger than 60 km inferred for this region by \citeA{ikedaetal84} using a four-layer instability analysis.
 It is possible that there is even some spatial locking of these features, with a persistent separation at the north tip of the Island, and a strong tendency for one at 49.5 N.
-Our study area also has evidence of a squirt in most of the years, with 2011 being the only clear exception.
+Our study area also has evidence of a separation in most of the years, with 2011 being the only clear exception.
 General baroclinic instability of the upwelling front is a possible mechanism to drive offshore exchange \cite<e.g.>{ikedaetal84,durskiallen05}, but this tends to be shallow, with smaller-scale instabilities that will not extend as far into the interior ocean as observed here.
@@
 Large-scale mixing between the shelf and open ocean has been evident since the satellite era.
-Here we demonstrate that the flow is originating on the shelf and separating from the bathymetry and being injected into the interior down to the bottom of the water column.
-A  similar observation was made by \citeA{barthetal00} downstream of Cape Blanco, Oregon, where the coastal current was observed to detach from the shelf in the lee of the cape, and squirt into the interior.
+Here we demonstrate that in this study area the flow is originating on the shelf and separating from the bathymetry and being injected into the interior down to the bottom of the water column.
+A  similar observation was made by \citeA{barthetal00} downstream of Cape Blanco, Oregon, where the coastal current was observed to detach from the shelf in the lee of the cape, and flow into the interior.
 They hypothesized that as the current moved offshore, it deepened, stretching isopycnals and creating cyclonic relative vorticity that would tend to push the current back onshelf, but then it was caught in the undercurrent and stalled, being pushed offshore.
@@
     \caption{
-      Late August sea-surface temperature, 8-day composites \cite{MODISSST8d}.   From 2011 to 2021; missing years had too much cloud cover or no satellite coverage.
-      \label{fig:SSTLateAug} }
+      Available late-August sea-surface temperature from 8-day composites, 2011 to 2021 \cite{MODISSST8d}; missing years had too much cloud cover or no satellite coverage.  \label{fig:SSTLateAug} }
   \end{center}
@@
 Our observations are a finer detailed representation of the kind of cross-shore transports inferred by \citeA{mackasyelland99} from hydrographic surveys, and definitively show that this water can originate from the shelf from relatively deep depths and be transported offshore.
-We do not have velocity measurements for the water that replaces it, but assuming that water also flows along-shelf, there is a large replacement of low-O2, low-nutrient water with offshore water at this location.
+We do not have velocity measurements for the water that replaces it, but assuming that water also flows along-shelf, there is a large replacement of shelf water with offshore water at this location.
 This emphasizes the importance of three dimensional observations and modeling of cross-shelf dynamics when thinking about biological processes on the shelf.

### 2022-11-08 PW13Paper.tex
Version to SW

@@
 
+\newcommand*{\Eddy}{{\sc Eddy}}
 \draftfalse
@@
 \affiliation{1}{School of Earth and Ocean Sciences, University of Victoria, BC, Canada}
-\affiliation{2}{Department of Physcics \& Astronomy, University of Victoria, BC, Canada}
+\affiliation{2}{Department of Physics \& Astronomy, University of Victoria, BC, Canada}
 \affiliation{3}{Department of Earth, Ocean and Atmospheric Sciences, University of British Columbia, Vancouver, BC, Canada}
@@
 \item Offshore water is drawn onto the shelf and forms a sharp semi-persistent front with the Juan de Fuca Eddy.
-\item The Eddy shows evidence of long residence times, and little evidence of deep-water origin.
+\item The \Eddy\ shows evidence of long residence times, and little evidence of deep-water origin.
 \end{keypoints}
@@
 The observations indicate that the water in this Eddy is a mixture of offshore water and water from a buoyant coastal current.
-The water in the  in the Eddy is well-mixed along a mixing line in $\theta$--$S$ space, though it retains stratification, and is either rapidly mixed (high $\kappa$), or has a long residence time (large $\tau$), such that $(\kappa \tau)^{1/2} \sim 50 -- 100\ \mathrm{m}$.
+The water in the  in the Eddy is well-mixed along a mixing line in $\theta$--$S$ space, though it retains stratification, and is either rapidly mixed (high $\kappa$), or has a long residence time (large $\tau$), such that $(\kappa \tau)^{1/2} \sim 50$ ---  $100\ \mathrm{m}$.
 There is a sharp temperature-salinity front on the offshore side of of this well-mixed water, with the front no more than 1-km wide, and no sign of instabilities.
@@
 The estuarine water flows poleward, hugging the coast as the Vancouver Island Coastal Current, and in the summer shelf water flows equatorward.
-Trapped between these two currents is the Juan de Fuca eddy, a stagnant region of water that occupies the shelf where it abruptly widens just north of the Strait of Juan de Fuca.
+Trapped between these two currents is the Juan de Fuca Eddy, a stagnant region of water that occupies the shelf where it abruptly widens just north of the Strait of Juan de Fuca.
 
-Here we present intensive sampling of the Juan de Fuca eddy region.
-The observations show the eddy is low in oxygen, and we find has undergone substantial vertical and lateral mixing.
+Here we present intensive sampling of the Juan de Fuca Eddy region.
+The observations show the Eddy is low in oxygen, and we find has undergone substantial vertical and lateral mixing.
 This indicates that the waters have a surprisingly long residence time for a coastal feature, and that this region extends the depth of the water column.
@@
 Water from the strait of Juan de Fuca flows poleward as a buoyant current that hugs the coast \cite{thomsonetal89, hickeyetal91}, while shelf water flows equatorward in the summer under upwelling conditions, both forced by local winds and via tele-connections with the long, homogenous shelves equatorward off Washington and Oregon \cite{hickeyetal91,thomsonkrassovski15,engidaetal16}.
-Trapped between these two currents is a region of relatively homogenous water that has been termed the Juan de Fuca Eddy \cite{freelanddenman82,freelandmcintosh89,foremanetal07,macfadyenhickey10}.
+Trapped between these two currents is a region of relatively homogenous water that has been termed the Juan de Fuca Eddy \cite{freelanddenman82,freelandmcintosh89,foremanetal07,macfadyenhickey10}, denoted ``\Eddy'' below.
 
-A goal of our study was to understand how the Juan de Fuca Eddy persists, and how it exchanges water properties with offshore water.
-Low oxygen has previously been found in the Eddy, causing it to be inferred that water in the eddy is upwelled from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
-This inference was based on the oxygen being conservative, and hence having to come from close to the oxygen minimum zone found further offshore \cite{mackasetal87} and lead to hypotheses that perhaps the eddy was fed by waters being drawn up a spur canyon (called the Spur Canyon) via ageostrophic transport from offshore to onshore due to the low pressure in the eddy center \cite{weaverhsieh87}.
+A goal of our study was to understand how the \Eddy persists, and how it exchanges water properties with offshore water.
+Low oxygen has previously been found in the \Eddy, causing it to be inferred that water in the \Eddy\ is upwelled from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
+This inference was based on the oxygen being conservative, and hence having to come from close to the oxygen minimum zone found further offshore \cite{mackasetal87} and lead to hypotheses that perhaps the \Eddy\ was fed by waters being drawn up a spur canyon (called the Spur Canyon) via ageostrophic transport from offshore to onshore due to the low pressure in the \Eddy\ center \cite{weaverhsieh87}.
 Below, we will argue that there is not substantial evidence of such transport, and that oxygen is likely low because of consumption due to respiration.
@@
 Net cross-shore transport of nutrients and chlorophyll have been found in hydrographic surveys off Vancouver Island \cite{mackasyelland99}, and associated with mesoscale features both in geostrophic velocities, and in satellite observations.
-The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics \cite{dalebarth01} leading to along-shore trapping of coastal waves,  wind stress curl variations \cite{castelaobarth07}, induced relative vorticity due to stretching of parcels that inertially overshoot their initial isobaths due to a sudden change in downstream bathymetry, and the non-linear breaking of large-scale meanders due to baroclinic instability between the wind-driven current and the California Undercurrent \cite{ikedaetal84, batteen97}.
+The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics \cite{dalebarth01} leading to along-shore trapping of coastal waves,  wind stress curl variations \cite{castelaobarth07}, induced relative vorticity due to stretching of parcels that overshoot their initial isobaths due to a sudden change in downstream bathymetry, and the non-linear breaking of large-scale meanders due to baroclinic instability between the wind-driven current and the California Undercurrent \cite{ikedaetal84, batteen97}.
 
-In this paper we present  observations from 2013 (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes, with a focus on the offshore side of the eddy region (\fref{sec:Observations}) where we consider the age of the eddy, the steadiness of the front separating the eddy and the offshore water, properties along the spur canyon that was believed to feed the eddy, and finally describe a large offshore tongue of the shelf-break current and its intrusion into the interior.
-We discuss the origin of the eddy and the implications of the separating jet (\fref{sec:Summary}).
+In this paper we present  observations from 2013 (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes, with a focus on the offshore side of the \Eddy\ region (\fref{sec:Observations}) where we consider the age of the \Eddy, the steadiness of the front separating the \Eddy\ and the offshore water, properties along the spur canyon that was believed to feed the \Eddy, and finally describe a large offshore tongue of the shelf-break current and its intrusion into the interior.
+We discuss the origin of the \Eddy\ and the implications of the separating jet (\fref{sec:Summary}).
 
@@
 Hydrographic sections along the LB and LC lines highlight summer conditions on the Southern Vancouver Island Shelf, and indicate some of the features we are focusing on in this paper (\Fref{fig:LaPerouse2013Ctd}).
-During both cruises and at both lines there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m depth to shallower than 85 m over the shelf.
+During both cruises, and at both lines, there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m depth offshore to shallower than 85 m over the shelf.
 This dense water tends to be low in oxygen and cool.
-Further onshore, isopycnals tilt back down, and the water warms, evidence of the Vancouver Island Coastal Current that hugs the coast here.
+Further onshore the Vancouver Island Coastal Current hugs the coast where isopycnals tilt back down, and the water is warmer than offshore.
 
@@
 
-The water upstream of the Eddy region, along the LC line appears less mixed than the Eddy water.
+The water upstream of the \Eddy\ region, along the LC line, is less mixed than the \Eddy\ water.
 It does not show as much drawdown of oxygen, and is warmer, at least offshore of La Perouse bank (\fref{fig:LaPerouse2013Ctd}, top three rows).
@@
 The $\theta$--$S$ properties show many of the features demonstrated above (\fref{fig:LaPerouse2013TS}a).
-Below $26.6\ \mathrm{kg\,m^{-3}}$ or so, the deep water masses are found offshore, and are largely homogenous.
-Shallower, there are distinct differences between the water found offshore, with it being warmer and saltier than water on the shelf at the same densities.
-These anomalies are hard to parse directly, so in this paper we use a spice anomaly, relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}).
-This line represents an approximate mixing line for water in the Eddy, and passes the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$.
+Below $26.6\ \mathrm{kg\,m^{-3}}$, the deep water masses found offshore are largely homogenous.
+Shallower, there are distinct differences between warm and salty offshore compared to water on the shelf at the same densities.
+In this paper we use a spice anomaly relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}) representing an approximate mixing line for water in the \Eddy.
+It passes through the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$.
 Spice anomaly for a given water sample at density $\sigma_{\theta}$ is given by $\gamma = \alpha \left(T - T_0\left(\sigma_{\theta}\right)\right) + \beta \left(S - S_0\left(\sigma_{\theta}\right)\right)$ where $T_0\left(\sigma_{\theta}\right)$, where $S_0\left(\sigma_{\theta}\right)$ are the temperature and salinity along the mixing line at the same density as the water sample, with the sign convention that positive is warmer and saltier than data along the mixing line.
@@
 
-Offshore water tends to be high is absolute spice anomaly (\fref{fig:LaPerouse2013Ctd}, middle rows), with a positive-anomaly layer (red) centered at $26.4\ \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and below.
+Offshore water tends to be high in absolute spice anomaly (\fref{fig:LaPerouse2013Ctd}, middle rows), with a positive-anomaly layer (red) centered at $26.4\ \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and below.
 On the shelf ($-10\, \mathrm{km} < X < 30\, \mathrm{km}$), the distinct $\theta$--$S$ masses are attenuated and much closer to the defined mixing line than the offshore water (closer to white in color).
@@
 
-The June and September cruises are quite similar in water properties, particularly in the offshore waters.
+The June and September cruises have quite similar water properties, particularly in the offshore waters.
 Water has upwelled from deeper depths in September, but has most of the same water mass characteristics as earlier in the year.
@@
   \begin{center}
-
      \includegraphics[width=3.5in]{TSdensSpice.pdf}
@@
 Grey contours are potential density relative to the surface, every $0.1\ \mathrm{kg\,m^{-3}}$, and the magenta contour is the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.
-Spice anomaly, $\gamma$, is defined as temperature anomaly scaled by the thermal expansion co-efficient, $\alpha$, along isopycnals from the mixing line found in the well-mixed region in the Eddy (white line), in $\mathrm{kg\,m^{-3}}$.
+Spice anomaly, $\gamma$, is defined as temperature anomaly scaled by the thermal expansion co-efficient, $\alpha$, along isopycnals from the mixing line found in the well-mixed region in the \Eddy\ (white line), in $\mathrm{kg\,m^{-3}}$.
     \label{fig:TSdensSpice}}
@@
 
-In synthetic cross sections of spice anomaly, the Eddy is clearly identifiable as water that is along the straight mixing line, and thus has low-spice anomaly ($\gamma \approx 0\ \mathrm{kg\,m^{-3}}$, \fref{fig:CrossSectionsSpice}c, d).
+In synthetic cross sections of spice anomaly, the \Eddy\ is clearly identifiable as water that is along the straight mixing line, and thus has low-spice anomaly ($\gamma \approx 0\ \mathrm{kg\,m^{-3}}$, \fref{fig:CrossSectionsSpice}c, d).
 This signature of well-mixed water extends from the seafloor to approximately 30 m depth, onshore of $x \approx 0\ \mathrm{km}$.
-Despite lying along a mixing line, the Eddy water is still stratified.
-This well-mixed water contrasts with water directly offshore on the other side of a sharp $\theta-S$ compensated front that is warmer and saltier than the well mixed water ($\gamma > 0.2\ \mathrm{kg\,m^{-3}}$) deeper than $26\,\mathrm{kg\,m^{-3}}$, and colder and fresher shallower.
-The thickness of this front is even sharper in individual sections than in these composite sections (see \fref{sec:frontsurvey}).
+Despite lying along a mixing line, the \Eddy\ water is still stratified.
+This partially-mixed water contrasts with water directly offshore, on the other side of a sharp $\theta-S$ compensated front, that is warmer and saltier than the mixed water ($\gamma > 0.2\ \mathrm{kg\,m^{-3}}$) deeper than $26\,\mathrm{kg\,m^{-3}}$, and colder and fresher shallower.
+This front is even sharper in individual sections than in these composite sections (see \fref{sec:frontsurvey}).
 
-In contrast, the third population of partially mixed water mass identified in \fref{fig:TSdensSpice} is all found upstream of the Eddy region (\fref{fig:CrossSectionsSpice}a, b) as a weaker spice anomaly ($\gamma \approx 0.1\ \mathrm{kg\,m^{-3}}$, ``pink'' colors).
+In contrast, the third population of partially mixed water mass identified in \fref{fig:TSdensSpice} is all found upstream of the \Eddy\ region (\fref{fig:CrossSectionsSpice}a, b) as a weaker spice anomaly ($\gamma \approx 0.1\ \mathrm{kg\,m^{-3}}$, ``pink'' colors).
 This water is generally found near the intersection of the upwelling water and La Perouse Bank, and is not as warm as offshore water, indicating some mixing has taken place.
@@
 
-Oxygen section show similar trends, with the low-spice anomaly water having low oxygen saturations, while the strong negative spice anomaly shallower than $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ has very high oxygen, and the deeper water is also more oxygenated.
-Of note is that compared with the coarser sampling along the LC Line (\fref{fig:LaPerouse2013Ctd}), we see that the abrupt transition between offshore and Eddy water is also present in oxygen.
+Oxygen section show similar trends, with the low-spice anomaly water also having low oxygen saturations, while the strong negative spice anomaly shallower than $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ has very high oxygen; deeper water is also more oxygenated.
+Of note is that compared with the coarser sampling along the LC Line (\fref{fig:LaPerouse2013Ctd}), we see that the abrupt transition between offshore and \Eddy\ water is also present in oxygen.
 
@@
 
-Upstream of La Perouse bank and Barkley Canyon, the water along the shelf the weaker spice anomaly of the third water mass ($\gamma \approx 0.1\,\mathrm{kg\,m^{-3}}$, light pink colors).
-This water is also somewhat lower in oxygen than water from offshore, though not as depleted as the water in the Eddy.
-This water appears to be pushed offshore just upstream of Barkley Canyon, and replaced on the shelf by the warmer offshore water.
+Upstream of La Perouse bank and Barkley Canyon, the water along the shelf has the weaker spice anomaly of the third water mass ($\gamma \approx 0.1\,\mathrm{kg\,m^{-3}}$, light pink colors).
+This water is also somewhat lower in oxygen than water from offshore, though not as depleted as the water in the \Eddy.
+This water appears to be pushed offshore just upstream of Barkley Canyon, and replaced on the shelf by the warmer (high spice) offshore water.
 
@@
 
-\subsubsection{Frontal survey}
-\label{sec:frontsurvey}
-
-A systematic survey through the front between the offshore water and the Eddy water demonstrates the structure of the front (\fref{fig:Frontsurvey}).
-The survey started close to shore and passed through the Vancouver Island Coastal Current (along-track 0-10 km).
-The coastal current water is buoyant and significantly fresher and cooler than water at the same density near the surface and forms a classic density front with
-
-The front between the offshore water and the ``eddy'' water was passed through 10 times, showing its evolution following the along-shore equatorward flow.
-First, as noted in the composite sections, isopycnals slope up from offshore onto the shelf.
-This is in agreement with a shelf-break current, though the isopycnals flatten out in the Eddy region.
-The front starts out very sharp, (along-track distance $50\ \mathrm{km}$) though two small tendrils can be seen separating from the front on the inshore side.
-Similar tendrils are found on the second pass, perhaps a bit more separated from the front ($\approx 80\ \mathrm{km}$), and on the third pass ($\approx 95\ \mathrm{km}$).
-Note that these tendrils are made of up partially mixed water.
-The subsequent passes have more of this partially mixed water, such that the partially mixed region is up to 5-km wide by the fifth pass ($\approx 150\ \mathrm{km}$).
-However, note that the deeper isopycnals retain a sharp front, and indeed the front appears sharp again by the seventh pass at all depths.
-
-There is evidence of some warmer water swirling into Eddy, particularly along isopycnals deeper than $26.4\,\mathrm{kg\,m^{-3}}$.
-Regions of warmer (and more oxygenated) water are found in tendrils at these depths (e.g.\ $\approx\,170\ \mathrm{km}$ and $\approx 185\,\mathrm{km}$).
-
-The overall effect is similar to what was seen in the Gulf Stream with similar observations \cite{klymaketal16}; there are two quite distinct water masses, the Eddy water and the offshore water, as seen in the $\theta$--$S$ plot (\fref{fig:Frontsurvey}c).
-There is only a small population of samples between these two, indicative of substantial isopycnal and vertical mixing, but even these populations are relatively cut off from the main water masses, indicating that they are well-mixed on their own, in short episodic events.
-
-\begin{figure*}[htbp]
-  \begin{center}
-    \includegraphics[width=6.2in]{Frontsurvey}
-    \caption{Data from a survey focusing on the front region (2013-08-28T15:11 to 2013-08-29T09:18) a) spice anomaly along $26.4\,\mathrm{kg\,m^{-3}}$.
-Grey-white alternating lines are the coordinate system, with 10-km alternating shades.
-b) oxygen saturation; magenta dots correspond to distance along-track.
-c) density of data points (logscale) in this section of data.
-d) cross-section of spice anomaly along track.
-Grey isopycnals are contoured every $0.5\,\mathrm{kg\,m^{-3}}$, and the $26.4\,\mathrm{kg\,m^{-3}}$ isopycnal is shown in magenta.
-e) cross-section of oxygen saturation.
-      \label{fig:Frontsurvey} }
-  \end{center}
-\end{figure*}
-
 \subsection{Spur Canyon}
 
-The Spur Canyon leading from the Strait of Juan de Fuca has been implicated in allowing dense water to be upwelled into the Eddy.
-The seasurface is low in the middle of the Eddy, so water has been hypothesized to move up the Spur Canyon due to ageostrophic motion \cite{weaverhsieh87,freelanddenman82}.
-This is difficult to infer from water properties alone.
+The Spur Canyon leading from the Strait of Juan de Fuca has been implicated in allowing dense water to be upwelled into the \Eddy.
+The sea surface is low in the middle of the \Eddy, so water has been hypothesized to move up the Spur Canyon due to ageostrophic motion \cite{weaverhsieh87,freelanddenman82}.
+This is difficult to infer from the observations collected here.
 Three transects up the canyon indicate that deep isopycnals slope up into the canyon (\fref{fig:CanyonPropertiesSpice}, to approximately 30 km).
-Across the eddy, isopycnals are largely flat until they intersect the Vancouver Island Coastal Current ($x\approx 80\ \mathrm{km}$).
-The deeper isopycnals do not make it into the deeper basin northeast of La Perouse bank ($x\approx 70\ \mathrm{km}$).
+Continuing across the \Eddy, isopycnals are largely flat until they intersect the Vancouver Island Coastal Current ($x\approx 80\ \mathrm{km}$).
+The deeper isopycnals do not make it into the deeper basin northeast of La Perouse bank ($x\approx 70\ \mathrm{km}$),  thus the bank is a natural northern boundary of the \Eddy.
 
@@
 
-Water properties along the canyon go from offshore water, to Eddy water.
-Oxygen behaves in a similar manner, though some of the incoming water has slightly lower oxygen than water inside the eddy (not shown).
-Based on these sections, it is difficult to infer water motion up the canyon in particular.
+Water properties along the canyon go from offshore water, to \Eddy\ water.
+Oxygen behaves in a similar manner, though some of the incoming water has slightly lower oxygen than water inside the \Eddy\ (not shown).
+Based on these sections, it is difficult to infer water motion up the canyon.
 
@@
 
-This cross-canyon mixing makes it ambiguous if there is water moving up the canyon or not.
-There is indeed a general tendency along the canyon for higher spice water to be found offshore (\fref{fig:CanyonPropertiesSpice}) but it seems likely the source of the higher spice is over the bank rather than water being advected up the canyon.
+The cross-canyon turbulence makes it ambiguous if there is water moving up the canyon or not.
+There is a general tendency along the canyon for higher spice water to be found offshore (\fref{fig:CanyonPropertiesSpice}) but it seems likely the source of the higher spice is over the bank rather than water being advected up the canyon.
+This tidally driven flow over the bank is the strongest source of high-spice offshore water into the eddy region identified during our surveys.
 
-\subsection{Offshore tongue}
+\subsubsection{Frontal survey}
+\label{sec:frontsurvey}
 
-As discussed above, the very sharp T/S front between the well-mixed Eddy water and the offshore water  persists, despite there being considerable partially-mixed water found upstream along the shelf.
-The partially-mixed water must go somewhere, and indeed it appears to largely be pushed offshore just upstream of Barkley Canyon (\fref{fig:IsopycnalsDepth}).
-There is a tongue of partially mixed ($\gamma$ is pink along $26.4\,\mathrm{kg\,m^{-3}}$) that flows almost due south just west of 126 W.
+A systematic survey through the front between the offshore water and the \Eddy\ water demonstrates the sharpness and persistence of this front (\fref{fig:Frontsurvey}), giving evidence that it has limited exchange with the offshore region.
+The survey started close to shore and passed through the Vancouver Island Coastal Current (along-track 0-10 km).
+The coastal current water is buoyant and significantly fresher and cooler than water at the same density near the surface and forms a classic density front with the water in the Eddy region.
+There is some partially mixed (pink in spice spice) water in this density front due to mixing with the warmer water near the surface.
+
+Offshore of this coastal current, measurements were collected through the front between the offshore water and the \Eddy\ water 10 times, showing its evolution following the along-shore equatorward flow.
+First, as noted in the composite sections, isopycnals slope up from offshore onto the shelf.
+The front starts out very sharp, (along-track distance $50\ \mathrm{km}$) though two small tendrils can be seen separating from the front on the inshore side.
+Similar tendrils are found on the second pass, perhaps a bit more separated from the front ($\approx 80\ \mathrm{km}$), and on the third pass ($\approx 95\ \mathrm{km}$).
+These tendrils are made of up partially mixed water.
+The subsequent passes have more of this partially mixed water, such that the partially mixed region is up to 5-km wide by the fifth pass ($\approx 150\ \mathrm{km}$).
+However, note that the deeper isopycnals retain a sharp front, and indeed the front appears sharp again by the seventh pass at all depths.
+
+There is evidence of some warmer water swirling into Eddy, particularly along isopycnals deeper than $26.4\,\mathrm{kg\,m^{-3}}$.
+Regions of warmer (and more oxygenated) water are found in tendrils at these depths (e.g.\ $\approx\,170\ \mathrm{km}$ and $\approx 185\,\mathrm{km}$).
+The overall effect is similar to what was seen in the Gulf Stream with similar observations \cite{klymaketal16}; there are two quite distinct water masses, the Eddy water and the offshore water, as seen in the $\theta$--$S$ plot (\fref{fig:Frontsurvey}c).
+There is only a small population of samples between these two, indicative of substantial isopycnal and vertical mixing, but even these populations are relatively cut off from the main water masses, indicating that they are well-mixed on their own, in short episodic events.
+However, there is little evidence of large-scale flow into or out of the Eddy at all depths during this observation period.
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=6.2in]{Frontsurvey}
+    \caption{Data from a survey focusing on the front region (2013-08-28T15:11 to 2013-08-29T09:18) a) spice anomaly along $26.4\,\mathrm{kg\,m^{-3}}$.
+Grey-white alternating lines are the coordinate system, with 10-km alternating shades.
+b) oxygen saturation; magenta dots correspond to distance along-track.
+c) density of data points (logscale) in this section of data.
+d) cross-section of spice anomaly along track.
+Grey isopycnals are contoured every $0.5\,\mathrm{kg\,m^{-3}}$, and the $26.4\,\mathrm{kg\,m^{-3}}$ isopycnal is shown in magenta.
+e) cross-section of oxygen saturation.
+      \label{fig:Frontsurvey} }
+  \end{center}
+\end{figure*}
+
+\subsection{Separation of coastal water}
+
+Upstream of the the sharp front between the low-spice \Eddy\ and the high-spice offshore water is a substantial mass of intermediate-spice water along the shelf (compare \fref{fig:CrossSectionsSpice}a and c).
+If in geostrophic balance, intermediate spice water is moving equatorward along the shelf upstream of the \Eddy, but is pushed offshore just upstream of Barkley Canyon (\fref{fig:IsopycnalsDepth}).
+There is a tongue of intermediate-spice water ($\gamma$ is pink along $26.4\,\mathrm{kg\,m^{-3}}$) that separates from the shelf just west of 126 W.
 We had trouble crossing the whole tongue with the scale of our surveys, but it is at least 30 km wide.
 It also appears to end at approximately 48.2 N.
-This partially mixed water reaches from relatively shallow isopycnals to at least $26.6\,\mathrm{kg\,m^{-3}}$ (\fref{fig:IsopycnalsDepth}, left panel).
+This intermediate-spice water reaches from relatively shallow isopycnals to at least $26.6\,\mathrm{kg\,m^{-3}}$ (\fref{fig:IsopycnalsDepth}, left panel).
 Unfortunately, we cannot track the fate of this water mass because isopycnals tilt down offshore, below the depth limit of the MVP.
@@
 
-This feature is embedded in the larger scale isopycnal tilt caused by the upwelling (\fref{fig:IsopycnalsDepth} right panels), so it is difficult to see dynamically what is driving this offshore push.
+This separating tongue is embedded in the larger scale isopycnal tilt caused by the upwelling (\fref{fig:IsopycnalsDepth} right panels), so it is difficult to see dynamically what is driving this offshore push.
 One possibility is that it is simple flow separation as the water is not able to make the sharp turn around La Perouse Bank.
@@
 There is cooler tendril streaming west at 48 N off the south end of this tongue.
-This feature is not as well-developed in the previous image, Aug 21, perhaps indicating that it is an evolving feature.
-By 31 Aug, there is no surface expression of the feature, though small tendrils of cooler water can be seen separating from La Perouse Bank.
-By Sep-05 the water has significantly warmed, but the offshore anomaly does not appear to have a surface signature.
+This feature is not as well-developed in the previous image (21 August) perhaps indicating that it is an evolving feature.
+By 31 August, there is no surface expression of the feature, though small tendrils of cooler water can be seen separating from La Perouse Bank.
+By 5 September, the water has significantly warmed, but the offshore anomaly does not appear to have a surface signature.
 
@@
 
-Satellite-based surface chlorophyll estimates show the same feature (\fref{fig:ChlA}) demonstrating the advection of high chlorophyll to the west side of the Eddy.
+Satellite-based surface chlorophyll estimates show the same feature (\fref{fig:ChlA}) demonstrating the advection of high chlorophyll to the west side of the \Eddy.
 It also shows a relatively high-chlorophyll tendril to the west, again exiting the study region at approximately 48 degrees N.
@@
 The intensive sampling discussed here has demonstrated a few important features of the South Vancouver Island Shelf.
-First, there is a sharp and persistent temperature-salinity compensated front between offshore water, and water that has been termed Eddy water.
-This is readily determined in $\theta-S$ space, with the Eddy water lying along a straight mixing line, and the offshore water having a sharp kink centered at $26.4\,\mathrm{kg\,m^{-3}}$.
-Second, there was not strong evidence of water moving up the Spur Canyon during our observations, but the Spur Canyon was a site of hydraulic cross-canyon flows in which we infer significant mixing.
-Finally, upstream of the Eddy region water in the equatorward shelf current is partially mixed in $\theta-S$ space, and is seen to separate from the shelf at the point of an abrupt bend in an underwater bank.
-The water mass crosses isobaths and is ejected into the interior; this can also be observed from satellites.
-Thus the water that is offshore of the Eddy appears to have been brought onto the shelf in exchange with this offshore tongue.
+The Juan de Fuca Eddy water is readily identified as well-mixed withe stratified water found along a straight line in $\theta-S$ space, compared with offshore water that was warmer and saltier (high-spice).
+There was not strong evidence of the \Eddy\ being supplied by water moving up the Spur Canyon during our observations, but the Spur Canyon was a site of hydraulic cross-canyon flows in which we infer significant mixing.
+There is a sharp and persistent temperature-salinity compensated front between offshore water, and the mixed \Eddy\ water.
+Finally, upstream of the \Eddy, water in the equatorward shelf current has intermediate spice, and is seen to separate from the shelf at the point of an abrupt bend in an underwater bank.
+The water mass crosses isobaths and is ejected into the interior, also be observed from satellites.
+Thus the water that is offshore of the \Eddy\ appears to have been brought onto the shelf in exchange with this offshore tongue.
 
-\subsection{Age and source of the eddy water}
+\subsection{Age and source of the \Eddy\ water}
 
 The source of water and formation of the Juan de Fuca Eddy has received substantial attention, however, the observations of a tight mixing line in $\theta$--$S$ space presented here are a new observation.
-The deepest water in the eddy could originate along the $\theta$--$S$ line from approximately 5.5 degrees to 7.5 degrees (\fref{fig:LaPerouse2013TS}a), which in the open ocean spans depths from 420 m to 70 m.
+The deepest water in the \Eddy\ could originate along the $\theta$--$S$ line from approximately 5.5 degrees to 7.5 degrees (\fref{fig:LaPerouse2013TS}a), which in the open ocean spans depths from 420 m to 70 m.
 \citeA{mackasetal87} attempted to determine the origin of the water by including oxygen as a third variable to resolve the ambiguous $\theta$--$S$ relation.
-However, as \fref{fig:LaPerouse2013TS}b makes clear, oxygen is definitely not a conserved property in the eddy, with concentrations up to $150 \ \mathrm{\mu mol\ kg^{1}}$ lower in the eddy than the water found offshore, and in a way that definitely cannot result from conservative mixing.
-As a best guess, if the water in the deepest part of the eddy came from a vertical mixture of the water at $26.5\,\mathrm{kg\,m^{-3}}$ and an equal distance down in density space of $26.7\,\mathrm{kg\,m^{-3}}$, then the deepest water in the eddy would be coming from a depth of 250 m.
-This is a typical upwelling depth for coastal flows, and may not require extra input up Spur canyon as posited by \citeA{freelanddenman82}.
+However, as \fref{fig:LaPerouse2013TS}b makes clear, oxygen is definitely not a conserved property in the \Eddy, with concentrations up to $150 \ \mathrm{\mu mol\ kg^{1}}$ lower in the \Eddy\ than the water found offshore, and in a way that definitely cannot result from conservative mixing.
+As a best guess, if the water in the deepest part of the \Eddy\ came from a vertical mixture of the water at $26.5\,\mathrm{kg\,m^{-3}}$ and an equal distance down in density space of $26.7\,\mathrm{kg\,m^{-3}}$, then the deepest water in the \Eddy\ may be coming from a depth of 250 m.
+This is a typical upwelling depth for coastal flows, and may not require extra input up the Spur Canyon as posited by \citeA{freelanddenman82}.
 
@@
 However, it is also of note that upwelling winds had ceased at this point (\fref{fig:LaPeWind}), meaning that the offshore surface pressure gradient may be reduced, leading to reduced ageostrophic upwelling in the canyon.
-Whether the cessation of winds would also lead to a reduction of the low sea level height in the center of the eddy is an open question.
+Whether the cessation of winds would also lead to a reduction of the low sea level height in the center of the \Eddy\ that may drive up-canyon flow is an open question.
 
-There is evidence of aging of the water in the eddy between late spring and late summer (\fref{fig:LaPerouse2013Ctd}), with a reduction of oxygen in the eddy.
+There is evidence of aging of the water in the \Eddy\ between late spring and late summer (\fref{fig:LaPerouse2013Ctd}), with a reduction of oxygen in the \Eddy.
 If we posited that the reduction was all in the same water, then the oxygen consumption rate over the time between the late-May and early September cruises would be on the order of $0.5\ \mathrm{\mu mol\, kg^{-1} d^{-1}}$.
 This seems on the low side of continental upwelling system estimates of apparent oxygen utilization rates, which are between $1$ and $5\  \mathrm{\mu mol\, kg^{-1} d^{-1}}$ \cite{dortchetal94}.
-So it seems possible the eddy has enough exchange with the surrounding water for the residence time to be less than the whole time period from late May to September.
-Note also that during the June cruise, the water in the Eddy is slightly cooler than the mixing line, and during the September cruise is slightly warmer than the mixing line, so the water in the eddy is evolving seasonally, and probably affected by the water temperature in the Vancouver Island Coastal Current, and the amount of water getting through the front.
+So it seems possible the \Eddy\ has enough exchange with the surrounding water for the residence time to be less than the whole time period from late May to September.
+Note also that during the June cruise, the water in the \Eddy\ is slightly cooler than the mixing line, and during the September cruise is slightly warmer than the mixing line, so the water in the \Eddy\ is evolving seasonally, and probably affected by the water temperature in the Vancouver Island Coastal Current, and the amount of water getting through the front.
 
-The water is  mixed enough in the Eddy that it falls along a mixing line, though it remains vertically stratified.
-The amount of homogenization is such that either the mixing is very strong, or the water is retained in the eddy for a long time.
+The water is  mixed enough in the \Eddy\ that it falls along a mixing line, though it remains vertically stratified.
+The amount of homogenization is such that either the mixing is very strong, or the water is retained in the \Eddy\ for a long time.
 The amount of turbulence required to homogenize 100 m of water over 90 days is $\kappa \approx 10^{-3}\ \mathrm{m^2\,s^{-1}}$.
@@
 
-The sharpness of the front with the eddy and the offshore water is also intriguing.
-It was persistent for the duration of our detailed survey (\fref{fig:SpiceO2264}), and, so far as we can tell with the limited resolution, was present during the La Perouse cruises.
+The sharpness of the front with the \Eddy\ and the offshore water is intriguing.
+It was persistent for the duration of our detailed survey (\fref{fig:SpiceO2264}), and, so far as we can tell with the limited resolution, was present during the bracketing La Perouse cruises.
 There is not any substantial bathymetry blocking the onshore incursion of water at this location, so there must be a dynamic barrier.
-In contrast, numerical simulations of this region reported by \cite{sahuetal22} using a NEMO 36th-degree regional model  indicate that there is more rapid exchange (order 40 days) between the deep eddy water and the rest of the coastal ocean.
-In this model, the region where the eddy resides has velocities equal or greater than other parts of the shelf, and water has an approximate residence time of less than 20 days.
-The eddy water in the model does develop a distinct $\theta$--$S$ signature, but not nearly as strong as observed.
-It also has a front with the offshore water, but the front is much wider and more gradual than that observed here.
+In contrast, numerical simulations of this region reported by \cite{sahuetal22} using a NEMO 36th-degree regional model  indicate that there is more rapid exchange (order 40 days) between the deep \Eddy\ water and the rest of the coastal ocean.
+In this model, the region where the \Eddy\ resides has velocities equal or greater than other parts of the shelf, and water has an approximate residence time of less than 20 days.
+The \Eddy\ water in the model does develop a distinct $\theta$--$S$ signature, but not nearly as strong as observed.
+It also has a front with the offshore water, but the front is more gradual than that observed here.
 
-Overall, it would be an improvement to our understanding of the eddy if we could sample the shelf more persistently.
-The eddy was already well-formed by the June La Perouse cruise, and seems to evolve slowly during that time.
-Capturing its formation, presumably earlier in the spring, and its evolution through the year would be valuable in undertanding retention and exchange on this productive part of the shelf.
+Overall, it would be an improvement to our understanding of the \Eddy\ if we could sample the shelf more persistently.
+The \Eddy\ was already well-formed by the June La Perouse cruise, and seems to evolve slowly during that time.
+Capturing its formation, presumably earlier in the spring, and its evolution through the year would be valuable in understanding retention and exchange on this productive part of the shelf.
 
@@
 
-The (apparent) displacement of shelf water from La Perouse Bank is a dramatic departure from geostrophically balanced isobath-following flow..
+The (apparent) displacement of shelf water from La Perouse Bank is a dramatic departure from geostrophically balanced isobath-following flow.
 Eddies have been known to separate from irregular coastal topography, both at the surface \cite{barthetal00}, and deeper in the water column \cite{pellandetal13}.
 It has been recognized that instabilities lead to exchange between the open ocean and the shelf at this location \cite{ikedaemery84, ikedaemery84}.
-However, observations of the wholesale replacement of shelf water by a new water mass from offshore, is rare.
+However, observations of the wholesale replacement of shelf water by a new water mass from offshore are rare.
 In our situation it is clear that water from as deep as 150 m is separating from the shelf and moving offshore (\fref{fig:SpiceO2264}, \fref{fig:IsopycnalsDepth}).
 
-Satellite imagery makes it clear that there is often an exchange between coastal and deepwater along the Vancouver Island shelf (\fref{fig:SSTLateAug}).
+Satellite imagery makes it clear that there is often an exchange between coastal and deep water along the Vancouver Island shelf (\fref{fig:SSTLateAug}).
 Most years there are three of four large excursions from the shelf into the open ocean, many of them over 100 km long.
-This lengthscale is somewhat larger than 60 km inferred for this region by \citeA{ikedaetal84} using a four-layer instability analysis.
-It is possible that there is even some spatial locking of these features, with a persistent separation at the north tip of the Island, and a strong tendency for one at 49.5 N.
+This length scale is longer than the 60 km inferred for this region by \citeA{ikedaetal84} using a four-layer instability analysis.
+It is possible that there is spatial locking of these features, with a persistent separation at the north tip of the Island, and a strong tendency for one at 49.5 N.
 Our study area also has evidence of a separation in most of the years, with 2011 being the only clear exception.

### 2023-01-05 PW13Paper.tex
FIX: new version to SW

@@
 
-\authors{Jody M. Klymak\affil{1,2}, Susan E. Allen\affil{3,4}, Stephanie N. Waterman\affil{3}}
+\authors{Jody M.\ Klymak\affil{1,2}, Susan E.\ Allen\affil{3,4}, Stephanie Waterman\affil{3}}
 
@@
 \item Offshore water is drawn onto the shelf and forms a sharp semi-persistent front with the Juan de Fuca Eddy.
-\item The \Eddy\ shows evidence of long residence times, and little evidence of deep-water origin.
+\item The Eddy shows evidence of long residence times, and little evidence of deep-water origin.
 \end{keypoints}
@@
 The observations indicate that the water in this Eddy is a mixture of offshore water and water from a buoyant coastal current.
-The water in the  in the Eddy is well-mixed along a mixing line in $\theta$--$S$ space, though it retains stratification, and is either rapidly mixed (high $\kappa$), or has a long residence time (large $\tau$), such that $(\kappa \tau)^{1/2} \sim 50$ ---  $100\ \mathrm{m}$.
-There is a sharp temperature-salinity front on the offshore side of of this well-mixed water, with the front no more than 1-km wide, and no sign of instabilities.
-Previous work had hypothesized the role of a spur canyon incising the shelf in feeding water to the Eddy; little evidence of this was found during the observations, however, these were collected during a relaxation and reversal of upwelling winds, which may have slowed the flow of water into the Eddy.
-Tidally resolved observations in the spur canyon also show very strong hydraulic flows in the cross-canyon direction accompanied by very high localized mixing rates in 50-m hydraulic jumps possibly exceeding $\kappa \sim 1 \ \mathrm{m^2s^{-1}}$.
+This water is well-mixed along a mixing line in temperature-salinity space, though it retains stratification, and is either rapidly mixed (high mixing rate $\kappa$), or has a long residence time (large $\tau$), such that the mixing length $(\kappa \tau)^{1/2} \sim 50$ ---  $100\ \mathrm{m}$.
+There is a sharp temperature-salinity front on the offshore side of of this well-mixed water.
+The front is no more than 1-km wide, and has no sign of instabilities.
+Previous work had hypothesized the role of a spur canyon incising the shelf in feeding water to the Eddy; little evidence of this was found in the observations, however, these were collected during a relaxation and reversal of upwelling winds, which may have slowed the flow of water into the Eddy.
+Tidally resolved observations in the spur canyon also show very strong hydraulic flows in the cross-canyon direction accompanied by very high localized mixing rates in 50-m hydraulic jumps, possibly exceeding $\kappa \sim 1 \ \mathrm{m^2s^{-1}}$.
 
@@
 The shelf current was observed to separate from the shelf in the lee of a bank, and is replaced by the distinct offshore water mass that forms the offshore front with the Eddy.
-The separation event can be seen in sea-surface temperatures in satellite images as tongue of water that is ejected offshore, possibly with the relaxation of upwelling, with the flow heading directly offshore.
-The cause of this offshore separation event is not determined, but likely involves instability between the shelfbreak current and the California Undercurrent, and possibly is catalyzed by separation from the abrupt underwater bank at the poleward end of the study region.
+The separation event can be seen in sea-surface temperatures in satellite images as a tongue of water that is ejected offshore.
+The cause of this offshore separation event is not determined, but possibly is catalyzed by separation from the abrupt underwater bank at the poleward end of the study region.
 
@@
 
-The southern Vancouver Island continental shelf is very productive due to high nutrient input from the Strait of Juan de Fuca and Salish Sea estuarine system, and substantial cross-shelf transport due to the complicated topography.
-The estuarine water flows poleward, hugging the coast as the Vancouver Island Coastal Current, and in the summer shelf water flows equatorward.
+The southern Vancouver Island continental shelf is very biologically productive due to high nutrient input from the Strait of Juan de Fuca and Salish Sea estuarine system and substantial cross-shelf transport due to the complicated topography.
+The estuarine water flows poleward, hugging the coast as the Vancouver Island Coastal Current and, in the summer, shelf water flows equatorward.
 Trapped between these two currents is the Juan de Fuca Eddy, a stagnant region of water that occupies the shelf where it abruptly widens just north of the Strait of Juan de Fuca.
@@
 Here we present intensive sampling of the Juan de Fuca Eddy region.
-The observations show the Eddy is low in oxygen, and we find has undergone substantial vertical and lateral mixing.
+The observations show that water in the Eddy is low in oxygen, and we find has undergone substantial vertical and lateral mixing.
 This indicates that the waters have a surprisingly long residence time for a coastal feature, and that this region extends the depth of the water column.
-In contrast to previous literature we find that the water likely has low oxygen from being used for respiration rather than being pulled from deep in the California Undercurrent.
+In contrast to previous literature we find that the water likely has low oxygen from being used for respiration rather than being pulled from low-oxygen water in the California Undercurrent.
 
-The sampling also shows a remarkable flow separation of the equatorward shelf current  The current is shown to detach and is pushed offshore.
-Such events are readily seen in satellite imagery, but the results here indicate that the separation extends the depth of the water column on the shelf, and that this separation may be partially driven by the local bathymetry.
+The observations also show a remarkable flow separation of the equatorward shelf current.
+The current is shown to detach and is pushed offshore.
+Such events are readily seen in satellite imagery, but our observations indicate that the separation extends the depth of the water column on the shelf, and that this separation may be partially driven by the local bathymetry.
 The separation is a very strong cross-shelf exchange event, and likely transports substantial nutrient-rich coastal water offshore to drive productivity in the deeper ocean adjacent to the continental slope.
@@
 
-Cross-shelf exchange is important to the health and productivity of continental shelf regions, where offshore oxygenated water is exchanged with nutrient-rich, but oxygen-depleted nearshore water.
+Cross-shelf exchange is important to the health and productivity of continental shelf regions, allowing for offshore oxygenated water to be exchanged with nutrient-rich, but oxygen-depleted nearshore water.
 Cross-shelf transport usually requires ageostrophic flow, since geostrophically balanced flow will tend to follow topographic contours, often providing a barrier to lateral exchange \cite{brink16}.
 Mechanisms for cross-shelf exchange include internal waves and instabilities in shelf-break fronts.
-Most-dramatically, however, topography can drive cross-slope exchanges, often seen most clearly in surface imagery, but occasionally observed \emph{in-situ} \cite{barthetal00}.
+However, these steady mechanisms are often smaller in magnitude to intermittent three-dimensional exchange, driven by eddying, and often catalyzed by topographic irregularity \cite{barthetal00}.
 
-Here we present detailed \emph{in-situ} observations from the southern Vancouver Island shelf, collected in summer 2013, and sampled with a rapid profiling vehicle equipped with a CTD and oxygen sensor, supplemented by traditional hydrographic surveys.
-The shelf has complicated bathymetry \fref{fig:LocMapBoth}, with a somewhat typical shelf poleward of our study site, but then widening in the lee of La Perouse Bank, and then hitting the Juan de Fuca canyon, equatorward.
-Water from the strait of Juan de Fuca flows poleward as a buoyant current that hugs the coast \cite{thomsonetal89, hickeyetal91}, while shelf water flows equatorward in the summer under upwelling conditions, both forced by local winds and via tele-connections with the long, homogenous shelves equatorward off Washington and Oregon \cite{hickeyetal91,thomsonkrassovski15,engidaetal16}.
+Here we present detailed \emph{in-situ} observations from the southern Vancouver Island shelf, collected in summer 2013, and sampled by a rapid profiling vehicle equipped with a CTD and oxygen sensor, supplemented by traditional hydrographic surveys bracketing the high resolution observations by a month.
+The shelf has complicated bathymetry (\fref{fig:LocMapBoth}), with a simple shelf geometry poleward of our study site, but widening in the lee of La Perouse Bank, and incised by the Juan de Fuca canyon, equatorward.
+Water from the Strait of Juan de Fuca flows poleward as a buoyant current that hugs the coast \cite{thomsonetal89, hickeyetal91}, while shelf water flows equatorward in the summer under upwelling conditions, both forced by local winds and via tele-connections with the long, homogenous shelves equatorward off Washington and Oregon \cite{hickeyetal91,thomsonkrassovski15,engidaetal16}.
 Trapped between these two currents is a region of relatively homogenous water that has been termed the Juan de Fuca Eddy \cite{freelanddenman82,freelandmcintosh89,foremanetal07,macfadyenhickey10}, denoted ``\Eddy'' below.
 
-A goal of our study was to understand how the \Eddy persists, and how it exchanges water properties with offshore water.
-Low oxygen has previously been found in the \Eddy, causing it to be inferred that water in the \Eddy\ is upwelled from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
-This inference was based on the oxygen being conservative, and hence having to come from close to the oxygen minimum zone found further offshore \cite{mackasetal87} and lead to hypotheses that perhaps the \Eddy\ was fed by waters being drawn up a spur canyon (called the Spur Canyon) via ageostrophic transport from offshore to onshore due to the low pressure in the \Eddy\ center \cite{weaverhsieh87}.
-Below, we will argue that there is not substantial evidence of such transport, and that oxygen is likely low because of consumption due to respiration.
+A goal of our study was to understand how the \Eddy persists and how it exchanges water properties with offshore water.
+Low oxygen has previously been found in the \Eddy, causing previous work to infer that water in the \Eddy\ is upwelled from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
+This inference was based on the assumption that oxygen is conservative, and hence water in the eddy had to come from  the oxygen minimum zone found further offshore \cite{mackasetal87}.
+This lead to hypotheses that perhaps the \Eddy\ was fed by waters being drawn up a spur canyon (called the Spur Canyon) via ageostrophic transport from offshore to onshore due to low pressure in the \Eddy\ center \cite{weaverhsieh87}.
+Below, we will argue that there is no evidence of such transport and that oxygen is likely low because of local consumption due to respiration.
 
 A second goal was to better understand cross-shore exchange with offshore water.
-While the typical schematic of shelf exchange is that there are Ekman layers near the boundaries, there is also evidence of filaments and wholesale separation of shelf currents into the deep ocean, both at this site, and at others.
-Such events are important as they tend to transport nutrient rich shelf water out into the open ocean, fueling offshore ecosystems.
+Cross-shore transport is supplied by Ekman layers in simple geometries under wind forcing.
+However, in many locations there is also evidence of filaments and wholesale separation of shelf currents into the deep ocean.
 Along the Vancouver Island shelf, large filaments or instabilities of the coastal current have been inferred using satellite images \cite{ikedaemery84,thomsongower98}.
 Direct observations of such filaments have been made further to the south off Cape Blanco \cite{barthetal00}, and other coasts \cite[e.g.]{relvasbarton05}.
-Net cross-shore transport of nutrients and chlorophyll have been found in hydrographic surveys off Vancouver Island \cite{mackasyelland99}, and associated with mesoscale features both in geostrophic velocities, and in satellite observations.
-The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics \cite{dalebarth01} leading to along-shore trapping of coastal waves,  wind stress curl variations \cite{castelaobarth07}, induced relative vorticity due to stretching of parcels that overshoot their initial isobaths due to a sudden change in downstream bathymetry, and the non-linear breaking of large-scale meanders due to baroclinic instability between the wind-driven current and the California Undercurrent \cite{ikedaetal84, batteen97}.
+Net cross-shore transport of nutrients and chlorophyll have been found in hydrographic surveys off Vancouver Island \cite{mackasyelland99}, and associated with mesoscale features in geostrophic velocities and in satellite observations.
+The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics leading to along-shore trapping of coastal waves \cite{dalebarth01},  wind stress curl variations \cite{castelaobarth07}, induced relative vorticity due to stretching of parcels that overshoot their initial isobaths due to a sudden change in downstream bathymetry \citep{dasaro88}, and the non-linear breaking of large-scale meanders due to baroclinic instability between the wind-driven current and the California Undercurrent \cite{ikedaetal84, batteen97}.
 
@@
 
-Observations from three summer cruises in 2013 are discussed here.
-Two were hydrographic surveys carried out on the \emph{CCGS Tully}, one in June, and the other in late September.
-The third was a finescale survey made in the Vancouver Island shelf between 21 and 30 August 2013, aboard the \emph{R/V Falkor}.
-The study site was the southern portion of the Vancouver Island Shelf (\fref{fig:LocMapBoth}a), a particularly complicated region due to the bathymetry and varied forcing.
-
 \begin{figure*}[htbp]
@@
 The dashed box indicates the approximate limits of the study area.
-b) The study area with hydrographic casts from the La Perouse cruises along the LB and LC Lines are green dots, hydrographic casts during the Falkor cruise indicated in yellow dots, and Moving Vessel Profiler casts as red dots.
-The coordinate system used for this paper is shown with alternating grey and white bands at 10-km intervals.}
+b) The study area with hydrographic casts from the La Perouse cruises along the LB and LC Lines are green dots, hydrographic casts during the Falkor cruise indicated in yellow dots, and finescale Moving Vessel Profiler casts as red dots.
+The coordinate system used for this paper is shown with alternating grey and white bands at 10-km intervals in the along- and cross-shore directions.}
   \label{fig:LocMapBoth}
@@
 
+The study site was the southern portion of the Vancouver Island Shelf (\fref{fig:LocMapBoth}a), a particularly complicated region due to the bathymetry and varied forcing.
 At the south end of the study site is the Juan de Fuca canyon, which feeds dense water into Juan de Fuca Strait.
-This dense water is mixed with fresh water from the Fraser River at the sills and archipelagos further inland, and fluxes out the Strait again, where it turns poleward along Vancouver Island to form the Vancouver Island Coastal Current.
+This dense water is mixed with fresh water from the Fraser River at the sills and archipelagos further inland and fluxes out the Strait again, where it turns poleward along Vancouver Island to form the Vancouver Island Coastal Current.
 The Juan de Fuca Canyon has a notable spur canyon (Spur Canyon), that incises the shelf towards the north into the study site (\fref{fig:LocMapBoth}b).
@@
 Notable is La Perouse bank, running along the shelf from the north, and moving isobaths offshore until it abruptly ends at around 48.5 N.
-The shelf break is characterized by a series of submarine canyons, in particular Nitnat Canyon and Barkley Canyon.
+The shelf break has a series of submarine canyons, in particular Nitnat Canyon and Barkley Canyon.
 
-We consider routine hydrographic surveys along the LB and LC lines, collected aboard the \emph{CCGS Tully} from 2013-05-30 to 2013-05-31  (``June''), and 2013-09-07 to 2013-09-09 (``September'').
-These lines span the depths from about 50 m to deep offshore, with casts every 7.5 km across shelf.
+We present routine La Perouse hydrographic surveys along the LB and LC lines, collected aboard the \emph{CCGS Tully} from 2013-05-30 to 2013-05-31  (``June''), and 2013-09-07 to 2013-09-09 (``September'').
+These lines span the 50 m isobath to deep offshore, with casts every $7.5\ \mathrm{km}$ across shelf.
 The data comes from a lowered Seabird 9-11 CTD, with an SBE 43 oxygen sensor.
-Oxygen data have been checked against bottle casts, and the CTD corrected for sensor offsets and thermal lags.
+Oxygen data have been corrected against bottle casts, and the CTD corrected for sensor offsets and thermal lags.
 
-We focus on fine-scale surveys carried out from 2013-08-21 to 2013-08-30.
+Below, we focus on finescale surveys carried out from 2013-08-21 to 2013-08-30.
 Data were collected from the \emph{R/V Falkor} with an ODIM Brooke-Ocean Moving Vessel Profiler (MVP).
 The MVP was equipped with an AML Oceanographic CTD, and a Rinko Oxygen sensor with a 7-s response time foil.
-The MVP profiles to depths of 200 m or to within 5 m of the seafloor, whichever is shallower, and drops at a speed of $3\ \mathrm{m\,s^{-1}}$.
-Data collection took place while the ship cruised at speeds between 5 and 8 kts, usually at around 6 kts to enable fine horizontal spacing of the casts, with typical spacing less than 800 m.
+The MVP profiled to depths of 200 m or to within 5 m of the seafloor, whichever was shallower, and dropped at a speed of approximately $3\ \mathrm{m\,s^{-1}}$.
+Data collection took place while the ship cruised at speeds between 5 and 8 kts, usually at around 6 kts, to enable fine horizontal spacing of the casts, with typical spacing  of 800 m in deep water, and less in water shallower than 200 m.
 Data is reported for the downcast, which mostly follows a vertical path.
-The rapid speed of the profiling makes the oxygen measurements somewhat coarse, and probably biased due to the phase lag of the sensor, so we treat these largely qualitatively in this paper.
+The rapid speed of the profiling makes the oxygen measurements somewhat coarse, and probably biased due to the phase lag of the sensor, so we treat these qualitatively in this paper.
 
-Unfortunately, neither vessel had an operational acoustic Doppler profiler during the cruises with which to make current measurements.
+Unfortunately, neither vessel had an operational acoustic Doppler profiler during the cruises with which to make water velocity measurements.
 
 Winds during the cruises were typical for the west coast of Vancouver Island, with upwelling favorable winds during July and early August (\fref{fig:LaPeWind}).
-During the finescale cruise, and for the week previous, the winds were intermittently downwelling favorable.
-Note that this locale is strongly affected by coastally trapped waves from further south, so doming of near-bottom isopycnals often persists despite local wind forcing \cite{thomsonkrassovski15, engidaetal16}, and takes a finite amount of time to spin down.
+During the finescale survey, and for the week previous, the winds were intermittently downwelling favorable.
+Note that this locale is strongly affected by coastally trapped waves from further south, so doming of near-bottom isopycnals often persists despite local wind forcing, and takes a finite amount of time to spin down \cite{thomsonkrassovski15, engidaetal16}.
 
@@
 The vertical direction is along-shelf (chosen as a heading of 315 N), so vectors pointing straight down represent upwelling winds.
-.
-The wind components has been low-pass filtered to one-day averages.
-The timing of the cruises discussed in the paper are shown as colored bands.
+The wind components have been low-pass filtered to one-day averages.
+The timing of the surveys discussed in the paper are shown as colored bands.
       \label{fig:LaPeWind} }
@@
 
-\subsection{Early and late summer broad-scale surveys}
+\subsection{Early and late summer hydrographic surveys}
 
-Hydrographic sections along the LB and LC lines highlight summer conditions on the Southern Vancouver Island Shelf, and indicate some of the features we are focusing on in this paper (\Fref{fig:LaPerouse2013Ctd}).
+Hydrographic sections along the LB and LC lines highlight summer conditions on the Southern Vancouver Island Shelf and indicate some of the features we are focusing on in this paper (\Fref{fig:LaPerouse2013Ctd}).
 During both cruises, and at both lines, there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m depth offshore to shallower than 85 m over the shelf.
-This dense water tends to be low in oxygen and cool.
-Further onshore the Vancouver Island Coastal Current hugs the coast where isopycnals tilt back down, and the water is warmer than offshore.
+This upwelled water tends to be low in oxygen and cool.
+Further onshore the Vancouver Island Coastal Current hugs the coast, where isopycnals tilt back down, and the water is warmer than offshore.
 
@@
 
-The Juan de Fuca Eddy region is along the LB Line (\fref{fig:LaPerouse2013Ctd}, bottom three rows), mostly inshore of 0 km.
+The Juan de Fuca Eddy region is observed along the LB Line (\fref{fig:LaPerouse2013Ctd}, bottom three rows), mostly inshore of 0 km.
 This water is cooler and lower in oxygen than along the same isopycnal further offshore.
-Also note that during the June cruise the water in this region was cooler and more oxygenated than later in the summer during the September cruise.
-Near the surface, the isopycnals are domed, but the low-oxygen anomaly extends as high as the thin surface mixed layer.
+Near the surface, the isopycnals are domed, and the low-oxygen anomaly extends as high as the thin surface mixed layer.
 
-The water upstream of the \Eddy\ region, along the LC line, is less mixed than the \Eddy\ water.
-It does not show as much drawdown of oxygen, and is warmer, at least offshore of La Perouse bank (\fref{fig:LaPerouse2013Ctd}, top three rows).
-Water does not appear to make it over La Perouse bank into the basin onshore, or if it does, it does so intermittently.
-There appears to be some mixing of the deeper and shallow water near $26.4\ \mathrm{kg\,m^{-3}}$, with the water along this isopycnal becoming cooler and lower in oxygen where it intersects the shelf, consistent with enhanced mixing near the shelf
+The water upstream of the \Eddy\ region, along the LC line, is less mixed than the \Eddy\ water (\fref{fig:LaPerouse2013Ctd}, top three rows).
+It does not show as much drawdown of oxygen, and is warmer, at least offshore of La Perouse bank ($X \approx 5\ \mathrm{km}$).
+Based on these properties, offshore water does not appear to make it over La Perouse bank into the basin onshore, or if it does, it does so intermittently.
+Note that the water along the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal becomes cooler and lower in oxygen where it intersects the shelf, consistent with both enhanced mixing near the shelf, and with drawdown of oxygen by respiration.
 
-The $\theta$--$S$ properties show many of the features demonstrated above (\fref{fig:LaPerouse2013TS}a).
-Below $26.6\ \mathrm{kg\,m^{-3}}$, the deep water masses found offshore are largely homogenous.
-Shallower, there are distinct differences between warm and salty offshore compared to water on the shelf at the same densities.
-In this paper we use a spice anomaly relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}) representing an approximate mixing line for water in the \Eddy.
-It passes through the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$.
+The contrast in onshore and offshore waters can be clearly traced in temperature-salinity ($\theta$--$S$) anomalies along isopycnals (\fref{fig:LaPerouse2013TS}a).
+Denser than $26.6\ \mathrm{kg\,m^{-3}}$, the deep water masses found offshore are largely homogenous.
+In the lighter water masses, there are distinct differences between warm and salty offshore compared to water on the shelf at the same densities.
+
+We use a spice anomaly relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}) representing an approximate mixing line for water in the \Eddy, and passsing through the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$.
 Spice anomaly for a given water sample at density $\sigma_{\theta}$ is given by $\gamma = \alpha \left(T - T_0\left(\sigma_{\theta}\right)\right) + \beta \left(S - S_0\left(\sigma_{\theta}\right)\right)$ where $T_0\left(\sigma_{\theta}\right)$, where $S_0\left(\sigma_{\theta}\right)$ are the temperature and salinity along the mixing line at the same density as the water sample, with the sign convention that positive is warmer and saltier than data along the mixing line.
@@
 
-Offshore water tends to be high in absolute spice anomaly (\fref{fig:LaPerouse2013Ctd}, middle rows), with a positive-anomaly layer (red) centered at $26.4\ \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and below.
+Using this measure of spice anomally, offshore water tends to have high absolute values (\fref{fig:LaPerouse2013Ctd}, middle rows), with a positive-anomaly layer (red) centered at $26.4\ \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and below.
 On the shelf ($-10\, \mathrm{km} < X < 30\, \mathrm{km}$), the distinct $\theta$--$S$ masses are attenuated and much closer to the defined mixing line than the offshore water (closer to white in color).
-Hugging the coast ($X > 30\, \mathrm{km}$), is the Vancouver Island Coastal Current, with a negative spice anomaly in June, and a positive anomaly in summer, consistent with seasonal warming.
 
-The June and September cruises have quite similar water properties, particularly in the offshore waters.
-Water has upwelled from deeper depths in September, but has most of the same water mass characteristics as earlier in the year.
-The water on the shelf is somewhat warmer than earlier in the year, and by September has started to fall off the mixing line.
+There are temporal changes over the summer.
+Away from the surface, water has upwelled from deeper depths by September, but the offshore water maintains the same water mass characteristics through the summer.
+Hugging the coast ($X > 30\, \mathrm{km}$), the Vancouver Island Coastal Current warms during the summer, and the spice anomaly goes from negative to postive.
+In the \Eddy, the water stays near the mixing line, but is cooler in the spring (\fref{fig:LaPerouse2013Ctd}, LB Line, left-hand column: slightly negative spice anomally) and warms during the summer (\fref{fig:LaPerouse2013Ctd}, LB Line, right-hand column: slightly positive spice anomally).
 
-In terms of oxygen, both cruises have a similar dichotomy between the onshelf and offshelf water, with water found in the eddy tracking at least $100\ \mathrm{\mu mol\, kg^{-1}}$ lower on the shelf than offshelf (\fref{fig:LaPerouse2013Ctd}, \fref{fig:LaPerouse2013TS}b).
+In terms of oxygen, both surveys have a similar dichotomy between the \Eddy and offshelf water.
+Water found in the \Eddy is at least $100\ \mathrm{\mu mol\, kg^{-1}}$ lower on the shelf than offshelf (\fref{fig:LaPerouse2013Ctd}, \fref{fig:LaPerouse2013TS}b).
 The very deepest water ($S\approx 33.9\ \mathrm{psu}$) on the shelf shows a further $50\ \mathrm{\mu mol\, kg^{-1}}$ drop between June and September along the LB line.
 
-\subsection{Fine-scale surveys}
+\subsection{Finescale surveys}
 
@@
 
-As shown in \fref{fig:LocMapBoth},  we conducted a large spatial survey of the shelf using the Moving Vessel Profiler, collecting over 2000 CTD casts in a relatively synoptic manner.
-This amount of data allows a reasonable census of the water masses that motivated our choice of the mixing line shown in \fref{fig:LaPerouse2013TS}.
-Binning in temperature and salinity, we can look at sample frequency of these observations \fref{fig:TSdensSpice}, and we see that there are distinct water masses with few observations between them.
-Looking along the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal, we see that there is a large population along the mixing line.
-At the opposite extreme, we see that there is a large population with a positive spice anomaly of approximately $0.2\ \mathrm{kg\,m^{-3}}$, representing the offshore water.
-Between these two end points there is a third water mass, somewhat indistinct from the offshore water mass, but very distinct from the water mass along the mixing line.
-These three water masses exist on deeper and shallower isopycnals as well, but become hard to distinguish from $\theta$--$S$ properties alone as the properties converge near $26.1\,\mathrm{kg\,m^{-3}}$.
+The finescale survey covered most of the \Eddy region, with an emphasis on the offshore edge near the shelfbreak front (\fref{fig:LocMapBoth}).
+For water denser than $26\ \mathrm{kg\,m^{-3}}$, there are three distinct water masses found on the shelf (\fref{fig:TSdensSpice}).
+The first is the offshore water, which tends to be warmer, and hence have a positive spice anomally ($\gamma \approx 0.2\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$).
+The second is water in the \Eddy, which during this survey was found along the straight mixing line in $\theta-S$ space (\fref{fig:TSdensSpice}, $\gamma \approx 0\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$).
+Between these two water masses there is a less populous mass (\fref{fig:TSdensSpice}, $\gamma \approx 0.1\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$) that we demonstrate below is found on the shelf poleward of the \Eddy.
 
@@
      \includegraphics[width=3.5in]{TSdensSpice.pdf}
-    \caption{Sample density of salinity and potential density data from the cruise, with a logarithmic color scale.
+    \caption{Binned sample density of salinity and potential density data from the cruise, with a logarithmic color scale.
 Grey contours are potential density relative to the surface, every $0.1\ \mathrm{kg\,m^{-3}}$, and the magenta contour is the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.
-Spice anomaly, $\gamma$, is defined as temperature anomaly scaled by the thermal expansion co-efficient, $\alpha$, along isopycnals from the mixing line found in the well-mixed region in the \Eddy\ (white line), in $\mathrm{kg\,m^{-3}}$.
+Colored contours are spice anomaly, $\gamma$, relative to the straight line labeled $\gamma=0.0$.
     \label{fig:TSdensSpice}}
@@
 
-In synthetic cross sections of spice anomaly, the \Eddy\ is clearly identifiable as water that is along the straight mixing line, and thus has low-spice anomaly ($\gamma \approx 0\ \mathrm{kg\,m^{-3}}$, \fref{fig:CrossSectionsSpice}c, d).
-This signature of well-mixed water extends from the seafloor to approximately 30 m depth, onshore of $x \approx 0\ \mathrm{km}$.
+In synthetic cross sections of spice anomaly, the \Eddy\ is clearly identifiable as having low-spice anomaly ($\gamma \approx 0\ \mathrm{kg\,m^{-3}}$, \fref{fig:CrossSectionsSpice}c, d).
+This signature of well-mixed water extends from the seafloor to approximately 30 m depth, onshore of $X \approx 0\ \mathrm{km}$.
 Despite lying along a mixing line, the \Eddy\ water is still stratified.
-This partially-mixed water contrasts with water directly offshore, on the other side of a sharp $\theta-S$ compensated front, that is warmer and saltier than the mixed water ($\gamma > 0.2\ \mathrm{kg\,m^{-3}}$) deeper than $26\,\mathrm{kg\,m^{-3}}$, and colder and fresher shallower.
+The high-spice water ($\gamma \approx 0.2\ \mathrm{kg\,m^{-3}}$) is found offshore of the \Eddy water, on the other side of a sharp $\theta-S$ compensated front.
 This front is even sharper in individual sections than in these composite sections (see \fref{sec:frontsurvey}).
@@
 
-\subsection{Spur Canyon}
+\subsubsection{Spur Canyon}
 

### 2023-01-18 PW13Paper.tex
After more comments from Susan and STephanie

@@
 This water is well-mixed along a mixing line in temperature-salinity space, though it retains stratification, and is either rapidly mixed (high mixing rate $\kappa$), or has a long residence time (large $\tau$), such that the mixing length $(\kappa \tau)^{1/2} \sim 50$ ---  $100\ \mathrm{m}$.
-There is a sharp temperature-salinity front on the offshore side of of this well-mixed water.
+There is a sharp temperature-salinity front on the offshore side of this well-mixed water.
 The front is no more than 1-km wide, and has no sign of instabilities.
@@
 Upstream of the Eddy, there is an along-shelf current flowing equatorward with partially mixed water properties compared to the well-mixed water in the Eddy.
-The shelf current was observed to separate from the shelf in the lee of a bank, and is replaced by the distinct offshore water mass that forms the offshore front with the Eddy.
+The shelf current was observed to separate from the shelf in the lee of a bank, and was replaced by a distinct offshore water mass.
 The separation event can be seen in sea-surface temperatures in satellite images as a tongue of water that is ejected offshore.
-The cause of this offshore separation event is not determined, but possibly is catalyzed by separation from the abrupt underwater bank at the poleward end of the study region.
+The cause of this offshore separation event is not determined, but possibly is catalyzed by separation of the current from the abrupt underwater bank at the poleward end of the study region.
 
@@
 Here we present intensive sampling of the Juan de Fuca Eddy region.
-The observations show that water in the Eddy is low in oxygen, and we find has undergone substantial vertical and lateral mixing.
-This indicates that the waters have a surprisingly long residence time for a coastal feature, and that this region extends the depth of the water column.
-In contrast to previous literature we find that the water likely has low oxygen from being used for respiration rather than being pulled from low-oxygen water in the California Undercurrent.
+The observations show that below the surface mixed layer, the water in the Eddy is low in oxygen, and has undergone substantial vertical and lateral mixing.
+In contrast to previous literature we find that the low oxygen in the eddy is likely because of respiration rather than being pulled from low-oxygen water in the California Undercurrent.
 
 The observations also show a remarkable flow separation of the equatorward shelf current.
-The current is shown to detach and is pushed offshore.
+The current is seen to detach and is pushed offshore.
 Such events are readily seen in satellite imagery, but our observations indicate that the separation extends the depth of the water column on the shelf, and that this separation may be partially driven by the local bathymetry.
@@
 
-Cross-shelf exchange is important to the health and productivity of continental shelf regions, allowing for offshore oxygenated water to be exchanged with nutrient-rich, but oxygen-depleted nearshore water.
+Cross-shelf exchange is important to the health and productivity of continental shelf regions, allowing for offshore oxygenated water to be exchanged with nutrient-rich, but oxygen-depleted, nearshore water.
 Cross-shelf transport usually requires ageostrophic flow, since geostrophically balanced flow will tend to follow topographic contours, often providing a barrier to lateral exchange \cite{brink16}.
-Mechanisms for cross-shelf exchange include internal waves and instabilities in shelf-break fronts.
-However, these steady mechanisms are often smaller in magnitude to intermittent three-dimensional exchange, driven by eddying, and often catalyzed by topographic irregularity \cite{barthetal00}.
+Mechanisms of cross-shelf exchange include internal waves and instabilities in shelf-break fronts.
+However, these mechanisms are often smaller in magnitude to intermittent three-dimensional exchange, driven by eddying, and often catalyzed by topographic irregularity \cite{barthetal00}.
 
@@
 
-A goal of our study was to understand how the \Eddy persists and how it exchanges water properties with offshore water.
-Low oxygen has previously been found in the \Eddy, causing previous work to infer that water in the \Eddy\ is upwelled from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
+A goal of our study was to understand how the \Eddy\ persists and how it exchanges water properties with offshore water.
+We focus on water deeper than $50 \mathrm{m}$, below the summer mixed layer, because this water is trackable with water mass properties, whereas the \Eddy\ is often studied from the point of view of the surface circulation \citep[e.g.][]{macfadyenhickey10}.
+Past studies have found low oxygen concentrations in the \Eddy, and inferred that water is upwelled from the California Undercurrent from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
 This inference was based on the assumption that oxygen is conservative, and hence water in the eddy had to come from  the oxygen minimum zone found further offshore \cite{mackasetal87}.
@@
 
-A second goal was to better understand cross-shore exchange with offshore water.
+A second goal of our study was to better understand cross-shore exchange between shelf and offshore water.
 Cross-shore transport is supplied by Ekman layers in simple geometries under wind forcing.
@@
 Along the Vancouver Island shelf, large filaments or instabilities of the coastal current have been inferred using satellite images \cite{ikedaemery84,thomsongower98}.
-Direct observations of such filaments have been made further to the south off Cape Blanco \cite{barthetal00}, and other coasts \cite[e.g.]{relvasbarton05}.
+Direct observations of such filaments have been made further to the south off Cape Blanco \cite{barthetal00}, and off other coasts \cite{relvasbarton05}.
 Net cross-shore transport of nutrients and chlorophyll have been found in hydrographic surveys off Vancouver Island \cite{mackasyelland99}, and associated with mesoscale features in geostrophic velocities and in satellite observations.
@@
 
-In this paper we present observations from 2013 (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes, with a focus on the offshore side of the \Eddy\ region (\fref{sec:Observations}) where we consider the age of the \Eddy, the steadiness of the front separating the \Eddy\ and the offshore water, properties along the spur canyon that was believed to feed the \Eddy, and finally describe a large offshore tongue of the shelf-break current and its intrusion into the interior.
+In this paper we present observations of water masses on the Southern Vancouver Island Shelf collected in 2013 (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes, with a focus on the offshore side of the \Eddy\ region (\fref{sec:Observations}) where we consider the age of the \Eddy, the steadiness of the front separating the \Eddy\ and the offshore water, properties along the spur canyon that was believed to feed the \Eddy, and a large offshore tongue of the shelf-break current and its intrusion into the interior.
 We discuss the origin of the \Eddy\ and the implications of the separating jet (\fref{sec:Summary}).
@@
   \caption{a) Study site on the Vancouver Island Shelf.
-The blue arrow indicates the direction of the Vancouver Island Coastal Current.
-The pink arrow indicates southward flow of the coastal upwelling current.
+The blue arrow indicates the direction of the Vancouver Island Coastal Current, and the pink arrow indicates southward flow of the coastal upwelling current.
 The dashed box indicates the approximate limits of the study area.
-b) The study area with hydrographic casts from the La Perouse cruises along the LB and LC Lines are green dots, hydrographic casts during the Falkor cruise indicated in yellow dots, and finescale Moving Vessel Profiler casts as red dots.
+b) The study area with hydrographic casts from the La Perouse cruises along the LB and LC Lines (green dots), hydrographic casts during the Falkor cruise (yellow diamonds), and finescale Moving Vessel Profiler casts (red dots).
 The coordinate system used for this paper is shown with alternating grey and white bands at 10-km intervals in the along- and cross-shore directions.}
@@
 The rest of the shelf is punctuated by a series of banks and shallow basins.
-Notable is La Perouse bank, running along the shelf from the north, and moving isobaths offshore until it abruptly ends at around 48.5 N.
-The shelf break has a series of submarine canyons, in particular Nitnat Canyon and Barkley Canyon.
+La Perouse bank  separates the outer shelf from a deeper inner basin and forms the poleward boundary off where the shelf widens abruptly south of 48.5 N.
+Equatorward, the shelf widens further, and the shelf break has a series of submarine canyons, in particular Nitnat Canyon and Barkley Canyon.
 
-We present routine La Perouse hydrographic surveys along the LB and LC lines, collected aboard the \emph{CCGS Tully} from 2013-05-30 to 2013-05-31  (``June''), and 2013-09-07 to 2013-09-09 (``September'').
+We present observations from La Perouse hydrographic surveys along the LB and LC lines, collected aboard the \emph{CCGS Tully} from 2013-05-30 to 2013-05-31  (``May''), and 2013-09-07 to 2013-09-09 (``September'').
 These lines span the 50 m isobath to deep offshore, with casts every $7.5\ \mathrm{km}$ across shelf.
@@
 Hydrographic sections along the LB and LC lines highlight summer conditions on the Southern Vancouver Island Shelf and indicate some of the features we are focusing on in this paper (\Fref{fig:LaPerouse2013Ctd}).
-During both cruises, and at both lines, there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m depth offshore to shallower than 85 m over the shelf.
+During both surveys, and along both lines, there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m depth offshore to shallower than 85 m over the shelf.
 This upwelled water tends to be low in oxygen and cool.
-Further onshore the Vancouver Island Coastal Current hugs the coast, where isopycnals tilt back down, and the water is warmer than offshore.
+Further onshore, the Vancouver Island Coastal Current hugs the coast, where isopycnals tilt back down and water is warmer than offshore.
 
@@
      \includegraphics[width=130mm]{LaPerouse2013Ctd}
-     \caption{Observations along LC (top three rows) and LB lines (bottom three rows) in Jun (left column) and Sep 2013 (right column).
+     \caption{Observations along LC Line in May (a--c) and Sep (g--i) and LB lines in May (d--f) and Sep (j--l).
 X is across-shelf distance in the coordinate system shown in \fref{fig:LocMapBoth}.
@@
 Based on these properties, offshore water does not appear to make it over La Perouse bank into the basin onshore, or if it does, it does so intermittently.
-Note that the water along the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal becomes cooler and lower in oxygen where it intersects the shelf, consistent with both enhanced mixing near the shelf, and with drawdown of oxygen by respiration.
+Note that the water along the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal becomes cooler and lower in oxygen where it intersects the shelf, consistent with both enhanced mixing near the shelf and with drawdown of oxygen by respiration.
 
@@
 Denser than $26.6\ \mathrm{kg\,m^{-3}}$, the deep water masses found offshore are largely homogenous.
-In the lighter water masses, there are distinct differences between warm and salty offshore compared to water on the shelf at the same densities.
+In the lighter water masses, there are distinct differences between warm and salty offshore water compared to water on the shelf at the same densities.
 
-We use a spice anomaly relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}) representing an approximate mixing line for water in the \Eddy, and passsing through the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$.
-Spice anomaly for a given water sample at density $\sigma_{\theta}$ is given by $\gamma = \alpha \left(T - T_0\left(\sigma_{\theta}\right)\right) + \beta \left(S - S_0\left(\sigma_{\theta}\right)\right)$ where $T_0\left(\sigma_{\theta}\right)$, where $S_0\left(\sigma_{\theta}\right)$ are the temperature and salinity along the mixing line at the same density as the water sample, with the sign convention that positive is warmer and saltier than data along the mixing line.
+We use a spice anomaly defined as relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}) representing an approximate mixing line for water in the \Eddy, and passsing through the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$.
+Spice anomaly for a given water sample at density $\sigma_{\theta}$ is given by $\gamma = \alpha \left(T - T_0\left(\sigma_{\theta}\right)\right) + \beta \left(S - S_0\left(\sigma_{\theta}\right)\right)$ where $T_0\left(\sigma_{\theta}\right)$, where $S_0\left(\sigma_{\theta}\right)$ are the temperature and salinity along the mixing line at the same density as the water sample.
+The sign convention is that a positive anomaly is warmer and saltier than data along the mixing line.
 
@@
      \includegraphics[width=5in]{LaPerouse2013TS}
-     \caption{Potential temperature versus salinity for the La Perouse data shown in \fref{fig:LaPerouse2013Ctd}; potential density is contoured every $0.2\ \mathrm{kg\,m^{-3}}$, with the $26.4\ \mathrm{kg\,m^{-3}}$ colored magenta.
+     \caption{a) Potential temperature, $\theta$, versus salinity, $S$, for the La Perouse data shown in \fref{fig:LaPerouse2013Ctd}; potential density is contoured every $0.2\ \mathrm{kg\,m^{-3}}$, with the $26.4\ \mathrm{kg\,m^{-3}}$ colored magenta.
 A definition of spice anomaly is shown in this plot, and discussed in the text.
-The rectangle is the $\theta-S$ range used in \fref{fig:TSdensSpice} below.}
+The rectangle is the $\theta-S$ range used in \fref{fig:TSdensSpice} below.
+b) The same data with oxygen concentration versus salinity.}
      \label{fig:LaPerouse2013TS}
@@
 
-Using this measure of spice anomally, offshore water tends to have high absolute values (\fref{fig:LaPerouse2013Ctd}, middle rows), with a positive-anomaly layer (red) centered at $26.4\ \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and below.
-On the shelf ($-10\, \mathrm{km} < X < 30\, \mathrm{km}$), the distinct $\theta$--$S$ masses are attenuated and much closer to the defined mixing line than the offshore water (closer to white in color).
+Using this metric of spice anomaly, offshore water tends to have high absolute values (\fref{fig:LaPerouse2013Ctd}, middle rows), with a positive-anomaly layer (red) centered at $26.4\ \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and below.
+On the shelf ($-10\, \mathrm{km} < X < 30\, \mathrm{km}$), the distinct $\theta$--$S$ masses are attenuated and much closer to the defined mixing line than the offshore water (closer to white in color in \fref{fig:LaPerouse2013Ctd}).
 
@@
 Hugging the coast ($X > 30\, \mathrm{km}$), the Vancouver Island Coastal Current warms during the summer, and the spice anomaly goes from negative to postive.
-In the \Eddy, the water stays near the mixing line, but is cooler in the spring (\fref{fig:LaPerouse2013Ctd}, LB Line, left-hand column: slightly negative spice anomally) and warms during the summer (\fref{fig:LaPerouse2013Ctd}, LB Line, right-hand column: slightly positive spice anomally).
+In the \Eddy, the water stays near the mixing line, but is cooler in the spring (\fref{fig:LaPerouse2013Ctd}, LB Line, left-hand column: slightly negative spice anomaly) and warms during the summer (\fref{fig:LaPerouse2013Ctd}, LB Line, right-hand column: slightly positive spice anomaly).
 
-In terms of oxygen, both surveys have a similar dichotomy between the \Eddy and offshelf water.
-Water found in the \Eddy is at least $100\ \mathrm{\mu mol\, kg^{-1}}$ lower on the shelf than offshelf (\fref{fig:LaPerouse2013Ctd}, \fref{fig:LaPerouse2013TS}b).
-The very deepest water ($S\approx 33.9\ \mathrm{psu}$) on the shelf shows a further $50\ \mathrm{\mu mol\, kg^{-1}}$ drop between June and September along the LB line.
+Oxygen concentration in both surveys has a similar dichotomy between the \Eddy\ and offshelf water.
+Water found in the \Eddy\ has oxygen concentrations $100\ \mathrm{\mu mol\, kg^{-1}}$ lower on the shelf than offshelf (\fref{fig:LaPerouse2013Ctd}, \fref{fig:LaPerouse2013TS}b).
+The very deepest water ($S\approx 33.9\ \mathrm{psu}$) on the shelf shows a further $50\ \mathrm{\mu mol\, kg^{-1}}$ decrease in concetration between May and September along the LB line.
 
@@
 
-The finescale survey covered most of the \Eddy region, with an emphasis on the offshore edge near the shelfbreak front (\fref{fig:LocMapBoth}).
-For water denser than $26\ \mathrm{kg\,m^{-3}}$, there are three distinct water masses found on the shelf (\fref{fig:TSdensSpice}).
-The first is the offshore water, which tends to be warmer, and hence have a positive spice anomally ($\gamma \approx 0.2\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$).
+The finescale surveys covered most of the \Eddy\ region, with an emphasis on the offshore edge near the shelfbreak front (\fref{fig:LocMapBoth}).
+For water denser than $26\ \mathrm{kg\,m^{-3}}$, there are three distinct water masses sampled on the shelf (\fref{fig:TSdensSpice}).
+The first is the offshore water, which tends to be warmer, and hence has a positive spice anomaly ($\gamma \approx 0.2\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$).
 The second is water in the \Eddy, which during this survey was found along the straight mixing line in $\theta-S$ space (\fref{fig:TSdensSpice}, $\gamma \approx 0\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$).
-Between these two water masses there is a less populous mass (\fref{fig:TSdensSpice}, $\gamma \approx 0.1\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$) that we demonstrate below is found on the shelf poleward of the \Eddy.
+Between these two water masses, there is a less populous mass (\fref{fig:TSdensSpice}, $\gamma \approx 0.1\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$) that we demonstrate below is found on the shelf poleward of the \Eddy.
 
@@
      \includegraphics[width=3.5in]{TSdensSpice.pdf}
-    \caption{Binned sample density of salinity and potential density data from the cruise, with a logarithmic color scale.
-Grey contours are potential density relative to the surface, every $0.1\ \mathrm{kg\,m^{-3}}$, and the magenta contour is the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.
-Colored contours are spice anomaly, $\gamma$, relative to the straight line labeled $\gamma=0.0$.
+    \caption{Binned sample density of salinity and potential temperature from the cruise, with a logarithmic color scale.
+Grey contours are potential density relative to the surface at intervals of $0.1\ \mathrm{kg\,m^{-3}}$; the magenta contour is the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.
+Colored contours are spice anomaly, $\gamma$, relative to the white line labeled $\gamma=0.0$.
     \label{fig:TSdensSpice}}
@@
 
-In synthetic cross sections of spice anomaly, the \Eddy\ is clearly identifiable as having low-spice anomaly ($\gamma \approx 0\ \mathrm{kg\,m^{-3}}$, \fref{fig:CrossSectionsSpice}c, d).
+In synthetic cross sections of spice anomaly, the \Eddy\ is clearly identifiable as having low spice anomaly ($\gamma \approx 0\ \mathrm{kg\,m^{-3}}$, \fref{fig:CrossSectionsSpice}c, d).
 This signature of well-mixed water extends from the seafloor to approximately 30 m depth, onshore of $X \approx 0\ \mathrm{km}$.
 Despite lying along a mixing line, the \Eddy\ water is still stratified.
-The high-spice water ($\gamma \approx 0.2\ \mathrm{kg\,m^{-3}}$) is found offshore of the \Eddy water, on the other side of a sharp $\theta-S$ compensated front.
+The high-spice water ($\gamma \approx 0.2\ \mathrm{kg\,m^{-3}}$) is found offshore of the \Eddy\ water, on the other side of a sharp $\theta-S$ compensated front.
 This front is even sharper in individual sections than in these composite sections (see \fref{sec:frontsurvey}).
 
-In contrast, the third population of partially mixed water mass identified in \fref{fig:TSdensSpice} is all found upstream of the \Eddy\ region (\fref{fig:CrossSectionsSpice}a, b) as a weaker spice anomaly ($\gamma \approx 0.1\ \mathrm{kg\,m^{-3}}$, ``pink'' colors).
-This water is generally found near the intersection of the upwelling water and La Perouse Bank, and is not as warm as offshore water, indicating some mixing has taken place.
-Of note, is that this intermediate water mass appears completely absent in the sections further equatorward (\fref{fig:CrossSectionsSpice}c, d), indicating that it is either mixed away, or that it is advected elsewhere; we argue below that it is advected offshore in a large-scale tongue.
+In contrast, the third population of partially mixed water between the \Eddy\ and offshore water (\fref{fig:TSdensSpice}) is all found poleward of the \Eddy\ region (\fref{fig:CrossSectionsSpice}a, b) as water with a weaker spice anomaly ($\gamma \approx 0.1\ \mathrm{kg\,m^{-3}}$, pink colors).
+This water is generally found near the intersection of the upwelling water and La Perouse Bank, and is not as warm as offshore water, indicating some mixing with offshore water has taken place.
+Of note is that this intermediate water mass appears completely absent in the sections further equatorward (\fref{fig:CrossSectionsSpice}c, d), indicating that it is either mixed away or that it is advected elsewhere; we argue below that it is advected offshore in a large-scale tongue.
 
@@
 
-Oxygen section show similar trends, with the low-spice anomaly water also having low oxygen saturations, while the strong negative spice anomaly shallower than $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ has very high oxygen; deeper water is also more oxygenated.
-Of note is that compared with the coarser sampling along the LC Line (\fref{fig:LaPerouse2013Ctd}), we see that the abrupt transition between offshore and \Eddy\ water is also present in oxygen.
+Oxygen saturation sections show similar trends, with the low-spice anomaly water also having low oxygen saturation, while the strong negative spice anomaly shallower than $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ has very high oxygen saturation; deeper water is also more oxygenated.
+Of note is that compared with the coarser sampling along the LC Line (\fref{fig:LaPerouse2013Ctd}), we see that the abrupt transition between offshore and \Eddy\ water is also present in disolved oxygen.
 
@@
     \includegraphics[width=6.2in]{CrossSectionsO2}
-    \caption{Synthetic sections of Oxygen, approximate percent saturation from the MVP data.
+    \caption{As in \fref{fig:CrossSectionsSpice}, for oxygen approximate saturation.
       \label{fig:CrossSectionsO2} }
@@
 
-The spatial patterns are very clear when considering a map-view of properties along the $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ isopycnal (\fref{fig:SpiceO2264}).
-There is a region onshore of the south tip of La Perouse Bank (approximately 44.5 N, 125.75 W), that consists of water that is found along the mixing line (spice anomaly $\gamma \approx 0 \mathrm{kg\,m^{-3}}$) with low oxygen saturations.
+The spatial patterns are very clear when considering a map view of properties along the $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ isopycnal (\fref{fig:SpiceO2264}).
+There is a region onshore of the south tip of La Perouse Bank (approximately 44.5\textdegree N, 125.75\textdegree W), that consists of water that is found along the mixing line (spice anomaly $\gamma \approx 0\ \mathrm{kg\,m^{-3}}$) with low oxygen saturation.
 Offshore of this region is water that is very warm and salty in comparison ($\gamma > 0.15 \mathrm{kg\,m^{-3}}$), and relatively high in oxygen (saturations of approximately 60\%).
-The region between these two water masses is very abrupt, and is concave with respect to the isobaths on the shelf,  stretching from La Perouse Bank to a shallow bank just above the Juan de Fuca canyon (48.3 N and 125.4 W).
+The region between these two water masses is very abrupt, and  stretches from La Perouse Bank to a shallow bank just above the Juan de Fuca canyon (48.3\textdegree N and 125.4\textdegree W).
 
-Upstream of La Perouse bank and Barkley Canyon, the water along the shelf has the weaker spice anomaly of the third water mass ($\gamma \approx 0.1\,\mathrm{kg\,m^{-3}}$, light pink colors).
+Poleward of La Perouse bank and Barkley Canyon, the water along the shelf has the weaker spice anomaly characteristic of the third water mass ($\gamma \approx 0.1\,\mathrm{kg\,m^{-3}}$, light pink colors).
 This water is also somewhat lower in oxygen than water from offshore, though not as depleted as the water in the \Eddy.
-This water appears to be pushed offshore just upstream of Barkley Canyon, and replaced on the shelf by the warmer (high spice) offshore water.
+This water appears to be pushed offshore just upstream of Barkley Canyon and replaced on the shelf by the warmer (high spice) offshore water.
 
@@
     \includegraphics[width=6.2in]{SpiceO2264}
-    \caption{Spatial overview of a) local spice (temperature anomaly on an isopycnal: see text), and b) oxygen saturation on the $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.
+    \caption{Spatial overview of a) the spice anomaly, and b) oxygen saturation on the $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.
 Grey cross-slope lines are cross sections indicated in \fref{fig:CrossSectionsSpice}.
-Along-slope grey lines are every 20 km in the cross-slope direction, with $x=0\ \mathrm{km}$ near the 100-m isobath at the north end of the observation area.
+Along-slope grey lines are every 20 km in the cross-slope direction, with $X=0\ \mathrm{km}$ near the 100-m isobath at the north end of the observation area.
    \label{fig:SpiceO2264}
@@
 The Spur Canyon leading from the Strait of Juan de Fuca has been implicated in allowing dense water to be upwelled into the \Eddy.
-The sea surface is low in the middle of the \Eddy, so water has been hypothesized to move up the Spur Canyon due to ageostrophic motion \cite{weaverhsieh87,freelanddenman82}.
+The sea surface is low in the middle of the \Eddy, so it has been hypothesized that water moves up the Spur Canyon due to ageostrophic motion \cite{weaverhsieh87,freelanddenman82}.
 This is difficult to infer from the observations collected here.
 Three transects up the canyon indicate that deep isopycnals slope up into the canyon (\fref{fig:CanyonPropertiesSpice}, to approximately 30 km).
-Continuing across the \Eddy, isopycnals are largely flat until they intersect the Vancouver Island Coastal Current ($x\approx 80\ \mathrm{km}$).
-The deeper isopycnals do not make it into the deeper basin northeast of La Perouse bank ($x\approx 70\ \mathrm{km}$),  thus the bank is a natural northern boundary of the \Eddy.
+Continuing across the \Eddy, isopycnals are largely flat until they intersect the Vancouver Island Coastal Current ($80\ \mathrm{km}$).
+The deeper isopycnals are not found in the deeper basin northeast of La Perouse bank ($70\ \mathrm{km}$),  thus the bank is a natural northern boundary of the \Eddy.
 
@@
     \includegraphics[width=6.2in]{CanyonPropertiesSpice}
-    \caption{Three surveys up the Spur Canyon, where $X$ is along-canyon as defined in the map.
+    \caption{Spice anomaly surveys up the Spur Canyon, where $X$ is along-canyon as defined in the map.
 Grey isopycnals are contoured every $0.5\,\mathrm{kg\,m^{-3}}$, and the $26.4\,\mathrm{kg\,m^{-3}}$ isopycnal is shown in magenta.
 The seafloor is indicated with the thick black line.
-      Map shows the path taken during each survey in chronological order blue, orange, and green.  Magenta line is path taken during a cross-canyon survey (\fref{fig:HydraulicsCanyon}).
+Map (lower right) shows the path taken during each survey in chronological order (blue, orange, and green).
+Magenta line is path taken during a cross-canyon survey (\fref{fig:HydraulicsCanyon}).
       \label{fig:CanyonPropertiesSpice} }
@@
 
-Water properties along the canyon go from offshore water, to \Eddy\ water.
-Oxygen behaves in a similar manner, though some of the incoming water has slightly lower oxygen than water inside the \Eddy\ (not shown).
+Spice anomaly along the canyon indicates a transition from offshore water to \Eddy\ water.
+Oxygen saturation behaves in a similar manner, though some of the incoming water has slightly lower oxygen than water inside the \Eddy\ (not shown).
 Based on these sections, it is difficult to infer water motion up the canyon.
@@
 Much of the modified water in the Spur Canyon appears to come from the shelf to the west, but heavily modified by tidal mixing.
-A repeated tidal survey over the bank on the west side of the canyon shows a strong hydraulic response (\fref{fig:HydraulicsCanyon}) during onshore flow (17:58--21:00).
-Dense ocean water passes from the offshore side into the Canyon, and plunges down the side wall before rebounding downstream.
+A repeated tidal survey over the bank on the west side of the canyon shows a strong hydraulic response during onshore flow (17:58--21:00, \fref{fig:HydraulicsCanyon}).
+Dense water passes from the offshore side into the canyon, and plunges down the side wall before rebounding downstream.
 Note that the tide here is largely diurnal, so this onslope flow only occurs once a day.
 Given the stratification of $N\approx 6\times10^{-3} \ \mathrm{rad\,s^{-1}}$ and an overturning scale of 50 m, we might expect dissipation rates reaching $\epsilon \sim L^2N^{3} \approx 5\times10^{-4}\ \mathrm{m^2\,s^{-3}}$, which is more than three orders of magnitude higher than dissipation observed west of this location by \citeA{deweycrawford88}.
-In terms of a diapycnal diffusivity $\kappa = \gamma \epsilon / N^2 \approx 1\ \mathrm{m^2\,s^{-1}}$, assuming a mixing efficiency of $\gamma=0.2$ The effect on the properties in the canyon is for water to spill over from the offshore front into the canyon during the tide.
+This estimate of turbulence dissipation rate implies a diapycnal diffusivity of $\kappa = \gamma \epsilon / N^2 \approx 1\ \mathrm{m^2\,s^{-1}}$, assuming a mixing efficiency of $\gamma=0.2$.
+Water spills over from the offshore front into the canyon during the onshore tide.
 This water is rapidly mixed with surrounding water such that its strong offshore spice values are attenuated.
@@
     \includegraphics[width=4in]{HydraulicsCanyon}
-    \caption{Repeated, tide-resolving survey across the Spur Canyon, time is indicated in the upper left of each plot (29 August, 2013).
-See magenta line on map \fref{fig:CanyonPropertiesSpice} for location of survey.
+    \caption{Spice anomaly observed in repeated, tide-resolving survey across the Spur Canyon, time is indicated in the upper left of each plot (29 August, 2013).
+Location of survey shown in \fref{fig:CanyonPropertiesSpice} as a magenta line.
       \label{fig:HydraulicsCanyon}
@@
 
-The cross-canyon turbulence makes it ambiguous if there is water moving up the canyon or not.
-There is a general tendency along the canyon for higher spice water to be found offshore (\fref{fig:CanyonPropertiesSpice}) but it seems likely the source of the higher spice is over the bank rather than water being advected up the canyon.
-This tidally driven flow over the bank is the strongest source of high-spice offshore water into the eddy region identified during our surveys.
+The turbulence found on the canyon rim makes it ambiguous if there is water moving up the canyon or not.
+There is a general tendency along the canyon for higher spice water to be found offshore (\fref{fig:CanyonPropertiesSpice}) but it seems likely the source of the higher spice water is from over the bank rather than water being advected up the canyon.
+This tidally driven flow over the bank is the most significant source of high-spice offshore water into the eddy region identified during our surveys.
 
@@
 
-A systematic survey through the front between the offshore water and the \Eddy\ water demonstrates the sharpness and persistence of this front (\fref{fig:Frontsurvey}), giving evidence that it has limited exchange with the offshore region.
-The survey started close to shore and passed through the Vancouver Island Coastal Current (along-track 0-10 km).
-The coastal current water is buoyant and significantly fresher and cooler than water at the same density near the surface and forms a classic density front with the water in the Eddy region.
-There is some partially mixed (pink in spice spice) water in this density front due to mixing with the warmer water near the surface.
+A systematic survey through the front between the offshore water and the \Eddy\ water demonstrates the sharpness and persistence of this front (\fref{fig:Frontsurvey}), suggesting that it has limited exchange with the offshore region.
 
-Offshore of this coastal current, measurements were collected through the front between the offshore water and the \Eddy\ water 10 times, showing its evolution following the along-shore equatorward flow.
+The survey started close to shore, and passed through the Vancouver Island Coastal Current (along-track 0-10 km).
+The coastal current forms a buoyant front, and is fresher and colder than water at the same density.
+This front is relatively thick, greater than 20 km wide, and has partially mixed water from the surface to the foot of the front ($\gamma\approx0.1$).
+
+Offshore of this coastal current, measurements were collected crossing the front between the offshore water and the \Eddy\ water 10 times, showing its evolution following the along-shore equatorward flow.
 First, as noted in the composite sections, isopycnals slope up from offshore onto the shelf.
-The front starts out very sharp, (along-track distance $50\ \mathrm{km}$) though two small tendrils can be seen separating from the front on the inshore side.
-Similar tendrils are found on the second pass, perhaps a bit more separated from the front ($\approx 80\ \mathrm{km}$), and on the third pass ($\approx 95\ \mathrm{km}$).
+In the first crossing, the front is very sharp, (along-track distance $50\ \mathrm{km}$) though two small tendrils can be seen separating from the front on the inshore side.
+Similar tendrils are found on the second crossing, perhaps a bit more separated from the front ($\approx 80\ \mathrm{km}$), and on the third crossing ($\approx 95\ \mathrm{km}$).
 These tendrils are made of up partially mixed water.
-The subsequent passes have more of this partially mixed water, such that the partially mixed region is up to 5-km wide by the fifth pass ($\approx 150\ \mathrm{km}$).
-However, note that the deeper isopycnals retain a sharp front, and indeed the front appears sharp again by the seventh pass at all depths.
+The subsequent passes have more of this partially mixed water, such that the partially mixed front is up to 5-km wide by the fifth pass ($\approx 150\ \mathrm{km}$).
+However, the deeper isopycnals retain a sharp front, and indeed the front appears sharp again by the seventh pass at all depths ($\approx 200\ \mathrm{km}$).
 
@@
 Regions of warmer (and more oxygenated) water are found in tendrils at these depths (e.g.\ $\approx\,170\ \mathrm{km}$ and $\approx 185\,\mathrm{km}$).
-The overall effect is similar to what was seen in the Gulf Stream with similar observations \cite{klymaketal16}; there are two quite distinct water masses, the Eddy water and the offshore water, as seen in the $\theta$--$S$ plot (\fref{fig:Frontsurvey}c).
-There is only a small population of samples between these two, indicative of substantial isopycnal and vertical mixing, but even these populations are relatively cut off from the main water masses, indicating that they are well-mixed on their own, in short episodic events.
-However, there is little evidence of large-scale flow into or out of the Eddy at all depths during this observation period.
+The overall effect is similar to what was seen in the Gulf Stream with similar observations \cite{klymaketal16}; there are two quite distinct water masses, the Eddy water and the offshore water, as seen in the $\theta$--$S$ plot (\fref{fig:Frontsurvey}c) with only a small population of samples between these two.
+This distribution of $\theta$--$S$ properties is indicative of substantial isopycnal and vertical mixing, but even these populations are relatively cut off from the main water masses, indicating that they are well-mixed on their own, in short episodic events.
+Regardless, this front is very sharp given that it has no density signature, indicating that there is not strong advection from offshore into the \Eddy\ region.
 
@@
     \includegraphics[width=6.2in]{Frontsurvey}
-    \caption{Data from a survey focusing on the front region (2013-08-28T15:11 to 2013-08-29T09:18) a) spice anomaly along $26.4\,\mathrm{kg\,m^{-3}}$.
+    \caption{Data from the frontal survey  a) spice anomaly along $26.4\,\mathrm{kg\,m^{-3}}$.
 Grey-white alternating lines are the coordinate system, with 10-km alternating shades.
-b) oxygen saturation; magenta dots correspond to distance along-track.
+b) oxygen saturation along $26.4\,\mathrm{kg\,m^{-3}}$; magenta dots correspond to distance along-track.
 c) density of data points (logscale) in this section of data.
@@
 Upstream of the the sharp front between the low-spice \Eddy\ and the high-spice offshore water is a substantial mass of intermediate-spice water along the shelf (compare \fref{fig:CrossSectionsSpice}a and c).
-If in geostrophic balance, intermediate spice water is moving equatorward along the shelf upstream of the \Eddy, but is pushed offshore just upstream of Barkley Canyon (\fref{fig:IsopycnalsDepth}).
-There is a tongue of intermediate-spice water ($\gamma$ is pink along $26.4\,\mathrm{kg\,m^{-3}}$) that separates from the shelf just west of 126 W.
-We had trouble crossing the whole tongue with the scale of our surveys, but it is at least 30 km wide.
-It also appears to end at approximately 48.2 N.
+If in geostrophic balance, this intermediate spice water is moving equatorward along the shelf upstream of the \Eddy, but is pushed offshore just upstream of Barkley Canyon (\fref{fig:IsopycnalsDepth}).
+There is a tongue of intermediate-spice water ($\gamma$ is pink along $26.4\,\mathrm{kg\,m^{-3}}$) that separates from the shelf just west of 126\textdegree W.
+The surveys do not cross the full extent of the  tongue, but it is at least 30 km wide.
+It also appears to end at approximately 48.2\textdegree N.
 This intermediate-spice water reaches from relatively shallow isopycnals to at least $26.6\,\mathrm{kg\,m^{-3}}$ (\fref{fig:IsopycnalsDepth}, left panel).
 Unfortunately, we cannot track the fate of this water mass because isopycnals tilt down offshore, below the depth limit of the MVP.
-The shallower isopycnal appears to have the tongue closer to the shelf than at $26.4\,\mathrm{kg\,m^{-3}}$, indicating strong three-dimensionality to this feature.
+The $25.5\,\mathrm{kg\,m^{-3}}$ isopycnal appears to have the tongue closer to the shelf than at $26.4\,\mathrm{kg\,m^{-3}}$, indicating strong three-dimensionality to this feature.
 
-This separating tongue is embedded in the larger scale isopycnal tilt caused by the upwelling (\fref{fig:IsopycnalsDepth} right panels), so it is difficult to see dynamically what is driving this offshore push.
-One possibility is that it is simple flow separation as the water is not able to make the sharp turn around La Perouse Bank.
+This separating tongue is embedded in the larger scale isopycnal tilt caused by the upwelling (\fref{fig:IsopycnalsDepth}, right panels), so it is difficult to see dynamically what is driving this offshore push.
+One possibility is that it is simple flow separation caused by the water not being able to make the sharp turn around La Perouse Bank.
 Whatever causes it, downstream of the separation, the water is replaced by offshore water with much higher spice values.
@@
 There is a clear surface expression of the separating tongue in satellite imagery (\fref{fig:SST}).
-Water flowing equatorward along the shelf tends to be cooler than offshore water, likely mixing with the colder coming out of the Strait of Juan de Fuca.
-On Aug 25, there is a cold tongue of water separating from La Perouse bank, crossing isobaths and pointing south at 125.8 W.
-There is cooler tendril streaming west at 48 N off the south end of this tongue.
+Water flowing equatorward along the shelf tends to be cooler than offshore water, likely due to mixing with the colder coming out of the Strait of Juan de Fuca.
+On August 25, there is a cold tongue of water separating from La Perouse bank, crossing isobaths and pointing south at 125.8\textdegree W.
+There is a cooler tendril streaming west at 48 N off the south end of this tongue.
 This feature is not as well-developed in the previous image (21 August) perhaps indicating that it is an evolving feature.
 By 31 August, there is no surface expression of the feature, though small tendrils of cooler water can be seen separating from La Perouse Bank.
-By 5 September, the water has significantly warmed, but the offshore anomaly does not appear to have a surface signature.
+By 5 September, the water has significantly warmed, and the offshore anomaly does not appear to have a surface signature.
 
@@
 Satellite-based surface chlorophyll estimates show the same feature (\fref{fig:ChlA}) demonstrating the advection of high chlorophyll to the west side of the \Eddy.
-It also shows a relatively high-chlorophyll tendril to the west, again exiting the study region at approximately 48 degrees N.
+They also show a relatively high-chlorophyll tendril to the west, again exiting the study region at approximately 48 degrees N.
 The feature is relatively long-lived, on the order of one month.
-Inspection of images before 2013-08-05 were too obscured by clouds, or did not show this feature.
-By 2013-09-06, we see the feature fading from the satellite.
-However, note that this feature is centered 0.2 degrees of latitude south of tongue that we observe deeper in the water column, again indicating that there is some depth-dependence to the feature.
+Inspection of images before August 5 were too obscured by clouds or did not show this feature.
+By September 6, we see the feature fading from the satellite image.
+Note that this feature is centered 0.2 degrees of latitude south of tongue that we observe deeper in the water column, again indicating that there is depth-dependen structure in the feature.
 
@@
     \includegraphics[width=6.2in]{ChlA}
-    \caption{Surface chlorophyll density estimated from ocean color \cite{Huetal12,MODISChlL3} over 8-day windows in 4-km bins. Gray regions had too many clouds for acceptable averages.
+    \caption{Surface chlorophyll density estimated from ocean color \cite{Huetal12,MODISChlL3} over 8-day windows in 4-km bins. Gray regions had too many clouds to compute averages.
       \label{fig:ChlA} }
@@
 The intensive sampling discussed here has demonstrated a few important features of the South Vancouver Island Shelf.
-The Juan de Fuca Eddy water is readily identified as well-mixed withe stratified water found along a straight line in $\theta-S$ space, compared with offshore water that was warmer and saltier (high-spice).
+The Juan de Fuca Eddy water is readily identified as falling along a mixing line in $\theta-S$ space, compared with offshore water that was warmer and saltier (high-spice anomaly).
 There was not strong evidence of the \Eddy\ being supplied by water moving up the Spur Canyon during our observations, but the Spur Canyon was a site of hydraulic cross-canyon flows in which we infer significant mixing.
-There is a sharp and persistent temperature-salinity compensated front between offshore water, and the mixed \Eddy\ water.
-Finally, upstream of the \Eddy, water in the equatorward shelf current has intermediate spice, and is seen to separate from the shelf at the point of an abrupt bend in an underwater bank.
-The water mass crosses isobaths and is ejected into the interior, also be observed from satellites.
-Thus the water that is offshore of the \Eddy\ appears to have been brought onto the shelf in exchange with this offshore tongue.
+There is a sharp and persistent temperature-salinity compensated front between offshore water and the partially mixed \Eddy\ water.
+Finally, upstream of the \Eddy, water in the equatorward shelf current has intermediate spice anomaly, and is seen to separate from the shelf at the point of an abrupt bend in an underwater bank.
+The water mass crosses isobaths and is ejected into the interior.
+This separation event can also be observed from satellite measurements of seasurface temperature and chlorophyll.
+Thus the water that is offshore of the \Eddy\ appears to have been brought onto the shelf in exchange for the offshore ejection of shelf water via this tongue.
 
@@
 
-The source of water and formation of the Juan de Fuca Eddy has received substantial attention, however, the observations of a tight mixing line in $\theta$--$S$ space presented here are a new observation.
-The deepest water in the \Eddy\ could originate along the $\theta$--$S$ line from approximately 5.5 degrees to 7.5 degrees (\fref{fig:LaPerouse2013TS}a), which in the open ocean spans depths from 420 m to 70 m.
+The source of water and formation mechanism of the Juan de Fuca Eddy has received substantial attention, however, the observations of \Eddy\ waters being found along a tight mixing line in $\theta$--$S$ space has not previously been noted.
+The deepest water in the \Eddy\ could originate along the $\theta$--$S$ line from approximately $5.5^\circ \mathrm{C}$ to $7.5^\circ \mathrm{C}$ (\fref{fig:LaPerouse2013TS}a), which in the open ocean spans depths from 420 m to 70 m.
 \citeA{mackasetal87} attempted to determine the origin of the water by including oxygen as a third variable to resolve the ambiguous $\theta$--$S$ relation.
-However, as \fref{fig:LaPerouse2013TS}b makes clear, oxygen is definitely not a conserved property in the \Eddy, with concentrations up to $150 \ \mathrm{\mu mol\ kg^{1}}$ lower in the \Eddy\ than the water found offshore, and in a way that definitely cannot result from conservative mixing.
+However, as \fref{fig:LaPerouse2013TS}b makes clear, oxygen does not appear to be a conserved property in the \Eddy, with concentrations up to $150 \ \mathrm{\mu mol\ kg^{1}}$ lower in the \Eddy\ than the water found offshore, and in a way that definitely cannot be the result of conservative mixing.
 As a best guess, if the water in the deepest part of the \Eddy\ came from a vertical mixture of the water at $26.5\,\mathrm{kg\,m^{-3}}$ and an equal distance down in density space of $26.7\,\mathrm{kg\,m^{-3}}$, then the deepest water in the \Eddy\ may be coming from a depth of 250 m.
@@
 
-We found little evidence of vigorous flow up the Spur Canyon, or if there is, then the mixing in the canyon is quick enough to remove the $\theta$--$S$ signature of offshore water within 20 km of the canyon mouth (\fref{fig:CanyonPropertiesSpice}).
-We did find substantial evidence of mixing in the canyon, but the primary route of high-spice water into the canyon appears to be due to tidal flow over the banks on the west side (\fref{fig:HydraulicsCanyon}), rather than evidence of flow up the canyon.
-However, it is also of note that upwelling winds had ceased at this point (\fref{fig:LaPeWind}), meaning that the offshore surface pressure gradient may be reduced, leading to reduced ageostrophic upwelling in the canyon.
+We found little evidence of flow up the Spur Canyon or, if there is, then the mixing in the canyon is intense enough to remove the $\theta$--$S$ signature of offshore water within 20 km of the canyon mouth (\fref{fig:CanyonPropertiesSpice}).
+We did find substantial evidence of mixing in the canyon, but the primary pathway of high-spice water into the canyon appears to be due to tidal flow over the banks on the west side (\fref{fig:HydraulicsCanyon}), rather than flow up the canyon.
+However, it is worthy of note that upwelling winds had ceased at the point of these observations (\fref{fig:LaPeWind}), so the offshore surface pressure gradient may be reduced, leading to reduced ageostrophic upwelling in the canyon.
 Whether the cessation of winds would also lead to a reduction of the low sea level height in the center of the \Eddy\ that may drive up-canyon flow is an open question.
 
-There is evidence of aging of the water in the \Eddy\ between late spring and late summer (\fref{fig:LaPerouse2013Ctd}), with a reduction of oxygen in the \Eddy.
+There is evidence of aging of the water in the \Eddy\ between late spring and late summer (\fref{fig:LaPerouse2013Ctd}), with a reduction of oxygen in the \Eddy\ over this time span.
 If we posited that the reduction was all in the same water, then the oxygen consumption rate over the time between the late-May and early September cruises would be on the order of $0.5\ \mathrm{\mu mol\, kg^{-1} d^{-1}}$.
-This seems on the low side of continental upwelling system estimates of apparent oxygen utilization rates, which are between $1$ and $5\  \mathrm{\mu mol\, kg^{-1} d^{-1}}$ \cite{dortchetal94}.
-So it seems possible the \Eddy\ has enough exchange with the surrounding water for the residence time to be less than the whole time period from late May to September.
-Note also that during the June cruise, the water in the \Eddy\ is slightly cooler than the mixing line, and during the September cruise is slightly warmer than the mixing line, so the water in the \Eddy\ is evolving seasonally, and probably affected by the water temperature in the Vancouver Island Coastal Current, and the amount of water getting through the front.
+This is on the low side of estimates of apparent oxygen utilization rates in continental upwelling systems, which are between $1$ and $5\  \mathrm{\mu mol\, kg^{-1} d^{-1}}$ \cite{dortchetal94}.
+So it seems possible the \Eddy\ had enough exchange with the surrounding water for the residence time to have been less than the full time period from late May to September.
+Note that during the May cruise, the water in the \Eddy\ is slightly cooler than the mixing line, and during the September cruise is slightly warmer than the mixing line, so the water in the \Eddy\ is evolving seasonally, and probably affected by the water temperature in the Vancouver Island Coastal Current and the amount of water getting through the onshore front.
 
@@
     \includegraphics[width=4in]{TSExercise}
-    \caption{Mixing model assuming constant eddy diffusivity of $\kappa = 10^{-3}\ \mathrm{m^2\,s^{-1}}$ acting on an offshore temperature and salinity profiles compressed from 250 thick to 150 m thick.
-Profiles are pinned to the deep value at ($34\ \mathrm{psu}$, $6.2\ \mathrm{^oC}$), and a shallow one at ($31.6\ \mathrm{psu}$, $12\ \mathrm{^oC}$).
+    \caption{Mixing model assuming constant eddy diffusivity of $\kappa = 10^{-3}\ \mathrm{m^2\,s^{-1}}$ acting on an offshore temperature and salinity profiles compressed from 250 thick to 150 m thick (grey dashed lines).
+Profiles are pinned to the deep $\theta-S$ value at ($34\ \mathrm{psu}$, $6.2\ \mathrm{^oC}$), and a shallow one at ($31.6\ \mathrm{psu}$, $12\ \mathrm{^oC}$).
+The green dashed line is the mixing line defined in the text.
       \label{fig:TSExercise} }
@@
 The sharpness of the front with the \Eddy\ and the offshore water is intriguing.
-It was persistent for the duration of our detailed survey (\fref{fig:SpiceO2264}), and, so far as we can tell with the limited resolution, was present during the bracketing La Perouse cruises.
+It was persistent for the duration of our detailed survey (\fref{fig:SpiceO2264}), and, so far as we can tell with the limited resolution of the hydrographic surveys, was present during the bracketing La Perouse cruises.
 There is not any substantial bathymetry blocking the onshore incursion of water at this location, so there must be a dynamic barrier.
-In contrast, numerical simulations of this region reported by \cite{sahuetal22} using a NEMO 36th-degree regional model  indicate that there is more rapid exchange (order 40 days) between the deep \Eddy\ water and the rest of the coastal ocean.
+
+Numerical simulations of this region reported by \cite{sahuetal22} using a NEMO 36th-degree regional model  indicate that there is more rapid exchange between the deep \Eddy\ water and the rest of the coastal ocean.
 In this model, the region where the \Eddy\ resides has velocities equal or greater than other parts of the shelf, and water has an approximate residence time of less than 20 days.
-The \Eddy\ water in the model does develop a distinct $\theta$--$S$ signature, but not nearly as strong as observed.
-It also has a front with the offshore water, but the front is more gradual than that observed here.
+The \Eddy\ water in the model does develop a distinct $\theta$--$S$ signature, but not along a sharp mixing line as observed.
+It also has a front with the offshore water, but the front is substantially wider than that observed here.
 
 Overall, it would be an improvement to our understanding of the \Eddy\ if we could sample the shelf more persistently.
-The \Eddy\ was already well-formed by the June La Perouse cruise, and seems to evolve slowly during that time.
-Capturing its formation, presumably earlier in the spring, and its evolution through the year would be valuable in understanding retention and exchange on this productive part of the shelf.
+The \Eddy\ was already well-formed by the May La Perouse cruise, and seems to evolve slowly during that time.
+Capturing its formation, presumably earlier in the spring, as well as its evolution through the year, would be valuable in understanding retention and exchange on this productive part of the shelf.
 
@@
 The (apparent) displacement of shelf water from La Perouse Bank is a dramatic departure from geostrophically balanced isobath-following flow.
-Eddies have been known to separate from irregular coastal topography, both at the surface \cite{barthetal00}, and deeper in the water column \cite{pellandetal13}.
+Eddies have been known to separate from irregular coastal topography, both at the surface \cite{barthetal00} and deeper in the water column \cite{pellandetal13}.
 It has been recognized that instabilities lead to exchange between the open ocean and the shelf at this location \cite{ikedaemery84, ikedaemery84}.
 However, observations of the wholesale replacement of shelf water by a new water mass from offshore are rare.
-In our situation it is clear that water from as deep as 150 m is separating from the shelf and moving offshore (\fref{fig:SpiceO2264}, \fref{fig:IsopycnalsDepth}).
+In these observations it is clear that water from as deep as 150 m is separating from the shelf and moving offshore (\fref{fig:SpiceO2264}, \fref{fig:IsopycnalsDepth}).
 
-Satellite imagery makes it clear that there is often an exchange between coastal and deep water along the Vancouver Island shelf (\fref{fig:SSTLateAug}).
-Most years there are three of four large excursions from the shelf into the open ocean, many of them over 100 km long.
+Satellite imagery makes it clear that there is often exchange between coastal and deep waters along the Vancouver Island shelf (\fref{fig:SSTLateAug}).
+Most years there are three of four large filaments from the shelf into the open ocean, many of them over 100 km long.
 This length scale is longer than the 60 km inferred for this region by \citeA{ikedaetal84} using a four-layer instability analysis.
-It is possible that there is spatial locking of these features, with a persistent separation at the north tip of the Island, and a strong tendency for one at 49.5 N.
-Our study area also has evidence of a separation in most of the years, with 2011 being the only clear exception.
-General baroclinic instability of the upwelling front is a possible mechanism to drive offshore exchange \cite<e.g.>{ikedaetal84,durskiallen05}, but this tends to be shallow, with smaller-scale instabilities that will not extend as far into the interior ocean as observed here.
+It is possible that there is spatial locking of these features, with a persistent separation at the north tip of Vancouver Island, and a strong tendency for one at 49.5 N.
+There is  also has evidence of separation events in most of the years, with 2011 being the only clear exception.
+General baroclinic instability of the upwelling front is a possible mechanism to drive offshore exchange \cite{ikedaetal84,durskiallen05}, but this tends to be shallow, with smaller-scale instabilities that will not extend as far into the interior ocean as observed here.
 Rather it seems likely that the topographic change engendered by the sudden turn to the east of La Perouse bank catalyses a larger scale instability at this location.
-In \citeA{durskiallen05}, when including realistic shelf bathymetry including Heceta Bank, the region around the bank catalyzes intermittent large-scale instabilities similar to the feature here and along the coast \cite{batteen97}.
+\citeA{durskiallen05}, when modelling the Oregon shelf, found that including realistic shelf bathymetry catalyzed intermittent large-scale instabilities similar to the feature here and likely other coastal locations \cite{batteen97}.
 
 Large-scale mixing between the shelf and open ocean has been evident since the satellite era.
-Here we demonstrate that in this study area the flow is originating on the shelf and separating from the bathymetry and being injected into the interior down to the bottom of the water column.
-A  similar observation was made by \citeA{barthetal00} downstream of Cape Blanco, Oregon, where the coastal current was observed to detach from the shelf in the lee of the cape, and flow into the interior.
+Here we demonstrate that in the Vancouver Island shelf the flow is originating on the shelf and separating from the bathymetry and being injected into the interior.
+A  similar observation was made by \citeA{barthetal00} downstream of Cape Blanco, Oregon, where the coastal current was observed to detach from the shelf in the lee of the cape and flow into the interior.
 They hypothesized that as the current moved offshore, it deepened, stretching isopycnals and creating cyclonic relative vorticity that would tend to push the current back onshelf, but then it was caught in the undercurrent and stalled, being pushed offshore.
-It is also possible that coastally trapped waves in the region experience a hydraulic control, and these meanders are the response \cite{dalebarth01}.
+It is also possible that coastally trapped waves in the region experience a hydraulic control, and these sepearation events are part of the response \cite{dalebarth01}.
 
@@
 
-Regardless of the dynamics, the offshore transport is substantial.
+Regardless of the dynamics of the separation events, the offshore transport can be substantial.
 If we assume the coastal current is approximately $0.1\ \mathrm{m\,s^{-1}}$ over 100 m in the vertical and 20 km in the horizontal, it represents 0.2 Sv of nutrient- and chlorophyll-rich shelf water transported offshore.
 Our observations are a finer detailed representation of the kind of cross-shore transports inferred by \citeA{mackasyelland99} from hydrographic surveys, and definitively show that this water can originate from the shelf from relatively deep depths and be transported offshore.
-We do not have velocity measurements for the water that replaces it, but assuming that water also flows along-shelf, there is a large replacement of shelf water with offshore water at this location.
+We do not have velocity measurements for the water that replaces it, but assuming that water also flows along-shelf, these separation events are associated with a large replacement of shelf water with offshore water at this location.
 This emphasizes the importance of three dimensional observations and modeling of cross-shelf dynamics when thinking about biological processes on the shelf.
@@
 
-%Open Research
-% AGU requires an Availability Statement for the underlying data needed to understand, evaluate, and build upon the reported research at the time of peer review and publication.
-
-%Additionally, authors should include an Availability Statement for the software that has a significant impact on the research.
-Details and templates are in the Availability Statement section of the Data & Software for Authors Guidance:
-% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#availability
-
-%For physical samples, use the IGSN persistent identifier, see the International Geo Sample Numbers section:
-% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#IGSN
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Derived data files (1-m vertical binned CTD files) and analysis scripts are available at \url{https://ocean-physics.seos.uvic.ca/~jklymak/PW13DataAnalysis/}.
+Raw CTD data is available on request.
+Data was processed using a scientific python toolchain as listed at \url{https://ocean-physics.seos.uvic.ca/~jklymak/PW13DataAnalysis/environment.yml}; major
+components include xarray \cite{xarray}, numpy \cite{numpy}, and Matplotlib \cite{Matplotlib36}, but those all leverage many smaller but vital projects.
 
 \acknowledgments
-This section is optional. Include any Acknowledgments here.
+
+Thank you to the officers and crew of the R/V Falkor and the Schmidt Foundation for funding the seagoing component of the cruise.
+Thank you to Richard Dewey, Benjamin Schiefle, and Rowan Fox for efforts during the cruise.
+Klymak was supported by Canadian NSERC grant 327920-2006 and a Canadian Foundation for Innovation Leading Edge Fund grant 36109.
 

### 2023-01-24 PW13Paper.tex
Version to SA, SNW after revisions

@@
 
-Comprehensive water column observations of temperature, salinity, and oxygen on the productive southern Vancouver Island shelf offer a unique characterization of processes affecting exchange on the shelf.
-A semi-permanent cyclonic recirculation (the Juan de Fuca Eddy) occupies the region in the lee of a bank where the shelf widens abruptly.
-The observations indicate that the water in this Eddy is a mixture of offshore water and water from a buoyant coastal current.
-This water is well-mixed along a mixing line in temperature-salinity space, though it retains stratification, and is either rapidly mixed (high mixing rate $\kappa$), or has a long residence time (large $\tau$), such that the mixing length $(\kappa \tau)^{1/2} \sim 50$ ---  $100\ \mathrm{m}$.
-There is a sharp temperature-salinity front on the offshore side of this well-mixed water.
-The front is no more than 1-km wide, and has no sign of instabilities.
-Previous work had hypothesized the role of a spur canyon incising the shelf in feeding water to the Eddy; little evidence of this was found in the observations, however, these were collected during a relaxation and reversal of upwelling winds, which may have slowed the flow of water into the Eddy.
-Tidally resolved observations in the spur canyon also show very strong hydraulic flows in the cross-canyon direction accompanied by very high localized mixing rates in 50-m hydraulic jumps, possibly exceeding $\kappa \sim 1 \ \mathrm{m^2s^{-1}}$.
-
-Upstream of the Eddy, there is an along-shelf current flowing equatorward with partially mixed water properties compared to the well-mixed water in the Eddy.
-The shelf current was observed to separate from the shelf in the lee of a bank, and was replaced by a distinct offshore water mass.
-The separation event can be seen in sea-surface temperatures in satellite images as a tongue of water that is ejected offshore.
-The cause of this offshore separation event is not determined, but possibly is catalyzed by separation of the current from the abrupt underwater bank at the poleward end of the study region.
+Observations of temperature, salinity, and oxygen on the southern Vancouver Island shelf show a large-scale exchange of shelf water with offshore water, just offshore of a semi-permanent recirculation.
+The semi-permanent cyclonic recirculation (the Juan de Fuca Eddy) occupies a region where the shelf widens abruptly in the lee of a bank.
+The water in this Eddy is a mixture of offshore water and water from a buoyant coastal current.
+This water is well-mixed along a mixing line in temperature-salinity space, though it retains stratification, and is either rapidly mixed or has a long residence time.
+There is a sharp temperature-salinity front on the offshore side of this well-mixed water, no more than 1-km wide, and has no sign of instabilities.
+The clearest evidence of cross-front transport is found during a short tidally resolved survey over a bank, and was due to flows in the cross-bank direction driving large mixing in 50-m tall hydraulic jumps.
+Upstream of the recirculation there is an along-shelf current flowing equatorward.
+However the whole current separates from the shelf before reaching the recirculation, in the lee of a bank, and is replaced by water from offshore.
+The separation event was also seen in sea-surface temperatures from satellite images as a tongue of cool coastal water that is ejected offshore.
 
@@
 
-The southern Vancouver Island continental shelf is very biologically productive due to high nutrient input from the Strait of Juan de Fuca and Salish Sea estuarine system and substantial cross-shelf transport due to the complicated topography.
-The estuarine water flows poleward, hugging the coast as the Vancouver Island Coastal Current and, in the summer, shelf water flows equatorward.
-Trapped between these two currents is the Juan de Fuca Eddy, a stagnant region of water that occupies the shelf where it abruptly widens just north of the Strait of Juan de Fuca.
-
+The southern Vancouver Island continental shelf is biologically productive due to high nutrient input from the Strait of Juan de Fuca and Salish Sea estuarine system and substantial cross-shelf transport due to the complicated topography.
 Here we present intensive sampling of the Juan de Fuca Eddy region.
@@
 Such events are readily seen in satellite imagery, but our observations indicate that the separation extends the depth of the water column on the shelf, and that this separation may be partially driven by the local bathymetry.
-The separation is a very strong cross-shelf exchange event, and likely transports substantial nutrient-rich coastal water offshore to drive productivity in the deeper ocean adjacent to the continental slope.
+The separation is a very strong cross-shelf exchange event, and transports substantial nutrient-rich coastal water offshore to drive productivity in the deeper ocean adjacent to the continental slope.
 
@@
 Mechanisms of cross-shelf exchange include internal waves and instabilities in shelf-break fronts.
-However, these mechanisms are often smaller in magnitude to intermittent three-dimensional exchange, driven by eddying, and often catalyzed by topographic irregularity \cite{barthetal00}.
+However, these mechanisms can be smaller in magnitude compared to intermittent three-dimensional exchange, driven by eddying, often catalyzed by topographic irregularity \cite{barthetal00}.
 
 Here we present detailed \emph{in-situ} observations from the southern Vancouver Island shelf, collected in summer 2013, and sampled by a rapid profiling vehicle equipped with a CTD and oxygen sensor, supplemented by traditional hydrographic surveys bracketing the high resolution observations by a month.
-The shelf has complicated bathymetry (\fref{fig:LocMapBoth}), with a simple shelf geometry poleward of our study site, but widening in the lee of La Perouse Bank, and incised by the Juan de Fuca canyon, equatorward.
-Water from the Strait of Juan de Fuca flows poleward as a buoyant current that hugs the coast \cite{thomsonetal89, hickeyetal91}, while shelf water flows equatorward in the summer under upwelling conditions, both forced by local winds and via tele-connections with the long, homogenous shelves equatorward off Washington and Oregon \cite{hickeyetal91,thomsonkrassovski15,engidaetal16}.
+The shelf has complicated bathymetry (\fref{fig:LocMapBoth}), with a relatively simple 50-km wide shelf poleward of our study site, that widens to over 75 km wide equatorward of La Perouse Bank.
+The South Vancouver Island Shelf is this wide shelf region, characterized by a number of banks, and finally is incised on the poleward side by the Juan de Fuca canyon.
+Water from the Strait of Juan de Fuca flows poleward as a buoyant current that hugs the coast \cite{thomsonetal89, hickeyetal91}, while shelf water flows equatorward in the summer both forced by local winds and via teleconnections with the long, homogenous shelves equatorward off Washington and Oregon \cite{hickeyetal91,thomsonkrassovski15,engidaetal16}.
 Trapped between these two currents is a region of relatively homogenous water that has been termed the Juan de Fuca Eddy \cite{freelanddenman82,freelandmcintosh89,foremanetal07,macfadyenhickey10}, denoted ``\Eddy'' below.
@@
 A goal of our study was to understand how the \Eddy\ persists and how it exchanges water properties with offshore water.
-We focus on water deeper than $50 \mathrm{m}$, below the summer mixed layer, because this water is trackable with water mass properties, whereas the \Eddy\ is often studied from the point of view of the surface circulation \citep[e.g.][]{macfadyenhickey10}.
+We focus on water deeper than $50 \mathrm{m}$, below the summer mixed layer, because this water is trackable with water mass properties, even though the \Eddy\ is often studied from the point of view of the surface circulation \citep[e.g.][]{macfadyenhickey10}.
 Past studies have found low oxygen concentrations in the \Eddy, and inferred that water is upwelled from the California Undercurrent from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
-This inference was based on the assumption that oxygen is conservative, and hence water in the eddy had to come from  the oxygen minimum zone found further offshore \cite{mackasetal87}.
+This inference was based on the assumption that oxygen is conservative, and hence water in the \Eddy\ had to come from  the oxygen minimum zone found further offshore \cite{mackasetal87}.
 This lead to hypotheses that perhaps the \Eddy\ was fed by waters being drawn up a spur canyon (called the Spur Canyon) via ageostrophic transport from offshore to onshore due to low pressure in the \Eddy\ center \cite{weaverhsieh87}.
@@
 Cross-shore transport is supplied by Ekman layers in simple geometries under wind forcing.
-However, in many locations there is also evidence of filaments and wholesale separation of shelf currents into the deep ocean.
+However, in many locations there is also evidence of eddies and filaments driving wholesale separation of shelf currents into the deep ocean.
 Along the Vancouver Island shelf, large filaments or instabilities of the coastal current have been inferred using satellite images \cite{ikedaemery84,thomsongower98}.
@@
 Net cross-shore transport of nutrients and chlorophyll have been found in hydrographic surveys off Vancouver Island \cite{mackasyelland99}, and associated with mesoscale features in geostrophic velocities and in satellite observations.
-The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics leading to along-shore trapping of coastal waves \cite{dalebarth01},  wind stress curl variations \cite{castelaobarth07}, induced relative vorticity due to stretching of parcels that overshoot their initial isobaths due to a sudden change in downstream bathymetry \citep{dasaro88}, and the non-linear breaking of large-scale meanders due to baroclinic instability between the wind-driven current and the California Undercurrent \cite{ikedaetal84, batteen97}.
+The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics leading to along-shore trapping of coastal waves \cite{dalebarth01},  wind stress curl variations \cite{castelaobarth07}, induced relative vorticity due to stretching of parcels that overshoot their initial isobaths due to a sudden change in downstream bathymetry \citep{dasaro88}, and, at this location, non-linear breaking of large-scale meanders due to baroclinic instability between the wind-driven current and the California Undercurrent \cite{ikedaetal84, batteen97}.
 
@@
 
-Below, we focus on finescale surveys carried out from 2013-08-21 to 2013-08-30.
-Data were collected from the \emph{R/V Falkor} with an ODIM Brooke-Ocean Moving Vessel Profiler (MVP).
+We focus on finescale surveys carried out between these hydrographic surveys, from 2013-08-21 to 2013-08-30.
+Data were collected from the \emph{R/V Falkor} with an AML Oceanographic Moving Vessel Profiler (MVP).
+Data was collected analogously to data collected during similar field campaigns \cite{klymaketal15a, klymaketal16, dasaroetal18}.
 The MVP was equipped with an AML Oceanographic CTD, and a Rinko Oxygen sensor with a 7-s response time foil.
@@
 
-Winds during the cruises were typical for the west coast of Vancouver Island, with upwelling favorable winds during July and early August (\fref{fig:LaPeWind}).
+Winds during the cruises were typical for the west coast of Vancouver Island, with equatorward upwelling-favorable winds during July and early August (\fref{fig:LaPeWind}).
 During the finescale survey, and for the week previous, the winds were intermittently downwelling favorable.
@@
 
-Hydrographic sections along the LB and LC lines highlight summer conditions on the Southern Vancouver Island Shelf and indicate some of the features we are focusing on in this paper (\Fref{fig:LaPerouse2013Ctd}).
+Hydrographic sections along the LB and LC hydrographic lines highlight summer conditions on the Southern Vancouver Island Shelf and indicate some of the features we are focusing on in this paper (\Fref{fig:LaPerouse2013Ctd}).
 During both surveys, and along both lines, there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m depth offshore to shallower than 85 m over the shelf.
 This upwelled water tends to be low in oxygen and cool.
-Further onshore, the Vancouver Island Coastal Current hugs the coast, where isopycnals tilt back down and water is warmer than offshore.
+Further onshore, the Vancouver Island Coastal Current hugs the coast, where isopycnals tilt down towards the shore,and water is warmer than offshore.
 
@@
 It does not show as much drawdown of oxygen, and is warmer, at least offshore of La Perouse bank ($X \approx 5\ \mathrm{km}$).
-Based on these properties, offshore water does not appear to make it over La Perouse bank into the basin onshore, or if it does, it does so intermittently.
+Based on these properties, offshore water does not appear to make it over La Perouse bank into the onshore basin, or if it does, it does so intermittently and with substantial mixing.
 Note that the water along the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal becomes cooler and lower in oxygen where it intersects the shelf, consistent with both enhanced mixing near the shelf and with drawdown of oxygen by respiration.
@@
 
-We use a spice anomaly defined as relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}) representing an approximate mixing line for water in the \Eddy, and passsing through the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$.
+We use a spice anomaly defined as relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}) representing an approximate mixing line for water in the \Eddy, and passing through the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$.
 Spice anomaly for a given water sample at density $\sigma_{\theta}$ is given by $\gamma = \alpha \left(T - T_0\left(\sigma_{\theta}\right)\right) + \beta \left(S - S_0\left(\sigma_{\theta}\right)\right)$ where $T_0\left(\sigma_{\theta}\right)$, where $S_0\left(\sigma_{\theta}\right)$ are the temperature and salinity along the mixing line at the same density as the water sample.
@@
 Away from the surface, water has upwelled from deeper depths by September, but the offshore water maintains the same water mass characteristics through the summer.
-Hugging the coast ($X > 30\, \mathrm{km}$), the Vancouver Island Coastal Current warms during the summer, and the spice anomaly goes from negative to postive.
+Hugging the coast ($X > 30\, \mathrm{km}$), the Vancouver Island Coastal Current warms during the summer, and the spice anomaly goes from negative to positive.
 In the \Eddy, the water stays near the mixing line, but is cooler in the spring (\fref{fig:LaPerouse2013Ctd}, LB Line, left-hand column: slightly negative spice anomaly) and warms during the summer (\fref{fig:LaPerouse2013Ctd}, LB Line, right-hand column: slightly positive spice anomaly).
 
-Oxygen concentration in both surveys has a similar dichotomy between the \Eddy\ and offshelf water.
-Water found in the \Eddy\ has oxygen concentrations $100\ \mathrm{\mu mol\, kg^{-1}}$ lower on the shelf than offshelf (\fref{fig:LaPerouse2013Ctd}, \fref{fig:LaPerouse2013TS}b).
+Oxygen concentration in both surveys has a similar dichotomy between the \Eddy\ and off-shelf water.
+Water found in the \Eddy\ has oxygen concentrations $100\ \mathrm{\mu mol\, kg^{-1}}$ lower on the shelf than off-shelf (\fref{fig:LaPerouse2013Ctd}, \fref{fig:LaPerouse2013TS}b).
 The very deepest water ($S\approx 33.9\ \mathrm{psu}$) on the shelf shows a further $50\ \mathrm{\mu mol\, kg^{-1}}$ decrease in concetration between May and September along the LB line.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Finescale surveys}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsubsection{Overview}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 For water denser than $26\ \mathrm{kg\,m^{-3}}$, there are three distinct water masses sampled on the shelf (\fref{fig:TSdensSpice}).
-The first is the offshore water, which tends to be warmer, and hence has a positive spice anomaly ($\gamma \approx 0.2\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$).
+The first is offshore water, which tends to be warmer, and hence has a positive spice anomaly ($\gamma \approx 0.2\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$).
 The second is water in the \Eddy, which during this survey was found along the straight mixing line in $\theta-S$ space (\fref{fig:TSdensSpice}, $\gamma \approx 0\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$).
 Between these two water masses, there is a less populous mass (\fref{fig:TSdensSpice}, $\gamma \approx 0.1\ \mathrm{kg\,m^{-3}}$ along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$) that we demonstrate below is found on the shelf poleward of the \Eddy.
-
-% TODO: shallower?
 
@@
 In contrast, the third population of partially mixed water between the \Eddy\ and offshore water (\fref{fig:TSdensSpice}) is all found poleward of the \Eddy\ region (\fref{fig:CrossSectionsSpice}a, b) as water with a weaker spice anomaly ($\gamma \approx 0.1\ \mathrm{kg\,m^{-3}}$, pink colors).
-This water is generally found near the intersection of the upwelling water and La Perouse Bank, and is not as warm as offshore water, indicating some mixing with offshore water has taken place.
+This water is  not as warm as offshore water, indicating some mixing with offshore water has taken place.
 Of note is that this intermediate water mass appears completely absent in the sections further equatorward (\fref{fig:CrossSectionsSpice}c, d), indicating that it is either mixed away or that it is advected elsewhere; we argue below that it is advected offshore in a large-scale tongue.
@@
 
-Oxygen saturation sections show similar trends, with the low-spice anomaly water also having low oxygen saturation, while the strong negative spice anomaly shallower than $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ has very high oxygen saturation; deeper water is also more oxygenated.
-Of note is that compared with the coarser sampling along the LC Line (\fref{fig:LaPerouse2013Ctd}), we see that the abrupt transition between offshore and \Eddy\ water is also present in disolved oxygen.
+Oxygen saturation sections show that the deep water in the  \Eddy\ has very low oxygen saturation (\fref{fig:CrossSectionsO2}c, d) compared to surrounding water (though again, we caution against the quality of these saturation values from such a fast profiling instrument with a slow response time).
+More oxygenated water is found offshore, and the oxygen deficit is not as strong in the poleward sections (\fref{fig:CrossSectionsO2}a, b).
+The T/S compensated front also shows an abrupt transition from the \Eddy\ water and the offshore.
+Note the excellent correlation between oxygen saturation and spice anomaly in \fref{fig:CrossSectionsSpice}.
 
@@
 The spatial patterns are very clear when considering a map view of properties along the $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ isopycnal (\fref{fig:SpiceO2264}).
-There is a region onshore of the south tip of La Perouse Bank (approximately 44.5\textdegree N, 125.75\textdegree W), that consists of water that is found along the mixing line (spice anomaly $\gamma \approx 0\ \mathrm{kg\,m^{-3}}$) with low oxygen saturation.
+There is a region onshore of the south tip of La Perouse Bank (approximately 44.5\textdegree N, 125.75\textdegree W), that consists of water that is found along the mixing line (spice anomaly $\gamma \approx 0\ \mathrm{kg\,m^{-3}}$) that also has low oxygen saturation.
 Offshore of this region is water that is very warm and salty in comparison ($\gamma > 0.15 \mathrm{kg\,m^{-3}}$), and relatively high in oxygen (saturations of approximately 60\%).
@@
 \subsubsection{Spur Canyon}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Continuing across the \Eddy, isopycnals are largely flat until they intersect the Vancouver Island Coastal Current ($80\ \mathrm{km}$).
-The deeper isopycnals are not found in the deeper basin northeast of La Perouse bank ($70\ \mathrm{km}$),  thus the bank is a natural northern boundary of the \Eddy.
+The deeper isopycnals are not found in the deeper basin northeast of La Perouse bank ($70\ \mathrm{km}$),  thus the bank is a natural poleward boundary of the \Eddy.
 
@@
 Much of the modified water in the Spur Canyon appears to come from the shelf to the west, but heavily modified by tidal mixing.
-A repeated tidal survey over the bank on the west side of the canyon shows a strong hydraulic response during onshore flow (17:58--21:00, \fref{fig:HydraulicsCanyon}).
+A repeat tidal survey over the bank on the west side of the canyon shows a strong hydraulic response during onshore flow (17:58--21:00, \fref{fig:HydraulicsCanyon}).
 Dense water passes from the offshore side into the canyon, and plunges down the side wall before rebounding downstream.
 Note that the tide here is largely diurnal, so this onslope flow only occurs once a day.
-Given the stratification of $N\approx 6\times10^{-3} \ \mathrm{rad\,s^{-1}}$ and an overturning scale of 50 m, we might expect dissipation rates reaching $\epsilon \sim L^2N^{3} \approx 5\times10^{-4}\ \mathrm{m^2\,s^{-3}}$, which is more than three orders of magnitude higher than dissipation observed west of this location by \citeA{deweycrawford88}.
+Given the stratification of $N\approx 6\times10^{-3} \ \mathrm{rad\,s^{-1}}$ and an overturning scale of 50 m, we might expect dissipation rates reaching $\epsilon \sim L^2N^{3} \approx 5\times10^{-4}\ \mathrm{m^2\,s^{-3}}$, which is three orders of magnitude higher than dissipation observed on the shelf west of this location by \citeA{deweycrawford88}.
 This estimate of turbulence dissipation rate implies a diapycnal diffusivity of $\kappa = \gamma \epsilon / N^2 \approx 1\ \mathrm{m^2\,s^{-1}}$, assuming a mixing efficiency of $\gamma=0.2$.
@@
 The turbulence found on the canyon rim makes it ambiguous if there is water moving up the canyon or not.
-There is a general tendency along the canyon for higher spice water to be found offshore (\fref{fig:CanyonPropertiesSpice}) but it seems likely the source of the higher spice water is from over the bank rather than water being advected up the canyon.
+There is a general tendency along the canyon for higher spice water to be found offshore (\fref{fig:CanyonPropertiesSpice}) but it seems likely that the source of the higher spice water is from over the bank rather than water being advected up the canyon.
 This tidally driven flow over the bank is the most significant source of high-spice offshore water into the eddy region identified during our surveys.
@@
 \label{sec:frontsurvey}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Separation of coastal water}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 Upstream of the the sharp front between the low-spice \Eddy\ and the high-spice offshore water is a substantial mass of intermediate-spice water along the shelf (compare \fref{fig:CrossSectionsSpice}a and c).
-If in geostrophic balance, this intermediate spice water is moving equatorward along the shelf upstream of the \Eddy, but is pushed offshore just upstream of Barkley Canyon (\fref{fig:IsopycnalsDepth}).
+This intermediate spice water is moving equatorward along the shelf upstream of the \Eddy, but is pushed offshore just upstream of Barkley Canyon (\fref{fig:IsopycnalsDepth}).
 There is a tongue of intermediate-spice water ($\gamma$ is pink along $26.4\,\mathrm{kg\,m^{-3}}$) that separates from the shelf just west of 126\textdegree W.
@@
 There is a cooler tendril streaming west at 48 N off the south end of this tongue.
-This feature is not as well-developed in the previous image (21 August) perhaps indicating that it is an evolving feature.
+This feature is not as well-developed in the previous image (Aug.\ 21) perhaps indicating that it is an evolving feature.
 By 31 August, there is no surface expression of the feature, though small tendrils of cooler water can be seen separating from La Perouse Bank.
@@
 By September 6, we see the feature fading from the satellite image.
-Note that this feature is centered 0.2 degrees of latitude south of tongue that we observe deeper in the water column, again indicating that there is depth-dependen structure in the feature.
+Note that this feature is centered 0.2 degrees of latitude south of tongue that we observe deeper in the water column, again indicating that there is depth-dependent structure in the feature.
 
@@
 The water mass crosses isobaths and is ejected into the interior.
-This separation event can also be observed from satellite measurements of seasurface temperature and chlorophyll.
+This separation event can also be observed from satellite measurements of sea surface temperature and chlorophyll.
 Thus the water that is offshore of the \Eddy\ appears to have been brought onto the shelf in exchange for the offshore ejection of shelf water via this tongue.
@@
 The amount of turbulence required to homogenize 100 m of water over 90 days is $\kappa \approx 10^{-3}\ \mathrm{m^2\,s^{-1}}$.
-Given that the diffusivity implied in the cross-channel surveys was (very) locally on the order of $\kappa_{\rho} = 10 \ \mathrm{m^2\,s^{-1}}$, this number is not outrageous if we think that such high dissipation is found in $10^{-4}$ of the water column.
+Given that the diffusivity implied in the cross-channel surveys was on the order of $\kappa_{\rho} = 10 \ \mathrm{m^2\,s^{-1}}$, this number is not outrageous if we think that such high dissipation is found in $10^{-4}$ of the water column.
 We can more carefully quantify this by considering a synthetic profile or temperature and salinity based on an offshore profile, extrapolated from the bottom of the 200-m cast to 250 m, assuming that the temperature and salinity at 250 m are $6.2\ \mathrm{[^oC]}$ and $34\ \mathrm{psu}$ respectively (\fref{fig:TSExercise}).
@@
 
-Numerical simulations of this region reported by \cite{sahuetal22} using a NEMO 36th-degree regional model  indicate that there is more rapid exchange between the deep \Eddy\ water and the rest of the coastal ocean.
-In this model, the region where the \Eddy\ resides has velocities equal or greater than other parts of the shelf, and water has an approximate residence time of less than 20 days.
+Numerical simulations of this region reported by \cite{sahuetal22} using a NEMO 36th-degree regional model contained relatively rapid exchange between the deep \Eddy\ water and the rest of the coastal ocean.
+The region where the \Eddy\ resides has velocities equal or greater than other parts of the shelf, and water has an approximate residence time of less than 20 days.
 The \Eddy\ water in the model does develop a distinct $\theta$--$S$ signature, but not along a sharp mixing line as observed.
@@
 
-The (apparent) displacement of shelf water from La Perouse Bank is a dramatic departure from geostrophically balanced isobath-following flow.
+The displacement of shelf water from La Perouse Bank is a dramatic departure from geostrophically balanced isobath-following flow.
 Eddies have been known to separate from irregular coastal topography, both at the surface \cite{barthetal00} and deeper in the water column \cite{pellandetal13}.
@@
 However, observations of the wholesale replacement of shelf water by a new water mass from offshore are rare.
-In these observations it is clear that water from as deep as 150 m is separating from the shelf and moving offshore (\fref{fig:SpiceO2264}, \fref{fig:IsopycnalsDepth}).
+In the  observations presented here, it is clear that water from as deep as 150 m is separating from the shelf and moving offshore (\fref{fig:SpiceO2264}, \fref{fig:IsopycnalsDepth}).
 
-Satellite imagery makes it clear that there is often exchange between coastal and deep waters along the Vancouver Island shelf (\fref{fig:SSTLateAug}).
+Satellite imagery shows that there is often exchange between coastal and deep waters along the Vancouver Island shelf (\fref{fig:SSTLateAug}).
 Most years there are three of four large filaments from the shelf into the open ocean, many of them over 100 km long.
@@
 It is possible that there is spatial locking of these features, with a persistent separation at the north tip of Vancouver Island, and a strong tendency for one at 49.5 N.
-There is  also has evidence of separation events in most of the years, with 2011 being the only clear exception.
+There is also evidence of separation events in most of the years, with 2011 being the only clear exception.
 General baroclinic instability of the upwelling front is a possible mechanism to drive offshore exchange \cite{ikedaetal84,durskiallen05}, but this tends to be shallow, with smaller-scale instabilities that will not extend as far into the interior ocean as observed here.
 Rather it seems likely that the topographic change engendered by the sudden turn to the east of La Perouse bank catalyses a larger scale instability at this location.
-\citeA{durskiallen05}, when modelling the Oregon shelf, found that including realistic shelf bathymetry catalyzed intermittent large-scale instabilities similar to the feature here and likely other coastal locations \cite{batteen97}.
+\citeA{durskiallen05}, when modelling the Oregon shelf, found that including realistic shelf bathymetry catalyzed intermittent large-scale instabilities similar to the feature here, a finding that is deemed likely to apply at other coastal locations \cite{batteen97}.
 
@@
 They hypothesized that as the current moved offshore, it deepened, stretching isopycnals and creating cyclonic relative vorticity that would tend to push the current back onshelf, but then it was caught in the undercurrent and stalled, being pushed offshore.
-It is also possible that coastally trapped waves in the region experience a hydraulic control, and these sepearation events are part of the response \cite{dalebarth01}.
+It is also possible that coastally trapped waves in the region experience a hydraulic control, and these separation events are part of the response \cite{dalebarth01}.
 
@@
 We do not have velocity measurements for the water that replaces it, but assuming that water also flows along-shelf, these separation events are associated with a large replacement of shelf water with offshore water at this location.
-This emphasizes the importance of three dimensional observations and modeling of cross-shelf dynamics when thinking about biological processes on the shelf.
+This emphasizes the importance of three dimensional observations and modeling of cross-shelf dynamics when thinking about physical and biological processes on the shelf.
 
@@
 Thank you to Richard Dewey, Benjamin Schiefle, and Rowan Fox for efforts during the cruise.
-Klymak was supported by Canadian NSERC grant 327920-2006 and a Canadian Foundation for Innovation Leading Edge Fund grant 36109.
-
-%% ------------------------------------------------------------------------ %%
-%% References and Citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% \bibliography{<name of your .bib file>} don't specify the file extension
-%
-% don't specify bibliographystyle
-
-% In the References section, cite the data/software described in the Availability Statement (this includes primary and processed data used for your research).
-For details on data/software citation as well as examples, see the Data & Software Citation section of the Data & Software for Authors guidance
-% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#citation
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Klymak was supported by NSERC Discovery grant RGPIN-2017-04050 and a Canadian Foundation for Innovation Leading Edge Fund grant 36109, Allen was supported by NSERC Discovery grant RGPIN-105694-11, and Waterman was supported by NSERC Discovery grant RGPIN-2020-05799.
 
@@
 
-%Reference citation instructions and examples:
-%
-% Please use ONLY \cite and \citeA for reference citations.
-% \cite for parenthetical references
-% ...as shown in recent studies (Simpson et al., 2019)
-% \citeA for in-text citations
-% ...Simpson et al. (2019) have shown...
-%
-%
-%...as shown by \citeA{jskilby}.
-%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
-%...has been shown \cite{jskilbye}.
-%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
-%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
-%...has been shown by \cite <e.g.,>[and others]{lewin76}.
-%
-% apacite uses < > for prenotes and [ ] for postnotes
-% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \citealp, etc.).
-% \nocite is okay to use to add references from your Supporting Information
-%
-
 \end{document}
-
-More Information and Advice:
-
-%% ------------------------------------------------------------------------ %%
-%
-%  SECTION HEADS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Capitalize the first letter of each word (except for
-% prepositions, conjunctions, and articles that are
-% three or fewer letters).
-
-% AGU follows standard outline style; therefore, there cannot be a section 1 without
-% a section 2, or a section 2.3.1 without a section 2.3.2.
-% Please make sure your section numbers are balanced.
-% ---------------
-% Level 1 head
-%
-% Use the \section{} command to identify level 1 heads;
-% type the appropriate head wording between the curly
-% brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-% ---------------
-% Level 2 head
-%
-% Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-% ---------------
-% Level 3 head
-%
-% Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-% Level 4 head
-%
-% Use the \subsubsubsection{} command to identify level 3 heads
-% An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  IN-TEXT LISTS
-%
-%% ------------------------------------------------------------------------ %%
-%
-% Do not use bulleted lists; enumerated lists are okay.
-% \begin{enumerate}
-% \item
-% \item
-% \item
-% \end{enumerate}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Single-line equations are centered.
-% Equation arrays will appear left-aligned.
-
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
-
-% To create multiline equations, use the
-% \begin{eqnarray} and \end{eqnarray} environment
-% as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-
-% Break each line at a sign of operation
-% (+, -, etc.) if possible, with the sign of operation
-% on the new line.
-
-% Indent second and subsequent lines to align with
-% the first character following the equal sign on the
-% first line.
-
-% Use an \hspace{} command to insert horizontal space
-% into your equation if necessary. Place an appropriate
-% unit of measure between the curly braces, e.g.
-% \hspace{1in}; you may have to experiment to achieve
-% the correct amount of space.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATION NUMBERING: COUNTER
-%
-%% ------------------------------------------------------------------------ %%
-
-% You may change equation numbering by resetting
-% the equation counter or by explicitly numbering
-% an equation.
-
-% To explicitly number an equation, type \eqnum{}
-% (with the desired number between the brackets)
-% after the \begin{equation} or \begin{eqnarray}
-% command.  The \eqnum{} command will affect only
-% the equation it appears with; LaTeX will number
-% any equations appearing later in the manuscript
-% according to the equation counter.
-%
-
-% If you have a multiline equation that needs only
-% one equation number, use a \nonumber command in
-% front of the double backslashes (\\) as shown in
-% the multiline equation above.
-
-% If you are using line numbers, remember to surround
-% equations with \begin{linenomath*}...\end{linenomath*}
-
-%  To add line numbers to lines in equations:
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}

### 2023-05-02 R1/ResponseReviewers.tex
R1

% LaTeX rebuttal letter example.
%
% Copyright 2019 Friedemann Zenke, zenkelab.org
%
% Based on examples by Dirk Eddelbuettel, Fran and others from
% https://tex.stackexchange.com/questions/2317/latex-style-or-macro-for-detailed-response-to-referee-report
%
% Licensed under cc by-sa 3.0 with attribution required.

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage{listings}

% import Eq and Section references from the main manuscript where needed
% \usepackage{xr}
% \externaldocument{manuscript}

% package needed for optional arguments
\usepackage{xifthen}
% define counters for reviewers and their points
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

% This refines the format of how the reviewer/point reference will appear.
\renewcommand{\thepoint}{P\,\thereviewer.\arabic{point}}

% command declarations for reviewer points and our responses
\newcommand{\reviewersection}{\stepcounter{reviewer} \bigskip \hrule
                  \section*{Reviewer \thereviewer}}

\newenvironment{point}
   {\refstepcounter{point} \bigskip \noindent {\textbf{Reviewer~Point~\thepoint} } ---\ }
   {\par }

\newcommand{\shortpoint}[1]{\refstepcounter{point}  \bigskip \noindent
	{\textbf{Reviewer~Point~\thepoint} } ---~#1\par }

\newenvironment{reply}
   {\medskip \noindent \begin{sf}\textbf{Reply}:\  }
   {\medskip \end{sf}}

\newcommand{\shortreply}[2][]{\medskip \noindent \begin{sf}\textbf{Reply}:\  #2
	\ifthenelse{\equal{#1}{}}{}{ \hfill \footnotesize (#1)}%
	\medskip \end{sf}}

\newenvironment{changes}[1]
{\texttt{#1}}
{\par}

\lstloadlanguages{TeX}

\title{Response to reviewers: Separation of an upwelling current bounding the Juan de Fuca Eddy}
\author{Jody Klymak, Susan Allen, and Stephanie Waterman}

\begin{document}

\lstset{language=TeX}

\maketitle

\section*{}
% General intro text goes here
Thank you to both reviewers and the Editor for their time and effort into this paper.  It is appreciated that this is a complicated story in this paper, and the reviewers' persistence in making sense of it with us have been very helpful in adding both better insights and (hopefully) better explanations.

% Let's start point-by-point with Reviewer 1
\reviewersection

% Point one description
\subsection*{Reviewer summary}
The authors present a detailed analysis of some very high spatial resolution surveys of the shelf region off southern Vancouver Island. It is gratifying to see this nice analysis of this data set. The analysis focuses on how shelf circulation and water masses are related to an important mesoscale circulation feature in this region, namely the Juan de Fuca Eddy (EDDY). The authors present new data and explore mechanisms associated with mixing between shelf waters and the EDDY, document the sharpness of the front between the EDDY and offshore shelf waters, and describe how upstream shelf water separates from the shelf before reaching, and influencing, the EDDY. Lastly, they provide good evidence that low oxygen levels in the EDDY are likely due to in situ respiration rather than advection of low-oxygen, deep undercurrent water into the eddy as previously suggested. While I have some suggestions for improving the manuscript and its clarity, I recommend that this paper be published in the Journal of Geophysical Research after minor revision.

\begin{reply}
    Thank you for your comments
\end{reply}

\subsection*{Suggestions to improve the manuscript}

\begin{point}
    The idea that the equatorward shelf current "separation is a very strong cross-shelf exchange event, and transports substantial nutrient-rich coastal water offshore to drive productivity in the deeper ocean" is like documented elsewhere in the California Current, for example as summarized by Strub et al. (1991). Maybe refer to that study?
\end{point}

\begin{reply}
    That reference is super helpful; the lead author is embarrassed to say that he had it in his bibliography, but failed to re-consider it when writing the paper.  Not only should we have cited it, but it summarizes a lot of the previous work off California, and it lays out the likely mechanisms of the separation we saw.  The fourth paragraph of the introduction has been revised to include this reference.

    \begin{lstlisting}[language=TeX, basicstyle=\small]
        A second goal of our study was to better understand cross-shore exchange
        between shelf and offshore water.  For a two-dimensional topography with
        wind forcing, cross-shore transport is supplied by Ekman layers. However,
        in many locations there is also evidence of eddies, meanders and filaments
        driving wholesale separation of shelf currents into the deep ocean. These
        have been well-studied off California \cite{strubetal91} where satellite
        and \emph{in-situ} observations show large meanders of coastal upwelling
        currents leading to coastal water injected into the offshore domain.
        Along the...
    \end{lstlisting}

    Further, in the discussion we have added reference to this paper again.

    \begin{lstlisting}[language=TeX, basicstyle=\small]
        In California, most of the cold filaments observed appear to be
        catalyzed by headlands and underwater topography \cite{strubetal91},
        though modelling studies find instabilities are possible even in
        two-dimensional flows \cite{pierceetal91}.
    \end{lstlisting}

\end{reply}

\begin{point}
    Lines 202-210: Water masses along 26.4 not clear in Fig 5; what are we supposed to see in the "\# of data points per bin" color map? I think you are highlighting the low \# of data points (more purple) between spiciness anomaly ~0.02-0.1 in between the populous offshore water (spiciness ~0.2) and EDDY water (spiciness 0). Maybe draw a label and arrow to this in between region?
\end{point}

\begin{reply}
    Yes, there are three relatively distinct water masses along 26.4 - note the logarithmic color mapping.  To (hopefully) make this more clear we added a second panel to Figure 5 that shows the PDF along 26.4, and labeled the peaks.  Hopefully this helps clarify these water masses (in addition to the maps show in Figures 6 and 8).
\end{reply}

\begin{point}
    Line 211 and on: I don't see synthetic across-shelf lines in Fig 1b, but I do see them in Fig 8; need to describe the projection method to make the synthetic sections somewhere in the paper
\end{point}

\begin{reply}
    Ooops - Figure 1 lost its synthetic sections at some point.  They have been re-added in figure 1a (1b was getting quite crowded).

    The naive projection used is described now in the Site and Methods:

    \begin{lstlisting}[language=TeX, basicstyle=\small]
    We use an along/across-shelf co-ordinate system with an origin at
    48.4\textdegree N and 125.6\textdegree W, with the across-shelf $x$
    oriented 39 degrees north of East (\fref{fig:LocMapBoth}b, grey/white
    alternating lines).  The projection is Cartesian with a central latitude at
    48.4\textdegree N, which is fine for the limited geographic extents
    discussed here.
    \end{lstlisting}

\end{reply}

\begin{point}
    Lines 260-272: can we see blobs of high spice as described in Fig 10 in the along-canyon transects of Fig 9, say at X=27 km in the 8/24 section? Might be nice to point these out.
\end{point}

\begin{reply}
  Fig 10 is at 15 km in Fig 9, so we think the overflow feeds the continuum at this point in the canyon.  However, it is interesting to consider how blobs are getting as far up the canyon as 27 km.
\end{reply}

\begin{point}
    3.3 Separation of coastal waters: I know that ADCP data are not available, but how about computing geostrophic velocities from the density data?
\end{point}

\begin{reply}
    The geostrophic shear indicates equatorward flow, as evidenced by the onshore upwards tilt of the isopycnals.  Absolute velocities are of course hard to infer.

    However, more to the point, inferring relatively subtle offshore velocities using along-isobath density gradients is challenging.   For a geostrophically balanced anti-cyclonic circulation, we would expect a sea-surface low at the pivot point between the offshore moving water, and the onshore moving water.  That would lead to a local high in the isopycnal heights. Considering Fig.\ 12, which has this information, we think its pretty hard to see that occurring against the larger-scale backdrop of the isopycnals sloping up onto the shelf, and with the complication of internal tide heaving, which is considerable at this location.  Similarly considering Fig 6, it would be hard to say the 26.4 isopycnal is much shallower at 28 km than 42 km.  "Along isobath" also is pretty hard to define in this location.  Frankly its not even clear how much ADCP data we would have needed to collect to capture this feature adequately.
\end{reply}

\begin{point}
    Fig 12 and discussion: I'm not sure the 25.5 sigma-t maps help much. The ~0 spiciness anomaly EDDY water is not apparent inshore; the inshore water is actually spicier than the tongue water, likely because this density surface is being influenced by shallow mixed layer processes.
\end{point}

\begin{reply}
    We agree with your interpretation that this layer is being influenced by the surface mixed layer.  The layer was included for completeness.

    In response to this comment, we have changed the isopycnal to a deeper one (25.8) which is in the "minty" water.  We think it shows the three-dimensionality pretty well.

    Note there was a rounding error in the labelling of the bottom row - this isopycnal was 26.55, not 26.6 as indicated in the title and text.

    Finally we decided to add all the data, rather than a subset to this plot.  There were some subsequent surveys excluded from the first that help solidify the case that the tongue is there.
\end{reply}

\begin{point}
    Line 357-359: this conjecture about the source of the shelf waters offshore of the EDDY isn't supported by the evidence. It likely is just deeper upwelled water, southward of where the tongue separates. So upwelling drives it, not compensation for offshore flux of tongue water.
\end{point}

\begin{reply}
    Agreed that there is upwelling, and the upwelling tends to push water up the shelf.  However, its not clear why that would be stronger here than further upstream (poleward) where the water has had time to partially mix.  Further, simple upwelling ageostrophically brings water up at the bottom, whereas this is happening mid-depth.  If anything, you'd expect the secondary circulation associated with upwelling to push mid-depth water offshore?
\end{reply}

\begin{point}
    Section 4.1, lines 385-397: so the inferred oxygen consumption rate is less than the estimated rates based on other continental shelf studies. [Aside: see Connolly et al., 2010, for oxygen consumption rates on a nearby shelf rather than from the Gulf of Mexico.] So that implies that physical processes are keeping the oxygen in the EDDY higher than would be expected from biological decay and oxygen consumption. The authors remark on this by suggesting that the residence time of the water is shorter than their estimate of the whole season. Another way to say this, and as done in Adams et al. (2013), the physics (advection, mixing) contributes about half to the measured DO change. Adams et al. (2013) found about 57-68\% was due to physics. I suggest the authors connect the dramatic over-bank hydraulic jump and mixing mechanism they document as another physical process that can contribute to maintaining oxygen levels in the EDDY.
\end{point}

\begin{reply}
    Thanks a lot for these references - they are very helpful.  We reference Connolly et al when we quote the consumption rate (it was surprisingly hard to find references on this).  We have modified the end of this section to say the following, which we think is in the spirit of your suggestion:

    \begin{lstlisting}[language=TeX, basicstyle=\small]
        This is consistent with findings on the Oregon shelf where physical
        processes are thought to account for 55-70\% of changes in dissolved
        oxygen concentrations \cite{adamsetal13}.  Definitively identifying the
        exchange mechanisms into the \Eddy\ is challenging from this data set.
        The offshore front does not appear to have much exchange, however as
        noted above there does appear to be strong tidal flows and mixing over
        the submarine banks that can transport more oxygenated water into the
        \Eddy.  Further, the Vancouver Island coastal current is oxygen rich,
        and has a much less sharp front than the offshore front
        (\fref{fig:Frontsurvey}).
    \end{lstlisting}

\end{reply}

\begin{point}
    Section 4.1, lines 398-414: This is a useful argument,
    but it could made clearer. First, I don't see where the estimate of $\kappa_\rho = 10\ m^2/s$ is reported in the
    results section. Second, what is meant by "found in $10^{-4}$ of the water column?" Perhaps state a percentage instead. Lastly, the conclusion is that the vigorous mixing contributes to the homogenization of the EDDY water onto a mixing line, rather than the alternative long residence time (with weaker mixing). This is consistent with the paragraph above based on oxygen, so maybe useful to mention this corroboration.
\end{point}

\begin{reply}
    Ooops sorry - we decided to be more conservative in the results section, but left the inconsistency here.  In both sections we've changed the numbers to reflect a relatively broad range $0.5-5 \mathrm{m^2\,s^{-1}}$, given the looseness of any of these estimates, and we changed the fractions to percentages.
\end{reply}

\begin{point}
    Section 4.1, lines 420-426: so what shall we conclude? Likely that model is too viscous, related to the coarse horizontal resolution of 3 km. This is high spatial resolution, but evidently not enough.
\end{point}

\begin{reply}
Fair enough.  Added a concluding sentence:

\begin{lstlisting}
    Overall it seems that the model sees stronger cross-shelf advection than are
    apparent in these observations, though the reason for this will require
    further study.
\end{lstlisting}

Its hard to say more, but we suspect that the model has a stronger mesoscale eddy field than observed, and these generate substantial cross-shore motions.  It would be a great project to try and test why the regional simulation is so much more advective, but beyond our scope here.
\end{reply}

\subsection*{Minor/editorial comments:}

\begin{point}
Abstract: use of "recirculation" is odd. Recirculating with respect to what? Why not just use the term "eddy?"
\end{point}

\begin{reply}
    This was just to differentiate the dynamics a bit, but we don't feel strongly about it.  A stagnant region between two currents is eddy-like, but the analogy with a freely propagating eddy seems strained.  Regardless, we use recirculation once now in the abstract, and then drop it for "eddy" since that is how this feature is referred to in previous literature.
\end{reply}

\shortpoint{Figure 1: draw notional Juan de Fuca Eddy on here like for Vancouver Island Coastal Current and the coastal upwelling current}
\shortreply[]{Figure 1 has been modified accordingly.}

\shortpoint{Line 234: 48.5N, not 44.5N}
\shortreply[]{Thanks}

\shortpoint{Fig 12: label Barclay Canyon}
\shortreply[]{We tried this, but its pretty hard to do without obscuring data. Barkley is labelled in Figure 1.  }

\shortpoint{References: many uncapitalized place names (e.g., U.S. states)}
\shortreply[]{Thanks}

\begin{reply}
    Thank you again for your help with the manuscript.
\end{reply}

\reviewersection

\begin{point}
Overall, this is a very compact manuscript that documents the structure of an upwelling current and coastal jet that comprise the bounding arms of the Juan de Fuca Eddy, and contrasts T/S and oxygen characteristics inside and outside of the feature to infer the origin of eddy interior water. The authors have a really cool combination of datasets that they have wrung out to explore the eddy feature from multiple angles, but the introduction to these datasets and features is a bit too compact for readers who may not be familiar with the region, its geography, and major features. The text and figures were at times difficult to follow, and some suggestions are given below for mostly minor revisions that would may improve the clarity of the manuscript, interpretation of its figures, and overall readability.
\end{point}

\begin{point}
    I spent a LOT of time trying to translate the descriptions of the data sets and key geographic features to the multiple cruises, cross-sections, and coordinate systems. It was very difficult to compare the different lines and coordinate axes to each other, and the text switched between primary references to lat/lon, cruise lines, position relative to cross-shore coordinate axes, and geographic features. Figure 1b attempted to put all of the La Perouse, Falkor, and MVP casts together with the primary cross-shore coordinate system used in several plots, but several of the layers got lost, so it was difficult to distinguish them from each other. Perhaps rearranging the layering so that the MVP lines are at the bottom, and the coordinate system on top would be sufficient, and perhaps moving (or repeating) the coordinate system to Figure 2a would make the plot more readable. Is the buoy plotted here too? I'm not very familiar with the area, and often had to flip among multiple figures to follow the descriptions of the features in the text. If the location of a key feature like a canyon or a bank is essential to understanding the description of a feature, it would be helpful to make sure it is easy to find in Figure 1 and/or the figure being described (for example, text describing Figure 8).
\end{point}

\begin{reply}
Indeed, keeping this all straight is challenging, and thanks for your suggestions.  Figure 1 now has three panels, and we changed the layering.  This is probably worth the extra space, and we hope that it helps with legibility.  Added the La Perouse Buoy to Fig 1c.  Hopefully this helps, though with such a complicated study region it will remain challenging.
\end{reply}

\begin{point}

    The text is largely well-written, but some mostly stylistic choices make sentences difficult to parse. Avoid the construction of starting sentences with "This", as it leads to loss of specificity. For example, 78-83, in which "This" could refer to the inference, the assumption, the OMZ, etc. Or 250: "this" could mean ageostrophy or movement. Other examples where the wording is unclear are given in the comments and questions below.
\end{point}

\begin{reply}
    Thanks cleaned this up in both these spots, and a few other places.  Agree that using "this" at the beginning of a sentence should be clear and unambiguous.  On the other hand, repeating a phrase from the previous sentence (or worse rephrasing it) is harder to parse.
\end{reply}

\begin{point}

    Editing suggestions: Consistent capitalization of EDDY, Canyon, Figure, Bank, etc. Revision can also correct several common grammar errors (farther vs further, that vs which).
\end{point}

\begin{reply}
    Thanks for the careful read of this.  We think we have made the location
    names consistent now.  Note that only "Eddy" gets capitalized by itself;
    for the banks and canyons, we capitalized if it was the full name, and left
    lower case if referring to "the canyon".

    As to the grammar points, we are not grammar experts, but wonder if the
    further/farther distinction is an Americanism.  Our five minutes of
    googling this subject indicates that both are in common usage to describe
    distances.  (The lead author will admit thinking he was illiterate between moves from Canada, to the US, and back again in his spelling of ``modeling/modelling'')
\end{reply}

\begin{point}
    Smaller notes:
    23-27: some grammar+clarity issues
\end{point}

\begin{reply}
    Thanks changed to:
    \begin{lstlisting}
        There is a less than 1-km wide temperature-salinity front on the
        offshore side of this well-mixed water that has no sign of
        instabilities. The clearest evidence of cross-front transport is found
        during a tidally resolved survey over a bank. The transport is due to
        flows in the cross-bank direction that also drive 50-m tall hydraulic
        jumps.
    \end{lstlisting}
\end{reply}

\shortpoint{61: if I understand the intent of this sentence, no comma between
    site, that}

\shortreply[]{Fixed}

\shortpoint{ 63: equatorward, right?}

\shortreply[]{ooops, yes - mistakes like that won't make it easier for readers to get oriented.  }

\begin{point}
    79: is the oxygen conservative, or do you mean the process that upwells it preserves its signature?
\end{point}

\begin{reply}
   The assumption in the previous work was that oxygen is a conservative tracer.
\end{reply}

\begin{point}
    102-107: very long sentence
\end{point}

\begin{reply}
    Indeed - split.  The second sentence is still rather long, but is a list, so should be easy enough to parse, but agree that two introductory clauses and a list was very hard to read.
\end{reply}

\begin{point}
    116: which not that
\end{point}

\begin{reply}
    Went the other way and removed the comma.
\end{reply}

\begin{point}
    126-127: error bars for oxygen post Winklers/calibration?
\end{point}

\begin{reply}
    TODO: get Whitney reference.
\end{reply}

\begin{point}
    138-140: a real shame, as is the lack of the ADCP! The sentence about the ADCP can likely be combined with paragraph above.
\end{point}

\begin{point}
    153-154: which features? We haven't described them yet.
\end{point}

\begin{reply}
    Rephrased to:

    \begin{lstlisting}
        Hydrographic sections along the LB and LC hydrographic lines highlight
        summer conditions on the Southern Vancouver Island Shelf and presented
        to give context to the more detailed survey carried out on the
        \emph{R/V Falkor}
    \end{lstlisting}

\end{reply}

\begin{point}
    154: drawdown? What do you mean here?
\end{point}

\begin{reply}
    rephrased to

    \begin{lstlisting}
        The water upstream of the \Eddy\ region, along the LC line, is less
        mixed than the \Eddy\ water (\fref{fig:LaPerouse2013Ctd}, top three
        rows) and has higher oxygen.
    \end{lstlisting}

\end{reply}

\begin{point}

    165: the coordinate system is not described in the main text, just in the figure caption. Probably worth adding to section 2? Also helpful to reference the 26.4 kg/m3 isopycnal as magenta in the text here, especially if it is marked in magenta the rest of the paper.
\end{point}

\begin{reply}
    Thanks, there is a new second paragraph in the intro explaining the co-ordinate system.
\end{reply}

\begin{point}
    187: values can be easier to follow than color names (white is okay but defining what "pink" gets harder later on, after the descriptions of Figs 6-8)
\end{point}

\begin{reply}
  Thanks, great suggestion.
\end{reply}

\begin{point}

    Figure 5: a bit more info would help the reader find the features in the plot. In a talk, you might draw a circle around the offshore, eddy, and other water masses, but that's harder to do in a manuscript. Would having X mark the spot for a given feature, highlighting select profiles or using variance ellipses in T-S space of the features in question help?
\end{point}

\begin{reply}
    The other reviewer had a similar suggestion - we have instead made a PDF of $gamma$ along 26.4, and identified the peaks.  Readers may be more comfortable with 1-D PDFs anyhow, so this seems a good way to get this across.
\end{reply}

\begin{point}
    211: 2-d interpolation?
\end{point}

\begin{reply}
    We hadn't defined the interpolation, so added a sentence:
    \begin{lstlisting}
        These lines are made with a two-dimensional interpolation where MVP
        casts at a given depth have a Gaussian weight $w = \exp^{-(r/r_0)^2}$,
        where $r$ is the distance from the line and $r_0 = 1.5\ \mathrm{km}$.
    \end{lstlisting}
\end{reply}

\begin{point}
    Figure 6: a is close to LC, right? And d to LB? Probably good to make this connection explicitly for comparison to the La Perouse cruise data. A subpanel of T/S for selected profiles or subsets of these data would be interesting for comparison to Figs. 4/5. Might be interesting to color-code by oxygen to really drive home the water mass relationships.
\end{point}

\begin{reply}
    b is actually close to LC and d to LB.  Thanks, this is a useful distinction to make.  We have added these labels to Fig 6 and 7, and note this in the caption.

    We could easily add another subpanel for with O2 color coded.  We hesitate to do so because there are already quite a few presentation methods, and we think the O2 relationship is pretty clear in the other presentation methods.
\end{reply}

\begin{point}
    230: would be better to quantify correlation/significance
\end{point}

\begin{reply}
    We didn't mean to imply linear correlation - a linear correlation could be fit, but its clearly not along a straight line.  This inspired adding Fig 8c. which is a nice way to show this.
\end{reply}

\begin{point}
    234: 48.5N?
\end{point}

\begin{reply}
    Thanks, fixed.
\end{reply}

\begin{point}
    237: wording: regions are not abrupt
\end{point}

\begin{reply}
    Thanks, changed to "transition".
\end{reply}

\begin{point}
    248: citation?
\end{point}

\begin{reply}
    Thanks, added Mackas et al 1987 and Weaver and Hsieh 1987.
\end{reply}

\begin{point}
    Figure 9: Give the coordinate another name ($X_C$), since X/Y are used for the cross-shore coordinates.
\end{point}

\begin{reply}
    Thanks, done.
\end{reply}

\begin{point}

    283-286: this description is hard to follow. It looks to me like the surface $\gamma$ is greater than 0 everywhere, even at the surface, where the buoyant front should sit.
\end{point}

\begin{reply}
    We should not have called the whole thing a ``front''.  The current itself has a slightly negative spice anomaly (cool and fresh), but the front has a positive spice anomaly.  This is changed to

    \begin{lstlisting}
        The survey started close to shore, and passed through the
        Vancouver Island Coastal Current (along-track 0-10 km).
        The coastal current forms a buoyant current, and is fresher
        and colder than water at the same density.  The front with
        the coastal water is relatively thick, greater than 20 km
        wide, and has entrained partially mixed water down from the
        surface to the foot of the front ($\gamma\approx0.1$).
    \end{lstlisting}
\end{reply}

\begin{point}

    291: tendrils of what?
\end{point}

\begin{reply}
    Thanks "higher spice water" is now added.
\end{reply}

\begin{point}
    303-304: Is the distribution indicative of substantial isopycnal and vertical mixing, or rather are the sharp cut-offs in $\theta-S$ properties what is indicative of mixing?
\end{point}

\begin{reply}
    The sharp distinction of t-s properties are indicative of mixing. Changed to ``These distinct $\theta$--$S$ properties''.
\end{reply}

\begin{point}
    313-315, 316-318: Neither of these features is apparent to the reader (at least this one!). I see pink at 25.5 and 26.6 kg/m3 but not 26.4 kg/m3.
\end{point}

\begin{reply}
    That is unfortunate because it is one of the main points of the paper!    The new version of Fig 12 has all the data, which we probably should have used in the original submission because hopefully it makes the tongue more clear.  We also now refer to the panel explicitly.  We could annotate it, but have refrained for the present.
\end{reply}

\begin{point}
    Figure 11: an arrow along direction of transport would be helpful in addition to the numbers in 11b.
\end{point}

\begin{reply}
    Thanks, done. Also changed the km markings to the turn locations and added ticks on panel d corresponding to the tick locations.
\end{reply}

\begin{point}

    Figure 12: Which data set is this? Worth including in caption and maybe text also.
\end{point}

\begin{reply}
    Fixed - it was the MVP data, though the previous version was a subsample.
\end{reply}

\begin{point}
    338: is advection demonstrated or suggested?
\end{point}

\begin{reply}
    Thanks, "suggesting" is definitely better here.
\end{reply}

\begin{point}
    351: infer significant mixing [has occurred].
\end{point}

\begin{reply}
    Thanks.
\end{reply}

\begin{point}

    393: Is this explicitly shown on a T/S diagram? Perhaps plucking out profiles through the eddy for comparison can make this point more cleanly. Is there preferential warming within the eddy relative to other water masses?
\end{point}

\begin{reply}
    Yes, but perhaps better shown in Fig 3e, k, which are now explicitly referenced.
\end{reply}

\begin{point}
    414: wording?
\end{point}

\begin{reply}
    Fixed (hopefully) to
    \begin{lstlisting}
        In this calculation, the T-S relationship does not approach
        a straight line until after approximately 60-100 days, or
        until the mixing affects a vertical length scale of
        $\lambda = \left(\tau \kappa\right)^{1/2} \gtrapprox 70\,
        \mathrm{m}$.
    \end{lstlisting}
\end{reply}

\begin{point}
    468: How is this value chosen?
\end{point}

\begin{reply}
    This was just a typical weak coastal current scaling.  We now cite Thomson and Kassovski and point out that this current could be substantially stronger.
\end{reply}

\end{document}

### 2023-05-02 R0/PW13Paper.tex
R1

--- 2023-01-24 PW13Paper.tex
@@
 
-\graphicspath{{./figs/}}
+% \graphicspath{{./figs/}}
 \DeclareGraphicsExtensions{%
@@
 A goal of our study was to understand how the \Eddy\ persists and how it exchanges water properties with offshore water.
-We focus on water deeper than $50 \mathrm{m}$, below the summer mixed layer, because this water is trackable with water mass properties, even though the \Eddy\ is often studied from the point of view of the surface circulation \citep[e.g.][]{macfadyenhickey10}.
+We focus on water deeper than $50 \mathrm{m}$, below the summer mixed layer, because this water is trackable with water mass properties, even though the \Eddy\ is often studied from the point of view of the surface circulation \cite{macfadyenhickey10}.
 Past studies have found low oxygen concentrations in the \Eddy, and inferred that water is upwelled from the California Undercurrent from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
@@
 Net cross-shore transport of nutrients and chlorophyll have been found in hydrographic surveys off Vancouver Island \cite{mackasyelland99}, and associated with mesoscale features in geostrophic velocities and in satellite observations.
-The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics leading to along-shore trapping of coastal waves \cite{dalebarth01},  wind stress curl variations \cite{castelaobarth07}, induced relative vorticity due to stretching of parcels that overshoot their initial isobaths due to a sudden change in downstream bathymetry \citep{dasaro88}, and, at this location, non-linear breaking of large-scale meanders due to baroclinic instability between the wind-driven current and the California Undercurrent \cite{ikedaetal84, batteen97}.
+The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics leading to along-shore trapping of coastal waves \cite{dalebarth01},  wind stress curl variations \cite{castelaobarth07}, induced relative vorticity due to stretching of parcels that overshoot their initial isobaths due to a sudden change in downstream bathymetry \cite{dasaro88}, and, at this location, non-linear breaking of large-scale meanders due to baroclinic instability between the wind-driven current and the California Undercurrent \cite{ikedaetal84, batteen97}.
 
@@
   \begin{center}
-     \includegraphics[width=130mm]{LaPerouse2013Ctd}
+     \includegraphics[width=5in]{LaPerouse2013Ctd}
      \caption{Observations along LC Line in May (a--c) and Sep (g--i) and LB lines in May (d--f) and Sep (j--l).

### 2023-05-02 R1/PW13Paper.tex
R1

--- 2023-05-02 R0/PW13Paper.tex
@@
 \usepackage{lineno}
+\usepackage[]{amssymb}
 % \graphicspath{{/Users/jklymak/Dropbox/PW13/satellite/doc/},
@@
 \usepackage[plain]{fancyref}
+\renewcommand*{\freffigname}{Fig.}
 % \usepackage[]{hyperref}
@@
 
-Observations of temperature, salinity, and oxygen on the southern Vancouver Island shelf show a large-scale exchange of shelf water with offshore water, just offshore of a semi-permanent recirculation.
-The semi-permanent cyclonic recirculation (the Juan de Fuca Eddy) occupies a region where the shelf widens abruptly in the lee of a bank.
+Observations of temperature, salinity, and oxygen on the southern Vancouver Island shelf show a large-scale exchange of shelf water with offshore water, just offshore of a semi-permanent recirculation, often termed the Juan de Fuca Eddy.
+The Eddy occupies a region where the shelf widens abruptly in the lee of a bank.
 The water in this Eddy is a mixture of offshore water and water from a buoyant coastal current.
 This water is well-mixed along a mixing line in temperature-salinity space, though it retains stratification, and is either rapidly mixed or has a long residence time.
-There is a sharp temperature-salinity front on the offshore side of this well-mixed water, no more than 1-km wide, and has no sign of instabilities.
-The clearest evidence of cross-front transport is found during a short tidally resolved survey over a bank, and was due to flows in the cross-bank direction driving large mixing in 50-m tall hydraulic jumps.
-Upstream of the recirculation there is an along-shelf current flowing equatorward.
-However the whole current separates from the shelf before reaching the recirculation, in the lee of a bank, and is replaced by water from offshore.
+There is a less than 1-km wide temperature-salinity front on the offshore side of this well-mixed water that has no sign of instabilities.
+The clearest evidence of cross-front transport is found during a tidally resolved survey over a bank.
+The transport is due to flows in the cross-bank direction that also drive 50-m tall hydraulic jumps.
+Upstream of the Eddy there is an along-shelf current flowing equatorward.
+However the whole current separates from the shelf before reaching the Eddy, in the lee of a bank, and is replaced by water from offshore.
 The separation event was also seen in sea-surface temperatures from satellite images as a tongue of cool coastal water that is ejected offshore.
@@
 The observations show that below the surface mixed layer, the water in the Eddy is low in oxygen, and has undergone substantial vertical and lateral mixing.
-In contrast to previous literature we find that the low oxygen in the eddy is likely because of respiration rather than being pulled from low-oxygen water in the California Undercurrent.
+In contrast to previous literature we find that the low oxygen in the Eddy is likely because of respiration rather than being pulled from low-oxygen water in the California Undercurrent.
 
@@
 Here we present detailed \emph{in-situ} observations from the southern Vancouver Island shelf, collected in summer 2013, and sampled by a rapid profiling vehicle equipped with a CTD and oxygen sensor, supplemented by traditional hydrographic surveys bracketing the high resolution observations by a month.
-The shelf has complicated bathymetry (\fref{fig:LocMapBoth}), with a relatively simple 50-km wide shelf poleward of our study site, that widens to over 75 km wide equatorward of La Perouse Bank.
-The South Vancouver Island Shelf is this wide shelf region, characterized by a number of banks, and finally is incised on the poleward side by the Juan de Fuca canyon.
+The shelf has complicated bathymetry (\fref{fig:LocMapBoth}), with a relatively simple 50-km wide shelf poleward of our study site that widens to over 75 km wide equatorward of La Perouse Bank.
+The South Vancouver Island Shelf is this wide shelf region, characterized by a number of banks, and finally is incised on the equatorward side by the Juan de Fuca Canyon.
 Water from the Strait of Juan de Fuca flows poleward as a buoyant current that hugs the coast \cite{thomsonetal89, hickeyetal91}, while shelf water flows equatorward in the summer both forced by local winds and via teleconnections with the long, homogenous shelves equatorward off Washington and Oregon \cite{hickeyetal91,thomsonkrassovski15,engidaetal16}.
@@
 We focus on water deeper than $50 \mathrm{m}$, below the summer mixed layer, because this water is trackable with water mass properties, even though the \Eddy\ is often studied from the point of view of the surface circulation \cite{macfadyenhickey10}.
-Past studies have found low oxygen concentrations in the \Eddy, and inferred that water is upwelled from the California Undercurrent from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
-This inference was based on the assumption that oxygen is conservative, and hence water in the \Eddy\ had to come from  the oxygen minimum zone found further offshore \cite{mackasetal87}.
-This lead to hypotheses that perhaps the \Eddy\ was fed by waters being drawn up a spur canyon (called the Spur Canyon) via ageostrophic transport from offshore to onshore due to low pressure in the \Eddy\ center \cite{weaverhsieh87}.
+Past studies found low oxygen concentrations in the \Eddy, and inferred that water is upwelled from the California Undercurrent from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.
+This inference was based on the assumption that oxygen is a conservative tracer, and hence water in the \Eddy\ had to come from  the oxygen minimum zone found further offshore \cite{mackasetal87}.
+It was hypothesized that the \Eddy\ was fed by waters being drawn up a spur canyon (called the Spur Canyon) via ageostrophic transport from offshore to onshore due to low pressure in the \Eddy\ center \cite{weaverhsieh87}.
 Below, we will argue that there is no evidence of such transport and that oxygen is likely low because of local consumption due to respiration.
@@
 A second goal of our study was to better understand cross-shore exchange between shelf and offshore water.
-Cross-shore transport is supplied by Ekman layers in simple geometries under wind forcing.
-However, in many locations there is also evidence of eddies and filaments driving wholesale separation of shelf currents into the deep ocean.
-Along the Vancouver Island shelf, large filaments or instabilities of the coastal current have been inferred using satellite images \cite{ikedaemery84,thomsongower98}.
+For a two-dimensional topography with wind forcing, cross-shore transport is supplied by Ekman layers.
+However, in many locations there is also evidence of eddies, meanders and filaments driving wholesale separation of shelf currents into the deep ocean.
+These have been well-studied off California \cite{strubetal91} where satellite and \emph{in-situ} observations show large meanders of coastal upwelling currents leading to coastal water injected into the offshore domain.
+Along the Vancouver Island shelf, large instabilities of the coastal current have also been inferred using satellite images \cite{ikedaemery84,thomsongower98}.
 Direct observations of such filaments have been made further to the south off Cape Blanco \cite{barthetal00}, and off other coasts \cite{relvasbarton05}.
@@
 
-In this paper we present observations of water masses on the Southern Vancouver Island Shelf collected in 2013 (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes, with a focus on the offshore side of the \Eddy\ region (\fref{sec:Observations}) where we consider the age of the \Eddy, the steadiness of the front separating the \Eddy\ and the offshore water, properties along the spur canyon that was believed to feed the \Eddy, and a large offshore tongue of the shelf-break current and its intrusion into the interior.
-We discuss the origin of the \Eddy\ and the implications of the separating jet (\fref{sec:Summary}).
+In this paper we present observations of water masses on the Southern Vancouver Island Shelf collected in 2013 (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes.
+We focus on the offshore side of the \Eddy\ region (\fref{sec:Observations}) where we consider the age of the \Eddy, the steadiness of the front separating the \Eddy\ and the offshore water, properties along the Spur Canyon that was believed to feed the \Eddy, and a large offshore tongue of the shelf-break current and its intrusion into the interior.
+We conclude by discussing the origin of the \Eddy\ and the implications of the separating jet (\fref{sec:Summary}).
 
@@
 
-\begin{figure*}[htbp]
-  \includegraphics[width=5in]{LocMapBoth.pdf}
+\begin{figure}[htbp]
+  \includegraphics[width=5.4in]{LocMapBoth.pdf}
   \caption{a) Study site on the Vancouver Island Shelf.
@@
 The dashed box indicates the approximate limits of the study area.
-b) The study area with hydrographic casts from the La Perouse cruises along the LB and LC Lines (green dots), hydrographic casts during the Falkor cruise (yellow diamonds), and finescale Moving Vessel Profiler casts (red dots).
-The coordinate system used for this paper is shown with alternating grey and white bands at 10-km intervals in the along- and cross-shore directions.}
+b) The study area with major bathymetric features labelled.
+The coordinate system used for this paper is shown with alternating grey and white bands at 10-km intervals in the along- and cross-shore directions.
+c) sample locations with hydrographic casts from the La Perouse cruises along the LB and LC Lines (green dots), hydrographic casts during the Falkor cruise (yellow diamonds), and finescale Moving Vessel Profiler casts (red dots).
+Magenta
+  square is La Perouse wind buoy.}
   \label{fig:LocMapBoth}
-\end{figure*}
+\end{figure}
 
 The study site was the southern portion of the Vancouver Island Shelf (\fref{fig:LocMapBoth}a), a particularly complicated region due to the bathymetry and varied forcing.
-At the south end of the study site is the Juan de Fuca canyon, which feeds dense water into Juan de Fuca Strait.
+At the south end of the study site is the Juan de Fuca Canyon, which feeds dense water into Juan de Fuca Strait.
 This dense water is mixed with fresh water from the Fraser River at the sills and archipelagos further inland and fluxes out the Strait again, where it turns poleward along Vancouver Island to form the Vancouver Island Coastal Current.
-The Juan de Fuca Canyon has a notable spur canyon (Spur Canyon), that incises the shelf towards the north into the study site (\fref{fig:LocMapBoth}b).
+The Juan de Fuca Canyon has a notable spur canyon (Spur Canyon) that incises the shelf towards the north into the study site (\fref{fig:LocMapBoth}b).
 The rest of the shelf is punctuated by a series of banks and shallow basins.
-La Perouse bank  separates the outer shelf from a deeper inner basin and forms the poleward boundary off where the shelf widens abruptly south of 48.5 N.
+La Perouse Bank  separates the outer shelf from a deeper inner basin and forms the poleward boundary off where the shelf widens abruptly south of 48.5 N.
 Equatorward, the shelf widens further, and the shelf break has a series of submarine canyons, in particular Nitnat Canyon and Barkley Canyon.
 
-We present observations from La Perouse hydrographic surveys along the LB and LC lines, collected aboard the \emph{CCGS Tully} from 2013-05-30 to 2013-05-31  (``May''), and 2013-09-07 to 2013-09-09 (``September'').
+We use an along/across-shelf co-ordinate system with an origin at
+48.4\textdegree N and 125.6\textdegree W, with the across-shelf $x$ oriented 39 degrees north of East (\fref{fig:LocMapBoth}b, grey/white alternating lines).
+The projection is Cartesian with a central latitude at 48.4\textdegree N, which is fine for the limited geographic extents discussed here.
+
+Observations from La Perouse hydrographic surveys along the LB and LC lines, collected aboard the \emph{CCGS Tully} from 2013-05-30 to 2013-05-31  (``May''), and 2013-09-07 to 2013-09-09 (``September'') are used to put finescale observations in context of routine hydrographic work in the study region.
 These lines span the 50 m isobath to deep offshore, with casts every $7.5\ \mathrm{km}$ across shelf.
@@
 
-We focus on finescale surveys carried out between these hydrographic surveys, from 2013-08-21 to 2013-08-30.
+The focus in this paper is on finescale surveys carried out between these hydrographic surveys, from 2013-08-21 to 2013-08-30.
 Data were collected from the \emph{R/V Falkor} with an AML Oceanographic Moving Vessel Profiler (MVP).
@@
     \caption{
-      Wind from the La Perouse buoy \cite{DFOWind2013C46206}.
+      Wind from the La Perouse buoy at 48.84 N, 126 W \cite{DFOWind2013C46206}.
 The vertical direction is along-shelf (chosen as a heading of 315 N), so vectors pointing straight down represent upwelling winds.
@@
 
-Hydrographic sections along the LB and LC hydrographic lines highlight summer conditions on the Southern Vancouver Island Shelf and indicate some of the features we are focusing on in this paper (\Fref{fig:LaPerouse2013Ctd}).
+Hydrographic sections along the LB and LC hydrographic lines highlight summer conditions on the Southern Vancouver Island Shelf and presented to give context to the more detailed survey carried out on the \emph{R/V Falkor} (\Fref{fig:LaPerouse2013Ctd}).
 During both surveys, and along both lines, there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m depth offshore to shallower than 85 m over the shelf.
-This upwelled water tends to be low in oxygen and cool.
+The upwelled water tends to be low in oxygen and cool.
 Further onshore, the Vancouver Island Coastal Current hugs the coast, where isopycnals tilt down towards the shore,and water is warmer than offshore.
@@
 The Juan de Fuca Eddy region is observed along the LB Line (\fref{fig:LaPerouse2013Ctd}, bottom three rows), mostly inshore of 0 km.
-This water is cooler and lower in oxygen than along the same isopycnal further offshore.
+The water in the \Eddy\ is cooler and lower in oxygen than along the same isopycnal further offshore.
 Near the surface, the isopycnals are domed, and the low-oxygen anomaly extends as high as the thin surface mixed layer.
 
-The water upstream of the \Eddy\ region, along the LC line, is less mixed than the \Eddy\ water (\fref{fig:LaPerouse2013Ctd}, top three rows).
-It does not show as much drawdown of oxygen, and is warmer, at least offshore of La Perouse bank ($X \approx 5\ \mathrm{km}$).
-Based on these properties, offshore water does not appear to make it over La Perouse bank into the onshore basin, or if it does, it does so intermittently and with substantial mixing.
-Note that the water along the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal becomes cooler and lower in oxygen where it intersects the shelf, consistent with both enhanced mixing near the shelf and with drawdown of oxygen by respiration.
+The water upstream of the \Eddy\ region, along the LC line, is less mixed than the \Eddy\ water (\fref{fig:LaPerouse2013Ctd}, top three rows) and has higher oxygen.
+Based on these properties, offshore water does not appear to make it over La Perouse Bank into the onshore basin, or if it does, it does so intermittently and with substantial mixing.
+Note that the water along the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal (\fref{fig:LaPerouse2013Ctd}, magenta contours) becomes cooler and lower in oxygen where it intersects the shelf, consistent with both enhanced mixing near the shelf and with drawdown of oxygen by respiration.
 
@@
 
-We use a spice anomaly defined as relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}) representing an approximate mixing line for water in the \Eddy, and passing through the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$.
-Spice anomaly for a given water sample at density $\sigma_{\theta}$ is given by $\gamma = \alpha \left(T - T_0\left(\sigma_{\theta}\right)\right) + \beta \left(S - S_0\left(\sigma_{\theta}\right)\right)$ where $T_0\left(\sigma_{\theta}\right)$, where $S_0\left(\sigma_{\theta}\right)$ are the temperature and salinity along the mixing line at the same density as the water sample.
-The sign convention is that a positive anomaly is warmer and saltier than data along the mixing line.
+We use a spice anomaly defined as relative to a straight line in $\theta$--$S$
+space (\fref{fig:LaPerouse2013TS}) representing an approximate mixing line for
+water in the \Eddy, and passing through the points $7.75\ \mathrm{^oC}$, $30\
+\mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$. Spice anomaly for a
+given water sample at density $\sigma_{\theta}$ is given by $\gamma = \alpha
+\left(T - T_0\left(\sigma_{\theta}\right)\right) + \beta \left(S -
+S_0\left(\sigma_{\theta}\right)\right)$ where
+$T_0\left(\sigma_{\theta}\right)$, where $S_0\left(\sigma_{\theta}\right)$ are
+the temperature and salinity along the mixing line at the same density as the
+water sample. The sign convention is that a positive anomaly is warmer and
+saltier than data along the mixing line.
 
@@
 
-Using this metric of spice anomaly, offshore water tends to have high absolute values (\fref{fig:LaPerouse2013Ctd}, middle rows), with a positive-anomaly layer (red) centered at $26.4\ \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and below.
-On the shelf ($-10\, \mathrm{km} < X < 30\, \mathrm{km}$), the distinct $\theta$--$S$ masses are attenuated and much closer to the defined mixing line than the offshore water (closer to white in color in \fref{fig:LaPerouse2013Ctd}).
+Using this metric of spice anomaly, offshore water tends to have high absolute
+values (\fref{fig:LaPerouse2013Ctd}, middle rows), with a positive-anomaly
+layer ($\gamma \approx 0.2 \mathrm{kg\,m^{-3}}$, red colors) centered at $26.4\
+\mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and
+below.  On the shelf ($-10\, \mathrm{km} < X < 30\, \mathrm{km}$), the distinct
+$\theta$--$S$ masses are attenuated and much closer to the defined mixing line
+than the offshore water ($\gamma \approx  0\,\mathrm{kg\,m^{-3}}$, almost
+white \fref{fig:LaPerouse2013Ctd}).
 
-There are temporal changes over the summer.
-Away from the surface, water has upwelled from deeper depths by September, but the offshore water maintains the same water mass characteristics through the summer.
-Hugging the coast ($X > 30\, \mathrm{km}$), the Vancouver Island Coastal Current warms during the summer, and the spice anomaly goes from negative to positive.
-In the \Eddy, the water stays near the mixing line, but is cooler in the spring (\fref{fig:LaPerouse2013Ctd}, LB Line, left-hand column: slightly negative spice anomaly) and warms during the summer (\fref{fig:LaPerouse2013Ctd}, LB Line, right-hand column: slightly positive spice anomaly).
+There are temporal changes over the summer.  Away from the surface, water has
+upwelled from deeper depths by September, but the offshore water maintains the
+same water mass characteristics through the summer.  Hugging the coast ($X >
+30\, \mathrm{km}$), the Vancouver Island Coastal Current warms during the
+summer, and the spice anomaly goes from negative to positive.  In the \Eddy,
+the water stays near the mixing line, but is cooler in the spring
+(\fref{fig:LaPerouse2013Ctd}, LB Line, left-hand column: slightly negative
+spice anomaly) and warms during the summer (\fref{fig:LaPerouse2013Ctd}, LB
+Line, right-hand column: slightly positive spice anomaly).
 
-Oxygen concentration in both surveys has a similar dichotomy between the \Eddy\ and off-shelf water.
-Water found in the \Eddy\ has oxygen concentrations $100\ \mathrm{\mu mol\, kg^{-1}}$ lower on the shelf than off-shelf (\fref{fig:LaPerouse2013Ctd}, \fref{fig:LaPerouse2013TS}b).
-The very deepest water ($S\approx 33.9\ \mathrm{psu}$) on the shelf shows a further $50\ \mathrm{\mu mol\, kg^{-1}}$ decrease in concetration between May and September along the LB line.
+Oxygen concentration in both surveys has a similar dichotomy between the \Eddy\
+and off-shelf water.  Water found in the \Eddy\ has oxygen concentrations $100\
+\mathrm{\mu mol\, kg^{-1}}$ lower on the shelf than off-shelf
+(\fref{fig:LaPerouse2013Ctd}, \fref{fig:LaPerouse2013TS}b).  The very deepest
+water ($S\approx 33.9\ \mathrm{psu}$) on the shelf shows a further $50\
+\mathrm{\mu mol\, kg^{-1}}$ decrease in concetration between May and September
+along the LB line.
 
@@
      \includegraphics[width=3.5in]{TSdensSpice.pdf}
-    \caption{Binned sample density of salinity and potential temperature from the cruise, with a logarithmic color scale.
+    \caption{a) Binned sample density of salinity and potential temperature from the cruise, with a logarithmic color scale.
 Grey contours are potential density relative to the surface at intervals of $0.1\ \mathrm{kg\,m^{-3}}$; the magenta contour is the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.
 Colored contours are spice anomaly, $\gamma$, relative to the white line labeled $\gamma=0.0$.
+b) Distribution of spice, $\gamma$ along the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.
     \label{fig:TSdensSpice}}
@@
 In synthetic cross sections of spice anomaly, the \Eddy\ is clearly identifiable as having low spice anomaly ($\gamma \approx 0\ \mathrm{kg\,m^{-3}}$, \fref{fig:CrossSectionsSpice}c, d).
-This signature of well-mixed water extends from the seafloor to approximately 30 m depth, onshore of $X \approx 0\ \mathrm{km}$.
+These lines are made with a two-dimensional interpolation where MVP casts at a given depth have a Gaussian weight $w = e^{-(r/r_0)^2}$, where $r$ is the distance from the line and $r_0 = 1.5\ \mathrm{km}$.
+Note that the section at   The low-spice anomaly water extends from the seafloor to approximately 30 m depth, onshore of $X \approx 0\ \mathrm{km}$.
 Despite lying along a mixing line, the \Eddy\ water is still stratified.
@@
     \includegraphics[width=6.2in]{CrossSectionsSpice}
-    \caption{Synthetic sections of spice anomaly from the MVP data, projected along four lines from poleward (upstream of coastal current) a) to equatorward (downstream); the lines are shown in \fref{fig:LocMapBoth}b).
+    \caption{Synthetic sections of spice anomaly from the MVP data, projected along four lines from poleward (upstream of coastal current) a) to equatorward (downstream); the lines are shown in \fref{fig:LocMapBoth}a).
 Isopycnals are contoured every $\sigma_{\theta} = 0.2\,\mathrm{kg\,m^{-3}}$ , with the $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ isopycnal highlighted in magenta.
@@
 "BC" in panel (c) refers to Barkley Canyon.
+Note that b) is along LC line (\fref{fig:LaPerouse2013Ctd}a--c,g--i), and d) is along the LB line (\fref{fig:LaPerouse2013Ctd}d--f,j--l).
       \label{fig:CrossSectionsSpice} }
@@
 The spatial patterns are very clear when considering a map view of properties along the $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ isopycnal (\fref{fig:SpiceO2264}).
-There is a region onshore of the south tip of La Perouse Bank (approximately 44.5\textdegree N, 125.75\textdegree W), that consists of water that is found along the mixing line (spice anomaly $\gamma \approx 0\ \mathrm{kg\,m^{-3}}$) that also has low oxygen saturation.
+There is a region onshore of the south tip of La Perouse Bank (approximately 48.5\textdegree N, 125.75\textdegree W), that consists of water that is found along the mixing line (spice anomaly $\gamma \approx 0\ \mathrm{kg\,m^{-3}}$) that also has low oxygen saturation.
 Offshore of this region is water that is very warm and salty in comparison ($\gamma > 0.15 \mathrm{kg\,m^{-3}}$), and relatively high in oxygen (saturations of approximately 60\%).
-The region between these two water masses is very abrupt, and  stretches from La Perouse Bank to a shallow bank just above the Juan de Fuca canyon (48.3\textdegree N and 125.4\textdegree W).
+The transition between these two water masses is very abrupt, and  stretches from La Perouse Bank to a shallow bank just above the Juan de Fuca Canyon (48.3\textdegree N and 125.4\textdegree W).
 
@@
 This water is also somewhat lower in oxygen than water from offshore, though not as depleted as the water in the \Eddy.
-This water appears to be pushed offshore just upstream of Barkley Canyon and replaced on the shelf by the warmer (high spice) offshore water.
+The water appears to be pushed offshore just upstream of Barkley Canyon and replaced on the shelf by the warmer (high spice) offshore water.
+
+The correspondence between low oxygen water and water on the mixing line that defines $\gamma \approx 0\ \mathrm{kg\,m^{-3}}$ is quite strong (\fref{fig:SpiceO2264}c).
+Eddy water is onshore (more red colors), low in oxygen, and along the mixing line.
+Offshore water is high in oxygen and has a high spice anomaly.
+Water that is found poleward on the shelf is still relatively high oxygen, but of intermediate spice $\gamma \approx 0.125\ \mathrm{kg\,m^{-3}}$.
 
@@
 Along-slope grey lines are every 20 km in the cross-slope direction, with $X=0\ \mathrm{km}$ near the 100-m isobath at the north end of the observation area.
+c) distribution of oxygen as a function of spice anomaly and $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$, colored by x-coordinate.
+Data from the Juan de Fuca Canyon region ($y<-12\ \mathrm{km}$) is not shown.
    \label{fig:SpiceO2264}
@@
 
-The Spur Canyon leading from the Strait of Juan de Fuca has been implicated in allowing dense water to be upwelled into the \Eddy.
+The Spur Canyon leading from the Strait of Juan de Fuca has been implicated in allowing dense water to be upwelled into the \Eddy \cite{mackasetal87,weaverhsieh87}.
 The sea surface is low in the middle of the \Eddy, so it has been hypothesized that water moves up the Spur Canyon due to ageostrophic motion \cite{weaverhsieh87,freelanddenman82}.
-This is difficult to infer from the observations collected here.
+It is difficult to infer flow up the Spur Canyon from the observations collected here.
 Three transects up the canyon indicate that deep isopycnals slope up into the canyon (\fref{fig:CanyonPropertiesSpice}, to approximately 30 km).
-Continuing across the \Eddy, isopycnals are largely flat until they intersect the Vancouver Island Coastal Current ($80\ \mathrm{km}$).
-The deeper isopycnals are not found in the deeper basin northeast of La Perouse bank ($70\ \mathrm{km}$),  thus the bank is a natural poleward boundary of the \Eddy.
+Continuing across the \Eddy, isopycnals are largely flat until they intersect the Vancouver Island Coastal Current ($X_{canyon} = 80\ \mathrm{km}$).
+The deeper isopycnals are not found in the deeper basin northeast of La Perouse Bank ($70\ \mathrm{km}$),  thus the bank is a natural poleward boundary of the \Eddy.
 
@@
     \includegraphics[width=6.2in]{CanyonPropertiesSpice}
-    \caption{Spice anomaly surveys up the Spur Canyon, where $X$ is along-canyon as defined in the map.
+    \caption{Spice anomaly surveys up the Spur Canyon, where $X_{canyon}$ is along-canyon as defined in the map.
 Grey isopycnals are contoured every $0.5\,\mathrm{kg\,m^{-3}}$, and the $26.4\,\mathrm{kg\,m^{-3}}$ isopycnal is shown in magenta.
@@
 Given the stratification of $N\approx 6\times10^{-3} \ \mathrm{rad\,s^{-1}}$ and an overturning scale of 50 m, we might expect dissipation rates reaching $\epsilon \sim L^2N^{3} \approx 5\times10^{-4}\ \mathrm{m^2\,s^{-3}}$, which is three orders of magnitude higher than dissipation observed on the shelf west of this location by \citeA{deweycrawford88}.
-This estimate of turbulence dissipation rate implies a diapycnal diffusivity of $\kappa = \gamma \epsilon / N^2 \approx 1\ \mathrm{m^2\,s^{-1}}$, assuming a mixing efficiency of $\gamma=0.2$.
+This estimate of turbulence dissipation rate implies a diapycnal diffusivity of $\kappa = \gamma \epsilon / N^2 \approx 0.5$ -- $5\ \mathrm{m^2\,s^{-1}}$, assuming a mixing efficiency of $\gamma=0.2$.
 Water spills over from the offshore front into the canyon during the onshore tide.
@@
 There is a general tendency along the canyon for higher spice water to be found offshore (\fref{fig:CanyonPropertiesSpice}) but it seems likely that the source of the higher spice water is from over the bank rather than water being advected up the canyon.
-This tidally driven flow over the bank is the most significant source of high-spice offshore water into the eddy region identified during our surveys.
+The tidally driven flow over the bank is the most significant source of high-spice offshore water into the \Eddy\ region identified during our surveys.
 
@@
 The survey started close to shore, and passed through the Vancouver Island Coastal Current (along-track 0-10 km).
-The coastal current forms a buoyant front, and is fresher and colder than water at the same density.
-This front is relatively thick, greater than 20 km wide, and has partially mixed water from the surface to the foot of the front ($\gamma\approx0.1$).
+The coastal current forms a buoyant current, and is fresher and colder than water at the same density.
+The front with the coastal water is relatively thick, greater than 20 km wide, and has entrained partially mixed water down from the surface to the foot of the front ($\gamma\approx0.1$).
 
@@
 First, as noted in the composite sections, isopycnals slope up from offshore onto the shelf.
-In the first crossing, the front is very sharp, (along-track distance $50\ \mathrm{km}$) though two small tendrils can be seen separating from the front on the inshore side.
+In the first crossing, the front is very sharp, (along-track distance $50\ \mathrm{km}$) though two small tendrils of higher spice water can be seen separating from the front on the inshore side.
 Similar tendrils are found on the second crossing, perhaps a bit more separated from the front ($\approx 80\ \mathrm{km}$), and on the third crossing ($\approx 95\ \mathrm{km}$).
@@
 
-There is evidence of some warmer water swirling into Eddy, particularly along isopycnals deeper than $26.4\,\mathrm{kg\,m^{-3}}$.
-Regions of warmer (and more oxygenated) water are found in tendrils at these depths (e.g.\ $\approx\,170\ \mathrm{km}$ and $\approx 185\,\mathrm{km}$).
-The overall effect is similar to what was seen in the Gulf Stream with similar observations \cite{klymaketal16}; there are two quite distinct water masses, the Eddy water and the offshore water, as seen in the $\theta$--$S$ plot (\fref{fig:Frontsurvey}c) with only a small population of samples between these two.
-This distribution of $\theta$--$S$ properties is indicative of substantial isopycnal and vertical mixing, but even these populations are relatively cut off from the main water masses, indicating that they are well-mixed on their own, in short episodic events.
+There is evidence of some warmer water swirling into \Eddy, particularly along isopycnals deeper than $26.4\,\mathrm{kg\,m^{-3}}$.
+Regions of warmer (and more oxygenated) water are found in tendrils at these depths (e.g.\ $\approx 170\ \mathrm{km}$ and $\approx 185\,\mathrm{km}$).
+The overall effect is similar to what was seen in the Gulf Stream with similar observations \cite{klymaketal16}; there are two quite distinct water masses, the \Eddy\ water and the offshore water, as seen in the $\theta$--$S$ plot (\fref{fig:Frontsurvey}c) with only a small population of samples between these two.
+These distinct $\theta$--$S$ properties are indicative of substantial isopycnal and vertical mixing, but even these populations are relatively cut off from the main water masses, indicating that they are well-mixed on their own, in short episodic events.
 Regardless, this front is very sharp given that it has no density signature, indicating that there is not strong advection from offshore into the \Eddy\ region.
@@
     \includegraphics[width=6.2in]{Frontsurvey}
-    \caption{Data from the frontal survey  a) spice anomaly along $26.4\,\mathrm{kg\,m^{-3}}$.
+    \caption{Data from the offshore front survey  a) spice anomaly along $26.4\,\mathrm{kg\,m^{-3}}$.
 Grey-white alternating lines are the coordinate system, with 10-km alternating shades.
-b) oxygen saturation along $26.4\,\mathrm{kg\,m^{-3}}$; magenta dots correspond to distance along-track.
+b) oxygen saturation along $26.4\,\mathrm{kg\,m^{-3}}$; magenta dots correspond to turn locations as distance along-track.
 c) density of data points (logscale) in this section of data.
@@
 Grey isopycnals are contoured every $0.5\,\mathrm{kg\,m^{-3}}$, and the $26.4\,\mathrm{kg\,m^{-3}}$ isopycnal is shown in magenta.
-e) cross-section of oxygen saturation.
+Ticks are turn locations e) cross-section of oxygen saturation.
       \label{fig:Frontsurvey} }
@@
 Upstream of the the sharp front between the low-spice \Eddy\ and the high-spice offshore water is a substantial mass of intermediate-spice water along the shelf (compare \fref{fig:CrossSectionsSpice}a and c).
-This intermediate spice water is moving equatorward along the shelf upstream of the \Eddy, but is pushed offshore just upstream of Barkley Canyon (\fref{fig:IsopycnalsDepth}).
-There is a tongue of intermediate-spice water ($\gamma$ is pink along $26.4\,\mathrm{kg\,m^{-3}}$) that separates from the shelf just west of 126\textdegree W.
+This intermediate spice water is moving equatorward along the shelf upstream of the \Eddy, but is pushed offshore just upstream of Barkley Canyon (\fref{fig:IsopycnalsDepth}d).
+There is a tongue of intermediate-spice water ($\gamma \approx 0.1\,\mathrm{kg\,m^{-3}}$, pink along $26.4\,\mathrm{kg\,m^{-3}}$) that separates from the shelf just west of 126\textdegree W.
 The surveys do not cross the full extent of the  tongue, but it is at least 30 km wide.
 It also appears to end at approximately 48.2\textdegree N.
-This intermediate-spice water reaches from relatively shallow isopycnals to at least $26.6\,\mathrm{kg\,m^{-3}}$ (\fref{fig:IsopycnalsDepth}, left panel).
+This intermediate-spice water reaches from relatively shallow isopycnals to at least $26.55\,\mathrm{kg\,m^{-3}}$ (\fref{fig:IsopycnalsDepth}h).
 Unfortunately, we cannot track the fate of this water mass because isopycnals tilt down offshore, below the depth limit of the MVP.
-The $25.5\,\mathrm{kg\,m^{-3}}$ isopycnal appears to have the tongue closer to the shelf than at $26.4\,\mathrm{kg\,m^{-3}}$, indicating strong three-dimensionality to this feature.
+The $25.8\,\mathrm{kg\,m^{-3}}$ isopycnal appears to have the tongue (\fref{fig:IsopycnalsDepth}a), but closer to the shelf than at $26.4\,\mathrm{kg\,m^{-3}}$, indicating strong three-dimensionality to this feature.
 
-This separating tongue is embedded in the larger scale isopycnal tilt caused by the upwelling (\fref{fig:IsopycnalsDepth}, right panels), so it is difficult to see dynamically what is driving this offshore push.
-One possibility is that it is simple flow separation caused by the water not being able to make the sharp turn around La Perouse Bank.
+This separating tongue is embedded in the larger scale isopycnal tilt caused by the upwelling (\fref{fig:IsopycnalsDepth}, c,f, and j), so it is difficult to see dynamically what is driving this offshore push.
+If the offshore motion were geostrophically balanced, we would expect the isopycnals to dome upwards from poleward towards Barkley Canyon, but if such doming is happening, it is weaker than variability from internal tides and synoptic unsteadiness.
+One possibility is that it is simple flow separation caused by the water not being able to make the sharp turn around La Perouse Bank, which would be an ageostrophic effect.
 Whatever causes it, downstream of the separation, the water is replaced by offshore water with much higher spice values.
@@
     \includegraphics[width=6.2in]{IsopycnalsDepth}
-    \caption{Isopycnal slices through showing the vertical structure of the water separating from the shelf, with the first row along $25.5 \,\mathrm{kg\,m^{-3}}$, second along $26.4\,\mathrm{kg\,m^{-3}}$, and the third at $26.6\,\mathrm{kg\,m^{-3}}$.
+    \caption{Isopycnal slices of the MVP data showing the vertical structure of the water separating from the shelf, with the first row along $25.5 \,\mathrm{kg\,m^{-3}}$, second along $26.4\,\mathrm{kg\,m^{-3}}$, and the third at $26.6\,\mathrm{kg\,m^{-3}}$.
 First column is the spice anomaly, second is oxygen saturation, and the last column is depth of each isopycnal.
@@
 Water flowing equatorward along the shelf tends to be cooler than offshore water, likely due to mixing with the colder coming out of the Strait of Juan de Fuca.
-On August 25, there is a cold tongue of water separating from La Perouse bank, crossing isobaths and pointing south at 125.8\textdegree W.
+On August 25, there is a cold tongue of water separating from La Perouse Bank, crossing isobaths and pointing south at 125.8\textdegree W.
 There is a cooler tendril streaming west at 48 N off the south end of this tongue.
@@
 
-Satellite-based surface chlorophyll estimates show the same feature (\fref{fig:ChlA}) demonstrating the advection of high chlorophyll to the west side of the \Eddy.
+Satellite-based surface chlorophyll estimates show the same feature (\fref{fig:ChlA}) suggesting the advection of high chlorophyll to the west side of the \Eddy.
 They also show a relatively high-chlorophyll tendril to the west, again exiting the study region at approximately 48 degrees N.
@@
 The Juan de Fuca Eddy water is readily identified as falling along a mixing line in $\theta-S$ space, compared with offshore water that was warmer and saltier (high-spice anomaly).
-There was not strong evidence of the \Eddy\ being supplied by water moving up the Spur Canyon during our observations, but the Spur Canyon was a site of hydraulic cross-canyon flows in which we infer significant mixing.
+There was not strong evidence of the \Eddy\ being supplied by water moving up the Spur Canyon during our observations, but the Spur Canyon was a site of hydraulic cross-canyon flows in which we infer significant mixing has occurred.
 There is a sharp and persistent temperature-salinity compensated front between offshore water and the partially mixed \Eddy\ water.
@@
 If we posited that the reduction was all in the same water, then the oxygen consumption rate over the time between the late-May and early September cruises would be on the order of $0.5\ \mathrm{\mu mol\, kg^{-1} d^{-1}}$.
-This is on the low side of estimates of apparent oxygen utilization rates in continental upwelling systems, which are between $1$ and $5\  \mathrm{\mu mol\, kg^{-1} d^{-1}}$ \cite{dortchetal94}.
-So it seems possible the \Eddy\ had enough exchange with the surrounding water for the residence time to have been less than the full time period from late May to September.
-Note that during the May cruise, the water in the \Eddy\ is slightly cooler than the mixing line, and during the September cruise is slightly warmer than the mixing line, so the water in the \Eddy\ is evolving seasonally, and probably affected by the water temperature in the Vancouver Island Coastal Current and the amount of water getting through the onshore front.
+This is on the low side of estimates of apparent oxygen utilization rates in continental upwelling systems, which are between $1$ and $5\  \mathrm{\mu mol\, kg^{-1} d^{-1}}$ \cite{dortchetal94,connollyetal10}.
+So it seems likely that the \Eddy\ has exchange with the surrounding water.
+Note that during the May cruise, the water in the \Eddy\ is slightly cooler than the mixing line, and during the September cruise is slightly warmer than the mixing line, further evidence that the water in the \Eddy\ is evolving seasonally (\fref{fig:LaPerouse2013Ctd}e,k).
+This is consistent with findings on the Oregon shelf where physical processes are thought to account for 55-70\% of changes in dissolved oxygen concentrations \cite{adamsetal13}.
+Definitively identifying the exchange mechanisms into the \Eddy\ is challenging from this data set.
+The offshore front does not appear to have much exchange, however as noted above there does appear to be strong tidal flows and mixing over the submarine banks that can transport more oxygenated water into the \Eddy.
+Further, the Vancouver Island coastal current is oxygen rich, and has a much less sharp front than the offshore front (\fref{fig:Frontsurvey}).
 
-The water is  mixed enough in the \Eddy\ that it falls along a mixing line, though it remains vertically stratified.
-The amount of homogenization is such that either the mixing is very strong, or the water is retained in the \Eddy\ for a long time.
-The amount of turbulence required to homogenize 100 m of water over 90 days is $\kappa \approx 10^{-3}\ \mathrm{m^2\,s^{-1}}$.
-Given that the diffusivity implied in the cross-channel surveys was on the order of $\kappa_{\rho} = 10 \ \mathrm{m^2\,s^{-1}}$, this number is not outrageous if we think that such high dissipation is found in $10^{-4}$ of the water column.
-We can more carefully quantify this by considering a synthetic profile or temperature and salinity based on an offshore profile, extrapolated from the bottom of the 200-m cast to 250 m, assuming that the temperature and salinity at 250 m are $6.2\ \mathrm{[^oC]}$ and $34\ \mathrm{psu}$ respectively (\fref{fig:TSExercise}).
-Water at the offshore station was warmer than onshore, so the profile was also linearly interpolated to a surface value of ($12\ \mathrm{^oC}$, $31.6\ \mathrm{psu}$).
-The profile was then linearly compressed into a depth range of 150 m representative of the shelf depth in the dense pool, and subjected to mixing with a constant diffusivity of $\kappa = 10^{-3}\ \mathrm{m^2\,s^{-1}}$, with the surface and bottom values pinned under the assumption that the near-bottom source and surface waters are replenished from a large reservoir.
-Similar to the naive scaling, the T-S relationship does not approach a straight line until after approximately 60-100 days, or until the mixing affects a vertical length scale of $\lambda = \left(\tau \kappa\right)^{1/2}$ that is greater than $\approx 70\,\mathrm{m}$.
+The water is  mixed enough in the \Eddy\ that it falls along a mixing line,
+though it remains vertically stratified. The amount of homogenization is such
+that either the mixing is very strong, or the water is retained in the \Eddy\
+for a long time.  The amount of turbulence required to homogenize 100 m of
+water over 90 days is $\kappa \approx 10^{-3}\ \mathrm{m^2\,s^{-1}}$.  Given
+that the diffusivity implied in the cross-channel surveys was on the order of
+$\kappa_{\rho} = 0.5$--$5\ \mathrm{m^2\,s^{-1}}$, this number is not outrageous
+if we think that such high dissipation is found in 0.2--0.02\% of the water
+column. We can more carefully quantify this by considering a synthetic profile
+or temperature and salinity based on an offshore profile, extrapolated from the
+bottom of the 200-m cast to 250 m, assuming that the temperature and salinity
+at 250 m are $6.2\ \mathrm{[^oC]}$ and $34\ \mathrm{psu}$ respectively
+(\fref{fig:TSExercise}).  Water at the offshore station was warmer than
+onshore, so the profile was also linearly interpolated to a surface value of
+($12\ \mathrm{^oC}$, $31.6\ \mathrm{psu}$).  The profile was then linearly
+compressed into a depth range of 150 m representative of the shelf depth in the
+dense pool, and subjected to mixing with a constant diffusivity of $\kappa =
+10^{-3}\ \mathrm{m^2\,s^{-1}}$, with the surface and bottom values pinned under
+the assumption that the near-bottom source and surface waters are replenished
+from a large reservoir.  In this calculation, the T-S relationship does not
+approach a straight line until after approximately 60-100 days, or until the
+mixing affects a vertical length scale of $\lambda = \left(\tau
+\kappa\right)^{1/2} \gtrsim 70\,\mathrm{m}$.
 
@@
 
-The sharpness of the front with the \Eddy\ and the offshore water is intriguing.
-It was persistent for the duration of our detailed survey (\fref{fig:SpiceO2264}), and, so far as we can tell with the limited resolution of the hydrographic surveys, was present during the bracketing La Perouse cruises.
-There is not any substantial bathymetry blocking the onshore incursion of water at this location, so there must be a dynamic barrier.
+Combined with the oxygen observations, the implication is that the water in the \Eddy\ likely experiences exchange with the outside water, but at a very modest rate.
+In terms of a volume flux, we might estimate the \Eddy\ area as $900\ \mathrm{km^2}$, over 100 m depth, so that 100-d residence time corresponds to a transport of $10^{4}\ \mathrm{m^3s^{-1}}$, which is remarkably weak for the transport in and out of such a large area.
 
-Numerical simulations of this region reported by \cite{sahuetal22} using a NEMO 36th-degree regional model contained relatively rapid exchange between the deep \Eddy\ water and the rest of the coastal ocean.
-The region where the \Eddy\ resides has velocities equal or greater than other parts of the shelf, and water has an approximate residence time of less than 20 days.
-The \Eddy\ water in the model does develop a distinct $\theta$--$S$ signature, but not along a sharp mixing line as observed.
-It also has a front with the offshore water, but the front is substantially wider than that observed here.
+The sharpness of the front with the \Eddy\ and the offshore water is
+intriguing.  It was persistent for the duration of our detailed survey
+(\fref{fig:SpiceO2264}), and, so far as we can tell with the limited resolution
+of the hydrographic surveys, was present during the bracketing La Perouse
+cruises. There is not any substantial bathymetry blocking the onshore incursion
+of water at this location, so there must be a dynamic barrier.
 
-Overall, it would be an improvement to our understanding of the \Eddy\ if we could sample the shelf more persistently.
-The \Eddy\ was already well-formed by the May La Perouse cruise, and seems to evolve slowly during that time.
-Capturing its formation, presumably earlier in the spring, as well as its evolution through the year, would be valuable in understanding retention and exchange on this productive part of the shelf.
+Numerical simulations of this region reported by \cite{sahuetal22} using a NEMO
+36th-degree regional model contained relatively rapid exchange between the deep
+\Eddy\ water and the rest of the coastal ocean.  The region where the \Eddy\
+resides has velocities equal or greater than other parts of the shelf, and
+water has an approximate residence time of less than 20 days.  The \Eddy\ water
+in the model does develop a distinct $\theta$--$S$ signature, but not along a
+sharp mixing line as observed.  It also has a front with the offshore water,
+but the front is substantially wider than that observed here.  Overall it seems
+that the model sees stronger cross-shelf advection than are apparent in these
+observations, though the reason for this will require further study.
+
+Overall, it would be an improvement to our understanding of the \Eddy\ if we
+could sample the shelf more persistently.  The \Eddy\ was already well-formed
+by the May La Perouse cruise, and seems to evolve slowly during that time.
+Capturing its formation, presumably earlier in the spring, as well as its
+evolution through the year, would be valuable in understanding retention and
+exchange on this productive part of the shelf.
 
@@
 It has been recognized that instabilities lead to exchange between the open ocean and the shelf at this location \cite{ikedaemery84, ikedaemery84}.
-However, observations of the wholesale replacement of shelf water by a new water mass from offshore are rare.
+However, observations of the wholesale replacement of shelf water by a new water mass from offshore are relatively rare.
 In the  observations presented here, it is clear that water from as deep as 150 m is separating from the shelf and moving offshore (\fref{fig:SpiceO2264}, \fref{fig:IsopycnalsDepth}).
@@
 General baroclinic instability of the upwelling front is a possible mechanism to drive offshore exchange \cite{ikedaetal84,durskiallen05}, but this tends to be shallow, with smaller-scale instabilities that will not extend as far into the interior ocean as observed here.
-Rather it seems likely that the topographic change engendered by the sudden turn to the east of La Perouse bank catalyses a larger scale instability at this location.
-\citeA{durskiallen05}, when modelling the Oregon shelf, found that including realistic shelf bathymetry catalyzed intermittent large-scale instabilities similar to the feature here, a finding that is deemed likely to apply at other coastal locations \cite{batteen97}.
+Rather it seems likely that the topographic change engendered by the sudden turn to the east of La Perouse Bank catalyses a larger scale instability at this location.
+In California, most of the cold filaments observed appear to be catalyzed by headlands and underwater topography \cite{strubetal91}, though modelling studies find instabilities are possible even in two-dimensional flows \cite{pierceetal91}.
+\citeA{durskiallen05}, when modelling the Oregon shelf, found that including realistic shelf bathymetry catalyzed intermittent large-scale instabilities, a finding that is deemed likely to apply at other coastal locations \cite{batteen97}.
 
@@
 If we assume the coastal current is approximately $0.1\ \mathrm{m\,s^{-1}}$ over 100 m in the vertical and 20 km in the horizontal, it represents 0.2 Sv of nutrient- and chlorophyll-rich shelf water transported offshore.
-Our observations are a finer detailed representation of the kind of cross-shore transports inferred by \citeA{mackasyelland99} from hydrographic surveys, and definitively show that this water can originate from the shelf from relatively deep depths and be transported offshore.
+Sometimes the along-shelf currents are substantially larger than this \cite{thomsonkrassovski15} reaching $0.4\ \mathrm{m\,s^{-1}}$.
+Our observations are a finer-detailed representation of the kind of cross-shore transports inferred by \citeA{mackasyelland99} from hydrographic surveys, and definitively show that this water can originate from the shelf from relatively deep depths and be transported offshore.
 We do not have velocity measurements for the water that replaces it, but assuming that water also flows along-shelf, these separation events are associated with a large replacement of shelf water with offshore water at this location.
@@
 Thank you to Richard Dewey, Benjamin Schiefle, and Rowan Fox for efforts during the cruise.
+Jack Barth and an anonymous reviewer provided valuable feedback on the manuscript.
 Klymak was supported by NSERC Discovery grant RGPIN-2017-04050 and a Canadian Foundation for Innovation Leading Edge Fund grant 36109, Allen was supported by NSERC Discovery grant RGPIN-105694-11, and Waterman was supported by NSERC Discovery grant RGPIN-2020-05799.

### 2023-05-25 R1/ResponseReviewers.tex
R1: submitted

@@
 % General intro text goes here
-Thank you to both reviewers and the Editor for their time and effort into this paper.
-It is appreciated that this is a complicated story in this paper, and the reviewers' persistence in making sense of it with us have been very helpful in adding both better insights and (hopefully) better explanations.
+Thank you to both Reviewers and the Editor for their time and effort into this paper.
+We appreciate that this paper has a complicated story, and the Reviewers' persistence in making sense of it with us have been very helpful in adding both better insights and (hopefully) better explanations.
 
@@
 \begin{point}
-    The idea that the equatorward shelf current "separation is a very strong cross-shelf exchange event, and transports substantial nutrient-rich coastal water offshore to drive productivity in the deeper ocean" is like documented elsewhere in the California Current, for example as summarized by Strub et al.
+    The idea that the equatorward shelf current ``separation is a very strong cross-shelf exchange event, and transports substantial nutrient-rich coastal water offshore to drive productivity in the deeper ocean'' is like documented elsewhere in the California Current, for example as summarized by Strub et al.
 (1991).
@@
 Not only should we have cited it, but it summarizes a lot of the previous work off California, and it lays out the likely mechanisms of the separation we saw.
-The fourth paragraph of the introduction has been revised to include this reference.
+The fourth paragraph of the Introduction has been revised to include this reference.
 
@@
         between shelf and offshore water.  For a two-dimensional topography with
-        wind forcing, cross-shore transport is supplied by Ekman layers. However,
-        in many locations there is also evidence of eddies, meanders and filaments
-        driving wholesale separation of shelf currents into the deep ocean. These
-        have been well-studied off California \cite{strubetal91} where satellite
+        along-shore wind forcing, cross-shore transport is supplied by Ekman layers.
+        However, in many locations there is also evidence of eddies, meanders and
+        filaments driving wholesale separation of shelf currents into the deep ocean.
+        These have been well-studied off California \cite{strubetal91} where satellite
         and \emph{in-situ} observations show large meanders of coastal upwelling
@@
 
-    Further, in the discussion we have added reference to this paper again.
+    Further, in the Discussion we have added reference to this paper again:
 
     \begin{lstlisting}[language=TeX, basicstyle=\small]
-        In California, most of the cold filaments observed appear to be
+        In California most of the cold filaments observed appear to be
         catalyzed by headlands and underwater topography \cite{strubetal91},
@@
 \begin{point}
-    Lines 202-210: Water masses along 26.4 not clear in Fig 5; what are we supposed to see in the "\# of data points per bin" color map?
+    Lines 202-210: Water masses along 26.4 not clear in Fig 5; what are we supposed to see in the ``\# of data points per bin'' color map?
 I think you are highlighting the low \# of data points (more purple) between spiciness anomaly ~0.02-0.1 in between the populous offshore water (spiciness ~0.2) and EDDY water (spiciness 0).
@@
     alternating lines).  The projection is Cartesian with a central latitude at
-    48.4\textdegree N, which is fine for the limited geographic extents
+    48.4\textdegree N, which is sufficient for the limited geographic extents
     discussed here.
@@
     However, more to the point, inferring relatively subtle offshore velocities using along-isobath density gradients is challenging.
-For a geostrophically balanced anti-cyclonic circulation, we would expect a sea-surface low at the pivot point between the offshore moving water, and the onshore moving water.
+For a geostrophically balanced, anti-cyclonic circulation, we would expect a sea-surface low at the pivot point between the offshore moving water and the onshore moving water.
 That would lead to a local high in the isopycnal heights.
-Considering Fig.\ 12, which has this information, we think its pretty hard to see that occurring against the larger-scale backdrop of the isopycnals sloping up onto the shelf, and with the complication of internal tide heaving, which is considerable at this location.
-Similarly considering Fig 6, it would be hard to say the 26.4 isopycnal is much shallower at 28 km than 42 km.
-"Along isobath" also is pretty hard to define in this location.
-Frankly its not even clear how much ADCP data we would have needed to collect to capture this feature adequately.
+Considering Fig.\ 12, which has this information, we think its pretty hard to see that occurring against the larger-scale backdrop of the isopycnals sloping up onto the shelf.
+Add to that the complication of internal tide heaving, which is considerable at this location, and this analysis proves quite challenging.
+Similarly considering Fig 6, it would be hard to say the 26.4 isopycnal is much shallower at 28 km than at 42 km.
+``Along isobath'' also is pretty hard to define in this location.
+Even if we had an ADCP, we suspect it would have been hard to see this clearly.
 \end{reply}
@@
 
-    In response to this comment, we have changed the isopycnal to a deeper one (25.8) which is in the "minty" water.  We think it shows the three-dimensionality pretty well.
+    In response to this comment, we have changed the isopycnal to a deeper one (25.8) which is in the ``minty'' water.  We think it shows the three-dimensionality pretty well.
 
@@
 
-    Finally we decided to add all the data, rather than a subset to this plot.  There were some subsequent surveys excluded from the first that help solidify the case that the tongue is there.
+    Finally we decided to add all the data, rather than a subset to this plot.
+There were some subsequent surveys excluded from the first version of the plot that help solidify the case that the tongue is there.
 \end{reply}
@@
 \begin{reply}
-    Agreed that there is upwelling, and the upwelling tends to push water up the shelf.
+    We agree that there is upwelling, and that this upwelling tends to push water onto the shelf.
 However, its not clear why that would be stronger here than further upstream (poleward) where the water has had time to partially mix.
-Further, simple upwelling ageostrophically brings water up at the bottom, whereas this is happening mid-depth.
-If anything, you'd expect the secondary circulation associated with upwelling to push mid-depth water offshore?
+Further, coastal upwelling  brings water up at shelf-break depth, whereas this is happening at mid-depth.
+If anything, you'd expect the secondary circulation associated with upwelling to push mid-depth water offshore.
+
+    To help clarify this, we have added to the end of the second paragraph of section 3.3:
+    \begin{lstlisting}
+        The high-spice water comes onto the shelf through much of the water column, so is
+        not just being carried in a near-bottom Ekman layer by wind driven upwelling.
+    \end{lstlisting}
 \end{reply}
@@
         This is consistent with findings on the Oregon shelf where physical
-        processes are thought to account for 55-70\% of changes in dissolved
-        oxygen concentrations \cite{adamsetal13}.  Definitively identifying the
-        exchange mechanisms into the \Eddy\ is challenging from this data set.
-        The offshore front does not appear to have much exchange, however as
-        noted above there does appear to be strong tidal flows and mixing over
-        the submarine banks that can transport more oxygenated water into the
-        \Eddy.  Further, the Vancouver Island coastal current is oxygen rich,
-        and has a much less sharp front than the offshore front
-        (\fref{fig:Frontsurvey}).
+        processes are thought to account for 55-70\% of changes in dissolved oxygen
+        concentrations \cite{adamsetal13}.  Definitively identifying the exchange
+        mechanisms into the \Eddy\ is challenging from this data set.  The offshore front
+        does not appear to have much exchange, however as noted above there does appear
+        to be strong tidal flows and mixing over the submarine banks that can transport
+        more oxygenated water into the \Eddy.  Further, the Vancouver Island Coastal
+        Current is oxygen-rich and has a much less sharp front than the offshore front
+        (\fref{fig:Frontsurvey}), and is a further source of oxygen to the \Eddy.
     \end{lstlisting}
@@
     results section.
-Second, what is meant by "found in $10^{-4}$ of the water column?" Perhaps state a percentage instead.
+Second, what is meant by ``found in $10^{-4}$ of the water column?'' Perhaps state a percentage instead.
 Lastly, the conclusion is that the vigorous mixing contributes to the homogenization of the EDDY water onto a mixing line, rather than the alternative long residence time (with weaker mixing).
@@
     Ooops sorry - we decided to be more conservative in the results section, but left the inconsistency here.
-In both sections we've changed the numbers to reflect a relatively broad range $0.5-5 \mathrm{m^2\,s^{-1}}$, given the looseness of any of these estimates, and we changed the fractions to percentages.
+In both sections we've changed our estimates to be a relatively broad range $0.5-5 \mathrm{m^2\,s^{-1}}$, given the looseness of any of these estimates, and we changed the fractions to percentages, as suggested.
 \end{reply}
@@
 
-Its hard to say more, but we suspect that the model has a stronger mesoscale eddy field than observed, and these generate substantial cross-shore motions.
+Its hard to say more, but we suspect that the model has a more energetic mesoscale eddy field than observed, and these eddies generate substantial cross-shore motions.
 It would be a great project to try and test why the regional simulation is so much more advective, but beyond our scope here.
@@
 \begin{point}
-Abstract: use of "recirculation" is odd. Recirculating with respect to what? Why not just use the term "eddy?"
+Abstract: use of ``recirculation'' is odd. Recirculating with respect to what? Why not just use the term ``eddy?''
 \end{point}
@@
 A stagnant region between two currents is eddy-like, but the analogy with a freely propagating eddy seems strained.
-Regardless, we use recirculation once now in the abstract, and then drop it for "eddy" since that is how this feature is referred to in previous literature.
+Regardless, we use recirculation once now in the abstract, and then drop it for ``eddy'' since that is how this feature is referred to in previous literature.
 \end{reply}
@@
 \shortpoint{Line 234: 48.5N, not 44.5N}
-\shortreply[]{Thanks}
+\shortreply[]{Thanks, corrected}
 
 \shortpoint{Fig 12: label Barclay Canyon}
-\shortreply[]{We tried this, but its pretty hard to do without obscuring data. Barkley is labelled in Figure 1.  }
+\shortreply[]{We tried this, but its pretty hard to do without obscuring data. Barkley Canyon is labelled in Figure 1.  }
 
 \shortpoint{References: many uncapitalized place names (e.g., U.S. states)}
-\shortreply[]{Thanks}
+\shortreply[]{Thanks, corrected.}
 
@@
 This is probably worth the extra space, and we hope that it helps with legibility.
-Added the La Perouse Buoy to Fig 1c.
-Hopefully this helps, though with such a complicated study region it will remain challenging.
+We added the La Perouse Buoy to Fig 1c.
+Hopefully this helps, though with such a complicated study region and diverse data sets, it will remain challenging.
 \end{reply}
@@
     The text is largely well-written, but some mostly stylistic choices make sentences difficult to parse.
-Avoid the construction of starting sentences with "This", as it leads to loss of specificity.
-For example, 78-83, in which "This" could refer to the inference, the assumption, the OMZ, etc.
-Or 250: "this" could mean ageostrophy or movement.
+Avoid the construction of starting sentences with ``This'', as it leads to loss of specificity.
+For example, 78-83, in which ``This'' could refer to the inference, the assumption, the OMZ, etc.
+Or 250: ``this'' could mean ageostrophy or movement.
 Other examples where the wording is unclear are given in the comments and questions below.
@@
 \begin{reply}
-    Thanks cleaned this up in both these spots, and a few other places.
-Agree that using "this" at the beginning of a sentence should be clear and unambiguous.
-On the other hand, repeating a phrase from the previous sentence (or worse rephrasing it) is harder to parse.
+    Thanks cleaned this up in a few other places.
+We gree that using ``this'' at the beginning of a sentence should be clear and unambiguous.
+On the other hand, repeating a phrase from the previous sentence (or worse rephrasing it) can be harder to parse.
 \end{reply}
@@
     Thanks for the careful read of this.  We think we have made the location
-    names consistent now.  Note that only "Eddy" gets capitalized by itself;
-    for the banks and canyons, we capitalized if it was the full name, and left
-    lower case if referring to "the canyon".
+    names consistent now.  Note that only ``Eddy'' gets capitalized by itself;
+    for the banks and canyons, we capitalized if it was used as a proper noun,
+    and left lower case if referring to a generic, e.g. ``the canyon''.
 
@@
     googling this subject indicates that both are in common usage to describe
-    distances.  (The lead author will admit thinking he was illiterate between moves from Canada, to the US, and back again in his spelling of ``modeling/modelling'')
+    distances.  (The lead author will admit thinking he was illiterate between
+    moves from Canada, to the US, and back again in his spelling of
+    ``modeling/modelling'')
 \end{reply}
@@
 \begin{reply}
-    Thanks changed to:
+    Thanks. We changed these lines to:
     \begin{lstlisting}
@@
 
-\shortreply[]{Fixed}
+\shortreply[]{Thanks, this is fixed.}
 
@@
 \begin{reply}
-    Indeed - split.  The second sentence is still rather long, but is a list, so should be easy enough to parse, but agree that two introductory clauses and a list was very hard to read.
+    Indeed, we have now split it.  The second sentence is still rather long,
+    but is a list, so should be easy enough to parse.
 \end{reply}
@@
 \begin{reply}
-    TODO: get Whitney reference.
+    They are generally about 2-4\%.  We are using oxygen very qualitatively in this paper, so we have removed the discussion of the recalibration that is routinely done.
 \end{reply}
@@
 
+\begin{reply}
+    We kept this the same just for ease of skimming by readers.
+The lack of an ADCP was very disappointing, and its something we expect readers will be looking for.
+Hence the standalone paragraph to make it easy to find.
+\end{reply}
+
 \begin{point}
@@
 \begin{reply}
-    Rephrased to:
+    Thanks. Rephrased to:
 
@@
         Hydrographic sections along the LB and LC hydrographic lines highlight
-        summer conditions on the Southern Vancouver Island Shelf and presented
+        summer conditions on the Southern Vancouver Island Shelf and are presented
         to give context to the more detailed survey carried out on the
@@
 \begin{reply}
-    rephrased to
+    Thanks, rephrased to:
 
     \begin{lstlisting}
-        The water upstream of the \Eddy\ region, along the LC line, is less
-        mixed than the \Eddy\ water (\fref{fig:LaPerouse2013Ctd}, top three
-        rows) and has higher oxygen.
+        The water upstream of the \Eddy\ region, along the LC line, is less mixed and has higher oxygen than the \Eddy\ water (\fref{fig:LaPerouse2013Ctd}, top three rows).
+Based on these properties, offshore water does not appear to make it over La Perouse Bank into the onshore basin, or if it does, it does so intermittently and with substantial mixing.
     \end{lstlisting}
@@
 Probably worth adding to section 2?
-Also helpful to reference the 26.4 kg/m3 isopycnal as magenta in the text here, especially if it is marked in magenta the rest of the paper.
+Also helpful to reference the $26.4\ \mathgrm{kg\,m^{-3}}$ isopycnal as magenta in the text here, especially if it is marked in magenta the rest of the paper.
 \end{point}
@@
 \begin{reply}
-    Thanks, there is a new second paragraph in the intro explaining the co-ordinate system.
+    Thanks. There is a new second paragraph in the Introduction explaining the co-ordinate system.
+
+    \begin{lstlisting}
+        We use an along/across-shelf co-ordinate system with an origin at
+        48.4\textdegree N and 125.6\textdegree W, with the across-shelf $x$ oriented
+        39 degrees north of East (\fref{fig:LocMapBoth}b, grey/white alternating
+        lines).  The projection is Cartesian with a central latitude at
+        48.4\textdegree N, which is sufficient for the limited geographic extents
+        discussed here.
+    \end{lstlisting}
+
+    The $26.4$ isopyncal is referenced in the text.
+
+    \begin{lstlisting}
+        Note that the water along the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal (\fref{fig:LaPerouse2013Ctd}, magenta contours) becomes cooler...
+    \end{lstlisting}
+
 \end{reply}
@@
 \begin{point}
-    187: values can be easier to follow than color names (white is okay but defining what "pink" gets harder later on, after the descriptions of Figs 6-8)
+    187: values can be easier to follow than color names (white is okay but defining what ``pink'' gets harder later on, after the descriptions of Figs 6-8)
 \end{point}
@@
 \begin{reply}
-  Thanks, great suggestion.
+  Thanks, great suggestion.  We now identify the water masses by values of spice anomaly.
 \end{reply}
@@
 \begin{reply}
-    The other reviewer had a similar suggestion - we have instead made a PDF of $gamma$ along 26.4, and identified the peaks.
-Readers may be more comfortable with 1-D PDFs anyhow, so this seems a good way to get this across.
+    The other reviewer had a similar suggestion - we have made a PDF of $gamma$ along 26.4, and identified the peaks.
+Readers may be more comfortable with 1-D PDFs anyhow, so this seems a good way to clarify how we are definiting the three water masses.
 \end{reply}
@@
         These lines are made with a two-dimensional interpolation where MVP
-        casts at a given depth have a Gaussian weight $w = \exp^{-(r/r_0)^2}$,
+        casts at a given depth are given a Gaussian weight $w = \exp^{-(r/r_0)^2}$,
         where $r$ is the distance from the line and $r_0 = 1.5\ \mathrm{km}$.
@@
 \begin{reply}
-    Thanks, changed to "transition".
+    Thanks, changed to ``transition''.
 \end{reply}
@@
 \begin{reply}
-    Thanks "higher spice water" is now added.
+    Thanks ``higher spice water'' is now added.
 \end{reply}
@@
     That is unfortunate because it is one of the main points of the paper!
-The new version of Fig 12 has all the data, which we probably should have used in the original submission because hopefully it makes the tongue more clear.
-We also now refer to the panel explicitly.
-We could annotate it, but have refrained for the present.
+The new version of Fig 12 has all the data, which we probably should have used in the original version because hopefully it makes the tongue more clear.
+We also now refer to this panel explicitly in the text.
+We could annotate it, but the plot is already relatively dense.
 \end{reply}
@@
 \begin{reply}
-    Thanks, done. Also changed the km markings to the turn locations and added ticks on panel d corresponding to the tick locations.
+    Thanks, done. We also changed the km markings to the turn locations and added ticks on panel d corresponding to the tick locations.
 \end{reply}
@@
 \begin{reply}
-    Fixed - it was the MVP data, though the previous version was a subsample.
+    Fixed; it was the MVP data, though the previous version was a subsample of the full dataset.
 \end{reply}
@@
 \begin{reply}
-    Thanks, "suggesting" is definitely better here.
+    Thanks, ``suggesting'' is definitely better here.
 \end{reply}
@@
 \begin{reply}
-    Thanks.
+    Thanks; changed as suggested.
 \end{reply}
@@
 \begin{reply}
-    Yes, but perhaps better shown in Fig 3e, k, which are now explicitly referenced.
+    Yes, but perhaps better shown in Fig 3e, k, which are now explicitly referenced here.
 \end{reply}
@@
 \begin{reply}
-    This was just a typical weak coastal current scaling.  We now cite Thomson and Kassovski and point out that this current could be substantially stronger.
+    This was just a typical weak coastal current scaling.  We now cite Thomson and Kassovski (2015), and point out that this current could be substantially stronger.
 \end{reply}

### 2023-05-25 R1/PW13Paper.tex
R1: submitted

@@
 A second goal of our study was to better understand cross-shore exchange between shelf and offshore water.
-For a two-dimensional topography with wind forcing, cross-shore transport is supplied by Ekman layers.
-However, in many locations there is also evidence of eddies, meanders and filaments driving wholesale separation of shelf currents into the deep ocean.
+For a two-dimensional topography with along-shore wind forcing, cross-shore transport is supplied by Ekman layers.
+However, in many locations there is also evidence of eddies, meanders, and filaments driving wholesale separation of shelf currents into the deep ocean.
 These have been well-studied off California \cite{strubetal91} where satellite and \emph{in-situ} observations show large meanders of coastal upwelling currents leading to coastal water injected into the offshore domain.
@@
 48.4\textdegree N and 125.6\textdegree W, with the across-shelf $x$ oriented 39 degrees north of East (\fref{fig:LocMapBoth}b, grey/white alternating lines).
-The projection is Cartesian with a central latitude at 48.4\textdegree N, which is fine for the limited geographic extents discussed here.
+The projection is Cartesian with a central latitude at 48.4\textdegree N, which is sufficient for the limited geographic extents discussed here.
 
@@
 The data comes from a lowered Seabird 9-11 CTD, with an SBE 43 oxygen sensor.
-Oxygen data have been corrected against bottle casts, and the CTD corrected for sensor offsets and thermal lags.
 
@@
 
-Hydrographic sections along the LB and LC hydrographic lines highlight summer conditions on the Southern Vancouver Island Shelf and presented to give context to the more detailed survey carried out on the \emph{R/V Falkor} (\Fref{fig:LaPerouse2013Ctd}).
+Hydrographic sections along the LB and LC hydrographic lines highlight summer conditions on the Southern Vancouver Island Shelf and are presented to give context to the more detailed survey carried out on the \emph{R/V Falkor} (\Fref{fig:LaPerouse2013Ctd}).
 During both surveys, and along both lines, there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m depth offshore to shallower than 85 m over the shelf.
@@
 
-The water upstream of the \Eddy\ region, along the LC line, is less mixed than the \Eddy\ water (\fref{fig:LaPerouse2013Ctd}, top three rows) and has higher oxygen.
+The water upstream of the \Eddy\ region, along the LC line, is less mixed and has higher oxygen than the \Eddy\ water (\fref{fig:LaPerouse2013Ctd}, top three rows).
 Based on these properties, offshore water does not appear to make it over La Perouse Bank into the onshore basin, or if it does, it does so intermittently and with substantial mixing.
@@
 values (\fref{fig:LaPerouse2013Ctd}, middle rows), with a positive-anomaly
-layer ($\gamma \approx 0.2 \mathrm{kg\,m^{-3}}$, red colors) centered at $26.4\
+layer ($\gamma \approx 0.2\ \mathrm{kg\,m^{-3}}$, red colors) centered at $26.4\
 \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and
@@
 In synthetic cross sections of spice anomaly, the \Eddy\ is clearly identifiable as having low spice anomaly ($\gamma \approx 0\ \mathrm{kg\,m^{-3}}$, \fref{fig:CrossSectionsSpice}c, d).
-These lines are made with a two-dimensional interpolation where MVP casts at a given depth have a Gaussian weight $w = e^{-(r/r_0)^2}$, where $r$ is the distance from the line and $r_0 = 1.5\ \mathrm{km}$.
+These lines are made with a two-dimensional interpolation where MVP casts at a given depth are given a Gaussian weight $w = e^{-(r/r_0)^2}$, where $r$ is the distance from the line and $r_0 = 1.5\ \mathrm{km}$.
 Note that the section at   The low-spice anomaly water extends from the seafloor to approximately 30 m depth, onshore of $X \approx 0\ \mathrm{km}$.
@@
 Whatever causes it, downstream of the separation, the water is replaced by offshore water with much higher spice values.
+The high-spice water comes onto the shelf through much of the water column, so is not just being carried onshelf by near-bottom Ekman layer by wind driven upwelling.
 
@@
 The offshore front does not appear to have much exchange, however as noted above there does appear to be strong tidal flows and mixing over the submarine banks that can transport more oxygenated water into the \Eddy.
-Further, the Vancouver Island coastal current is oxygen rich, and has a much less sharp front than the offshore front (\fref{fig:Frontsurvey}).
+Further, the Vancouver Island Coastal Current is oxygen-rich and has a much less sharp front than the offshore front (\fref{fig:Frontsurvey}), and is a likely source of oxygen to the \Eddy.
 
@@
 Rather it seems likely that the topographic change engendered by the sudden turn to the east of La Perouse Bank catalyses a larger scale instability at this location.
-In California, most of the cold filaments observed appear to be catalyzed by headlands and underwater topography \cite{strubetal91}, though modelling studies find instabilities are possible even in two-dimensional flows \cite{pierceetal91}.
+In California most of the cold filaments observed appear to be catalyzed by headlands and underwater topography \cite{strubetal91}, though modelling studies find instabilities are possible even in two-dimensional flows \cite{pierceetal91}.
 \citeA{durskiallen05}, when modelling the Oregon shelf, found that including realistic shelf bathymetry catalyzed intermittent large-scale instabilities, a finding that is deemed likely to apply at other coastal locations \cite{batteen97}.
@@
 
-\bibliography{main}
+\bibliography{pw13}
 
